---
title: "Regressions for paper_no MI"
author: "Ale Garcia"
date: "2025-04-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rio)
library(here)
library(tidyverse)
library(haven)
library(janitor)
library(knitr)
library(surveytoolbox)
library(sjPlot)
library(kableExtra)
library(psych)
library(sjmisc)
library(MVN)
library(stats)
library(AICcmodavg)
library(rsq)

theme_set(theme_minimal())
```


### The dataset ###

```{r}
d <- read_sav(here("outcomes paper/data_outcomes", "p_y_w1_for_analysesjan25.sav")) # 94 rows

# view_df(d)
```


```{r}
d1 <- d %>%
  select(school_id, family_id, participant_id, parent_eng_comf, youth_age, youth_gender, acadhwi_1, checkhwi_1, checkhwi_1, struct_1, convfut_1, knowschool_1, selfeffic_1, parnets_1, welcome_1, endorse_1, supported_1, ptrel_1, y_sch_eng, y_parsupp, y_parinv, y_hw, gen_grades, grades)

# view_df(d1)
```


```{r}
d2 <- d1 %>%
  mutate(school_id = factor(school_id),
         parent_eng_comf = as.numeric(parent_eng_comf),
         youth_gender = factor(youth_gender),
         youth_age = as.numeric(youth_age))

```


```{r}
d3 <- d2 %>%
  mutate(across(where(is.numeric), ~ scale(.)[,1]))
d3
```

### School engagement - 10 rows deleted

With MI: checkhwi_1  beta = 0.24, p < 0.01
```{r, warning=FALSE}
fit_engage <- lm(y_sch_eng ~ parent_eng_comf + youth_age + youth_gender + acadhwi_1 + checkhwi_1 + struct_1 + convfut_1 + knowschool_1 + parnets_1 + welcome_1 + endorse_1 + supported_1 + ptrel_1, d3)

summary(fit_engage) 

partial_r2 <- rsq.partial(fit_engage)
print(partial_r2)
```

### Youth-rep grades - 11 rows deleted

With MI:
youth being female: beta = -0.48, p = 0.03
convfut_1: beta = -0.26, p = 0.04
```{r, warning=FALSE}
fit_grades <- lm(grades ~ parent_eng_comf + youth_age + youth_gender + acadhwi_1 + checkhwi_1 + struct_1 + convfut_1 + knowschool_1 + parnets_1 + welcome_1 + endorse_1 + supported_1 + ptrel_1, d3)

summary(fit_grades) 

partial_r2 <- rsq.partial(fit_grades)
print(partial_r2)
```

### Youth-hw practices - 10 rows deleted

With MI:
no predictors were significantly related with the outcome
```{r, warning=FALSE}
fit_yhw <- lm(y_hw ~ parent_eng_comf + youth_age + youth_gender + acadhwi_1 + checkhwi_1 + struct_1 + convfut_1 + knowschool_1 + parnets_1 + welcome_1 + endorse_1 + supported_1 + ptrel_1, d3)

summary(fit_yhw) 

partial_r2 <- rsq.partial(fit_yhw)
print(partial_r2)
```

### Youth-parent support - 10 rows deleted

With MI:
knowschool_1  beta = 0.41 p = 0.02
```{r, warning=FALSE}
fit_psupp <- lm(y_parsupp ~ parent_eng_comf + youth_age + youth_gender + acadhwi_1 + checkhwi_1 + struct_1 + convfut_1 + knowschool_1 + parnets_1 + welcome_1 + endorse_1 + supported_1 + ptrel_1, d3)

summary(fit_psupp) 

partial_r2 <- rsq.partial(fit_psupp)
print(partial_r2)
```

### Youth-parent involvement - 10 rows deleted

With MI:
parent_eng_comf beta = -0.29, p = 0.01
knowschool_1  beta = 0.30, p = 0.05
```{r, warning=FALSE}
fit_pinv <- lm(y_parinv ~ parent_eng_comf + youth_age + youth_gender + acadhwi_1 + checkhwi_1 + struct_1 + convfut_1 + knowschool_1 + parnets_1 + welcome_1 + endorse_1 + supported_1 + ptrel_1, d3)

summary(fit_pinv) 

partial_r2 <- rsq.partial(fit_pinv)
print(partial_r2)
```

### Parent self-efficacy - 10 rows deleted

With MI:
parent_eng_comf beta = 0.16, p = 0.03
welcome_1 beta = 0.26, p = 0.02
endorse_1  beta = 0.46, p < 0.001
```{r, warning=FALSE}
fit_efficacy <- lm(selfeffic_1 ~ parent_eng_comf + youth_age + youth_gender + acadhwi_1 + checkhwi_1 + struct_1 + convfut_1 + knowschool_1 + parnets_1 + welcome_1 + endorse_1 + supported_1 + ptrel_1, d3)

summary(fit_efficacy) 

partial_r2 <- rsq.partial(fit_efficacy)
print(partial_r2)
```

### Parent-rep grades - 13 rows deleted

With MI:
youth being female: beta = -0.48, p = 0.03
convfut_1: beta = -0.26, p = 0.04
```{r, warning=FALSE}
fit_pgrades <- lm(gen_grades ~ parent_eng_comf + youth_age + youth_gender + acadhwi_1 + checkhwi_1 + struct_1 + convfut_1 + knowschool_1 + parnets_1 + welcome_1 + endorse_1 + supported_1 + ptrel_1, d3)

summary(fit_pgrades) 

format(summary(fit_pgrades)$coefficients, scientific = FALSE) 

partial_r2 <- rsq.partial(fit_pgrades)
print(partial_r2)
```

#### Select variables of interest
```{r}
vars_of_interest <- d3[, c("parent_eng_comf", "youth_age", "acadhwi_1", "checkhwi_1", "struct_1", "convfut_1", "knowschool_1", "parnets_1", "welcome_1", "endorse_1", "supported_1", "ptrel_1", "y_sch_eng", "y_parsupp", "y_parinv", "y_hw", "grades", "gen_grades", "selfeffic_1")] # removed youth gender because is a factor
```


#### Convert to long format for ggplot
```{r}
long_data <- pivot_longer(vars_of_interest, cols = everything(), names_to = "Variable", values_to = "Value")
```


#### Plot histograms
```{r}
ggplot(long_data, aes(x = Value)) + 
  geom_histogram(bins = 30, fill = "skyblue", color = "black", alpha = 0.7) + 
  facet_wrap(~ Variable, scales = "free") +  # One plot per variable
  theme_minimal() + 
  labs(title = "Distribution of Variables")
```