---
title: "Juntos Project Initial Report"
subtitle: "Baseline Assessment Cleaning and Recommendations"
author: "Alejandra Garcia Isaza"
date: "March 2021"
output: 
  pagedreport::paged_windmill:
    front_img: "C:/Users/alega/Documents/1. GRADUATE SCHOOL/Academics/Year IV/EDLD 609/juntos/front.PNG"
    logo: "https://mk0rfortheresto0o08q.kinstacdn.com/wp-content/uploads/2020/08/rru-logo-blue.svg"
    img_to_dark: TRUE
    logo_to_white: TRUE
knit: pagedown::chrome_print
main-color: "#65955E"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Juntos Project Description {-}

## Study and Intervention Details {-}

Say something about the project.


# Data Cleaning procedures {-}

I spent a loooot of time cleaning!

## Educator's dataset {-}

...

## Parent dataset {-}

...

## Youth dataset {-}

...

```{r}

```


# Participant descriptives  {-}

Say something about participant characteristics

## Educator's characteristics {-}

...

## Parent characteristics {-}

...

## Youth characteristics {-}

...


# Scale creation and Testing {-}

Say something about scales

## Educator's scales {-}

...

## Parent scales {-}

...

## Youth scales {-}

...

# Recommendations  {-}

I recommend...