---
title: "creating dataset_youth"
output: html_document
date: "2025-01-14"
---
This script is menat to be added on while I identify additional variables of interest. initially created on jan.12.25


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rio)
library(here)
library(tidyverse)
library(haven)
library(janitor)
library(knitr)
library(surveytoolbox)
library(sjPlot)
library(kableExtra)
library(psych)
library(sjmisc)
library(MVN)

theme_set(theme_minimal())
```

# youth w1
```{r}
youth_w1_raw <- read_sav(here("outcomes paper/data_outcomes", "all_youth_wave1.sav")) # it has 88, 99, -99, etc.

youth_w1 <- youth_w1_raw %>%
  select(school_id, condition, id, wave, q2, q4, q7:q13, q18:q22, q36:q40, q83:q91, q98, q231, q232_2, q232_3, q232_99) %>%
   rename(c("family_id" = "id"),
         c("grade_lev" = "q2"),
         c("grades" = "q4"),
         c("hw_1" = "q7"),
         c("hw_2" = "q8"),
         c("hw_3" = "q9"),
         c("hw_4" = "q10"),
         c("hw_5" = "q11"),
         c("hw_6" = "q12"),
         c("hw_7" = "q13"),
         c("pi_1" = "q18"),
         c("pi_2" = "q19"),
         c("pi_3" = "q20"),
         c("pi_4" = "q21"),
         c("pi_5" = "q22"),
         c("pi_6" = "q36"),
         c("pi_7" = "q37"),
         c("pi_8" = "q38"),
         c("pi_9" = "q39"),
         c("pi_10" = "q40"),
         c("se_1" = "q83"),
         c("se_2" = "q84"),
         c("se_3" = "q85"),
         c("se_4" = "q86"),
         c("se_5" = "q87"),
         c("se_6" = "q88"),
         c("se_7" = "q89"),
         c("se_8" = "q90"),
         c("se_9" = "q91"),
         c("youth_eng_comf" = "q98"),
         c("youth_gender" = "q231"),
         c("y_latino" = "q232_2"),
         c("y_indigenous_americas" = "q232_3"),
         c("y_race_ethnicity_no_response" = "q232_99")) 

view_df(youth_w1_raw) 
```

# Data prep: recoding missing variables as N/A
```{r include=FALSE}
# recoding missing values as N/A with function

# vector with missing values in dataset
missing_vals <- c(0, 88, 99, -99) # note: very specific to student dataset

# function that returns true if values in vector are equal to missing_vals. The function takes a vector x, and specified values of missing data
recode_missing <- function(x, missing_vals = c(0, 88, 99, -99)) {
  test <- x %in% missing_vals
  ifelse(test, NA, x)
}

# function that recodes missing values to NA. The function takes a dataframe with variables with missing data, and specified values of missing data
recode_missing_df <- function(df, missing_vals = c(0, 88, 99, -99)) {
  modify(df, ~recode_missing(.x, missing_vals)) # here uses the function created above
}

youth_w1_2 <- recode_missing_df(youth_w1) # the function strips out variable labels
```

# Data prep: reverse scoring negatively worded items
```{r}
youth_w1_3 <- youth_w1_2 %>%
  mutate(se_2 = likert_reverse(se_2, top = 5, bottom = 1),
         se_4 = likert_reverse(se_4, top = 5, bottom = 1),
         se_5 = likert_reverse(se_5, top = 5, bottom = 1))

view_df(youth_w1_3) 
```

```{r}
youth_hw <- youth_w1_3 %>%
  select(hw_2, hw_3, hw_4, hw_5, hw_6, hw_7)

alpha(youth_hw) # 0.83

# Poly corr matrix
poly <- polychoric(youth_hw)
cor.plot(poly$rho, numbers = T, upper = F, main = "Polychoric", show.legend = F)
```

*youth's perception of parental support (in the last 3 months...) 
q18 - find ways to take part in school activities (likert: 1-4)
q19 - find ways to do volunteer activities in the community (likert: 1-4)
q20 - take leadership roles in school or in the community (likert: 1-4)
q21 - do your best in school (likert: 1-4)
q22 - make a schedule for completing homework and school projects on time (likert: 1-4)
```{r}

youth_par_sup <- youth_w1_3 %>%
  select(pi_1, pi_2, pi_3, pi_4, pi_5)

alpha(youth_par_sup) # 0.72

# Poly corr matrix
poly <- polychoric(youth_par_sup)
cor.plot(poly$rho, numbers = T, upper = F, main = "Polychoric", show.legend = F)

```
*youth's perception of parental involvement (in the last 3 months...) 
q36.	… visited your school for a special event (activity nights, choir performance, dances, etc.) (likert: 1-4)
q37.	… attended a parent-teacher conference (likert: 1-4)
q38.	… attended a parent group meeting (likert: 1-4)
q39.	…volunteered for a school activity, event, or project (likert: 1-4)
q40.	… had at least some type of contact with at least one of your teachers (by telephone, text, email, or an in-person meeting)? (likert: 1-4)

```{r}
# youth_par_inv <- youth_w1_3 %>%
#   select(pi_6, pi_7, pi_8, pi_9, pi_10) # item 9 very low cor with other items
# 
# alpha(youth_par_inv) # 0.674
# 
# youth_par_inv <- youth_w1_3 %>%
#   select(pi_6, pi_7, pi_8, pi_9)
# 
# alpha(youth_par_inv) # 0.669

youth_sch_inv <- youth_w1_3 %>%
  select(pi_6, pi_7, pi_8, pi_10) # item 9 very low cor with other items

alpha(youth_sch_inv) # 0.667

# Poly corr matrix
poly <- polychoric(youth_sch_inv)
cor.plot(poly$rho, numbers = T, upper = F, main = "Polychoric", show.legend = F)
```
# outputing youth wave 1 dataset 
```{r}
# youth_w1_3 %>%
#   haven::write_sav(here("outcomes paper/data_outcomes", "youth_w1_jan2025.sav"))
```
