---
title: "Creating dataset"
author: "Alejandra Garcia Isaza"
date: "2023-03-20"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rio)
library(here)
library(tidyverse)
library(haven)
library(janitor)
library(knitr)
library(surveytoolbox)
library(sjPlot)
library(kableExtra)
library(psych)
library(sjmisc)
library(MVN)

theme_set(theme_minimal())
```

##### PARENT WAVE 1 #####

# Loading the dataset
```{r}
d_par_w1_raw <- read_sav(here("outcomes paper/data_outcomes", "all_parent_wave1.sav")) # This dataset has ALL parents in wave 1 (n = 124)
```

# filtering primary caregiver only (wave 1) - TBD if removing parent with less data waves among parents is ok
```{r}
d_par_w1 <- d_par_w1_raw %>%
  filter(participant_id != "405_1") %>% # removing mother who has less waves of data than father
  filter(participant_id != "614_1") %>% # removing mother who has less waves of data than father
  distinct(family_id, .keep_all = TRUE) # 95 primary caregivers
```

# Number of participants types in Wave 1 primary caregiver dataset
```{r}
# 95 total primary caregiver participants in w1, from which...

d_par_w1 %>%
  filter(participant == 1) # 89 were moms

d_par_w1 %>%
  filter(participant == 2) # 5 were dads

d_par_w1 %>%
  filter(participant == 4) # 1 were "other" (grandparent)

unique(d_par_w1$family_id) # 95 unique id values, thus 95 families in W1. 
```

```{r}
view_df(d_par_w1) # always load this one to check variable labels
```

# Data prep: recoding missing variables as N/A
```{r include=FALSE}
# recoding missing values as N/A with function

# vector with missing values in dataset
missing_vals <- c(77, 88, 99, -99)

# function that returns true if values in vector are equal to missing_vals. The function takes a vector x, and specified values of missing data
recode_missing <- function(x, missing_vals = c(77, 88, 99, -99)) {
  test <- x %in% missing_vals
  ifelse(test, NA, x)
}

# function that recodes missing values to NA. The function takes a dataframe with variables with missing data, and specified values of missing data
recode_missing_df <- function(df, missing_vals = c(77, 88, 99, -99)) {
  modify(df, ~recode_missing(.x, missing_vals)) # here uses the function created above
}

d1_par_w1 <- recode_missing_df(d_par_w1) # the function strips out variable labels
```

# Data prep: reverse scoring negatively worded items
```{r}
d2_par_w1 <- d1_par_w1 %>%
  mutate(q82 = likert_reverse(q82, top = 4, bottom = 1),
         q83 = likert_reverse(q83, top = 4, bottom = 1),
         q131 = likert_reverse(q131, top = 4, bottom = 1),
         q133 = likert_reverse(q133, top = 4, bottom = 1),
         q136 = likert_reverse(q136, top = 4, bottom = 1))
```

# checking reverse scoring
```{r}
d1_par_w1 %>%
  select(q136)

d2_par_w1 %>%
  select(q136)
```

# outputing primary caregiver parent wave 1 dataset 
```{r}
# d2_par_w1 %>%
#   haven::write_sav(here("outcomes paper/data_outcomes", "primary_parent_w1.sav"))
```

##
##
##

##### PARENT WAVE 2 #####

# Loading the dataset
```{r}
d_par_w2_raw <- read_sav(here("outcomes paper/data_outcomes", "all_parent_wave2.sav")) # This dataset has ALL parents in wave 2 (n = 108)
```

# filtering primary caregiver only (wave 1) - TBD if removing parent with less data waves among parents is ok
```{r}
d_par_w2 <- d_par_w2_raw %>%
  filter(participant_id != "614_1") %>% # removing mother who has less waves of data than father
  distinct(family_id, .keep_all = TRUE) # 84 primary caregivers
```

# Number of participants types in Wave 2 primary caregiver dataset
```{r}
# 84 total primary caregiver participants in w2, from which...

d_par_w2 %>%
  filter(participant == 1) # 80 were moms

d_par_w2 %>%
  filter(participant == 2) # 3 were dads

d_par_w2 %>%
  filter(participant == 4) # 1 were "other" (grandparent)

unique(d_par_w2$family_id) # 84 unique id values, thus 84 families in W1. 
```

```{r}
view_df(d_par_w2) # always load this one to check variable labels
```

# Data prep: recoding missing variables as N/A
```{r include=FALSE}
d1_par_w2 <- recode_missing_df(d_par_w2) # the function strips out variable labels
```

# Data prep: reverse scoring negatively worded items
```{r}
d2_par_w2 <- d1_par_w2 %>%
  mutate(q82 = likert_reverse(q82, top = 4, bottom = 1),
         q83 = likert_reverse(q83, top = 4, bottom = 1),
         q131 = likert_reverse(q131, top = 4, bottom = 1),
         q133 = likert_reverse(q133, top = 4, bottom = 1),
         q136 = likert_reverse(q136, top = 4, bottom = 1))
```

# checking reverse scoring
```{r}
d1_par_w2 %>%
  select(q136)

d2_par_w2 %>%
  select(q136)
```

# outputing primary caregiver parent wave 2 dataset 
```{r}
# d2_par_w2 %>%
#   haven::write_sav(here("outcomes paper/data_outcomes", "primary_parent_w2.sav"))
```

###
###
###

##### PARENT WAVE 3 #####

# Loading the dataset
```{r}
d_par_w3_raw <- read_sav(here("outcomes paper/data_outcomes", "all_parent_wave3.sav")) # This dataset has ALL parents in wave 3 (n = 108)
```

# filtering primary caregiver only (wave 1) - TBD if removing parent with less data waves among parents is ok
```{r}
d_par_w3 <- d_par_w3_raw %>%
  distinct(family_id, .keep_all = TRUE) # 86 primary caregivers
```

# Number of participants types in Wave 3 primary caregiver dataset
```{r}
# 86 total primary caregiver participants in w3, from which...

d_par_w3 %>%
  filter(participant == 1) # 80 were moms

d_par_w3 %>%
  filter(participant == 2) # 5 were dads

d_par_w3 %>%
  filter(participant == 4) # 1 were "other" (grandparent)

unique(d_par_w3$family_id) # 86 unique id values, thus 86 families in W1. 
```

```{r}
view_df(d_par_w3) # always load this one to check variable labels
```

# Data prep: recoding missing variables as N/A
```{r include=FALSE}
d1_par_w3 <- recode_missing_df(d_par_w3) # the function strips out variable labels
```

# Data prep: reverse scoring negatively worded items
```{r}
d2_par_w3 <- d1_par_w3 %>%
  mutate(q82 = likert_reverse(q82, top = 4, bottom = 1),
         q83 = likert_reverse(q83, top = 4, bottom = 1),
         q131 = likert_reverse(q131, top = 4, bottom = 1),
         q133 = likert_reverse(q133, top = 4, bottom = 1),
         q136 = likert_reverse(q136, top = 4, bottom = 1))
```

# checking reverse scoring
```{r}
d1_par_w3 %>%
  select(q136)

d2_par_w3 %>%
  select(q136)
```

# outputing primary caregiver parent wave 2 dataset 
```{r}
# d2_par_w3 %>%
#   haven::write_sav(here("outcomes paper/data_outcomes", "primary_parent_w3.sav"))
```


# need to create dataset with all variables of all waves to conduct CFA (to check time invariance)
```{r}
# Jaime's code to try
tidyr::pivot_longer(data=it2, cols= c("pre", "post"), names_to = "time", values_to = "val")
```

