---
title: "EFA for parent scales"
author: "Ale Garcia"
date: "2025-02-11"
output: html_document
---

Note: on may 5, 2025: I made a few changes on some scales, based on more painstakingly reviewing model assumptions for EFA. These changes are reflected in paper I worked with Seth. 

## From Flora, LaBrish, & Chalmers (2012)

Check collineraity: "refers to perfect or near-perfect linear relationships among observed variables. With multiple regression, the focus is on collinearity among explanatory variables, but with factor analysis, the concern is collinearity among dependent variables, that is, the set of variables being factor analyzed. When collinear variables are included, the product-moment correlation matrix R will be singular, or non-positive definite."

"Collinearity in factor analysis is relatively simple to diagnose: if any eigen values of a product-moment R equal zero or are negative, then R is non-positive definite and collinearity is present (and software will likely produce a warning message)."

Heywood cases: "Improper solutions (e.g.,a model solution with at least one negative estimated residual variance term, or“Heywoodcase”) are also likely to occur in the presence of one or more unusual cases (Bollen, 1987), which can lead to a researcher unwittingly revising a model or removing an observed variable from the analysis."

Multicollenearity will produce 1 of 2 things:
- aberrant factorial load
- negative variance (more for CFA) - non-positive definite matrix.
- Check for collinearity between items included on an EFA model using the correlation plot

### Rationale for factor extraction method and type of rotation.  
Since Juntos assessments variables targeted for EFA were ordinal (4- and 5-point Likert response scales) and EFA of Pearson product-moment correlation among polytomous variables can produce misleading results (Flora et al., 2012), a polychoric correlation approach was used in this dissertation. An unweighted least squares (ULS) estimation method was used for EFA as recommended by Flora et al. (2012). To determine the number of factors to retain in each scale, examination of scree plots, Kaiser’s rule (eigenvalues > 1), parallel analysis, and analysis of the pattern of factor loadings were considered. If more than one factor emerged in a given scale, an oblique quartimin rotation was used as suggested by Flora et al. (2012). To reach a simple factor structure, items with factor loadings that fell below .3 and items with cross-loadings on more than one factor were discarded.

### Steps to identify # of factors
1. Evaluate multicollinearity with matrix polychoric
2. Check sampling adequacy* with kmo and Bartlett’s test of sphericity 
3. Observe scree plot and eigen values
4. Conduct parallel analysis (ideally with no warning messages, but if there are, it's ok). If it shows more factors than sensible: look for communality above .4 (h2); factor loading above 0.4; 
5. Run fa() with number of factors that seem reasonable. Warnings are UNacceptable and changes in the scale need to be done. 

*Sampling adequacy evaluates whether your data have the right characteristics—enough sample size and inter-item correlations—to make factor analysis appropriate and meaningful.

```{r setup, include=FALSE}

library(rio)
library(here)
library(tidyverse)
library(haven)
library(janitor)
library(knitr)
library(surveytoolbox)
library(sjPlot)
library(kableExtra)
library(psych)
library(sjmisc)
library(MVN)

options(scipen = 999)
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# The dataset
```{r}
parent_w1 <- read_sav(here("outcomes paper/data_outcomes", "primary_parent_w1.sav"))
# 95 cases, primary parent only
```


# Homework Involvement 

conteste la frecuencia con la que ha hecho lo siguiente para ayudarle a su hija/o con la tarea en los últimos tres meses…
84	Me aseguré de que tenga un cierto tiempo para hacer la tarea
85	Me aseguré de que él / ella tenga un cierto lugar para hacer la tarea
86	Me senté con mi joven mientras él / ella hacía su tarea
87	Me hice disponible para mi joven cuando tenía preguntas sobre su tarea
88	Ayudé a mi joven cuando estaba atascado/a en una tarea
89	Ayudé a mi joven a prepararse para los exámenes haciéndole preguntas sobre el tema, etc.
90	Le di a mi joven incentivos como privilegios especiales, regalos especiales, etc.
91	Utilicé consecuencias, castigos, acciones disciplinarias
92	Le recordé a mi joven que hiciera su tarea
93	Comprobé con mi joven para asegurarme de que él / ella hiciera su tarea
94	Comprobé y corregí la tarea de mi hija/o 
95	Comprobé con la maestra/o para asegurarme que mi joven terminó su tarea 
96	Le ayudé a mi joven acceder a otros recursos (biblioteca, materiales de arte, laboratorio de computación, etc.)
97	Le ayudé a mi joven practicar habilidades (lectura, revisión, problemas de práctica, etc.)
98	Limité el ruido y/o las distracciones mientras hacia su tarea

## The scale to factor analyze
```{r}
hw_inv <- parent_w1 %>%
  select(q84, q85, q86, q87, q88, q89, q90, q91, q92, q93, q94, q95, q96, q97, q98)
```

## kmo Kaiser-Meyer-Olkin factor sampling adequacy (above 0.5, closer 1 is best)
```{r}
KMO(hw_inv) # 0.71 --> ok
```
```{r}
poly_corr <- polychoric(hw_inv)$rho
cortest.bartlett(poly_corr, n = nrow(hw_inv))
```
## Poly corr matrix
```{r}
poly <- polychoric(hw_inv)
cor.plot(poly$rho, numbers = T, upper = F, main = "Polychoric", show.legend = F)
# items 84, 89 neg corr with other items
# 84	Me aseguré de que tenga un cierto tiempo para hacer la tarea
# 89	Ayudé a mi joven a prepararse para los exámenes haciéndole preguntas sobre el tema, etc.
```

## scree plot
```{r}
scree(hw_inv, factors = TRUE, pc = FALSE, main = "Scree plot", hline = NULL, add = FALSE)
# Scree plot shows 3 factor solution, but 1 eigenvalue would be less than 1.0
```

## EFA using ULS 1 factor
```{r}
factor_test_uls <- fa(hw_inv, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 1) 
factor_test_uls
```

## EFA using ULS 2 factors
```{r}
factor_test_uls <- fa(hw_inv, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 2) 
factor_test_uls
```

## new efa iteration without q90, q91, q92, q95, q98 -> below 0.3 loadings or cross-loading
90	Le di a mi joven incentivos como privilegios especiales, regalos especiales, etc.
91	Utilicé consecuencias, castigos, acciones disciplinarias
92	Le recordé a mi joven que hiciera su tarea
95	Comprobé con la maestra/o para asegurarme que mi joven terminó su tarea 
98	Limité el ruido y/o las distracciones mientras hacia su tarea

```{r}
hw_inv_1 <- parent_w1 %>%
  select(q84, q85, q86, q87, q88, q89, q93, q94, q96, q97)
```


```{r}
KMO(hw_inv_1) # 0.73 --> ok

poly_corr <- polychoric(hw_inv_1)$rho
cortest.bartlett(poly_corr, n = nrow(hw_inv_1))
```

## Poly corr matrix
```{r}
poly <- polychoric(hw_inv_1)
cor.plot(poly$rho, numbers = T, upper = F, main = "Polychoric", show.legend = F) # looks better, 84 and 85 show some neg corr with other items.
```

```{r}
scree(hw_inv_1, factors = TRUE, pc = FALSE, main = "Scree plot", hline = NULL, add = FALSE)
# Scree plot shows 3 factor solution, but 1 eigenvalue would be less than 1.0
```

## parallel test
```{r}
fa.parallel(hw_inv_1, n.obs=NULL, fm="uls", fa="fa", main="Parallel Analysis Scree Plots", cor="poly", use="pairwise", plot=TRUE)
# Parallel analysis suggests that the number of factors =  2, multiple warnings
```

## EFA using ULS 2 factors
```{r}
factor_test_uls <- fa(hw_inv_1, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 2) 
factor_test_uls
fa.diagram(factor_test_uls)
```

## factor # 1: tutoring hwi
86	Me senté con mi joven mientras él / ella hacía su tarea
87	Me hice disponible para mi joven cuando tenía preguntas sobre su tarea
88	Ayudé a mi joven cuando estaba atascado/a en una tarea
89	Ayudé a mi joven a prepararse para los exámenes haciéndole preguntas sobre el tema, etc
94	Comprobé y corregí la tarea de mi hija/o 
96	Le ayudé a mi joven acceder a otros recursos (biblioteca, materiales de arte, laboratorio de computación, etc.)
97	Le ayudé a mi joven practicar habilidades (lectura, revisión, problemas de práctica, etc.)
```{r}
hw_inv_fa1 <- parent_w1 %>%
  select(q86, q87, q88, q89, q94, q96, q97)

# alpha
alpha(hw_inv_fa1) # 0.78
```

## factor # 2: checking hwi
84	Me aseguré de que tenga un cierto tiempo para hacer la tarea
85	Me aseguré de que él / ella tenga un cierto lugar para hacer la tarea
93	Comprobé con mi joven para asegurarme de que él / ella hiciera su tarea
```{r}
hw_inv_fa2 <- parent_w1 %>%
  select(q84, q85, q93)
# alpha
alpha(hw_inv_fa2) # 0.77
```

## Omega 
```{r}
omega(hw_inv_1, nfactors = 2, n.obs = 95, flip = T, plot = T) # alpha = 0.76
```

## structure 
Los últimos tres meses, he tomado medidas para apoyar a mi joven a...
23.	… desarrollar un plan para estudiar y hacer la tarea
24.	… encontrar formas de involucrarse en las actividades escolares
25.	… encontrar maneras de involucrarse en actividades de voluntario/a en la comunidad
26.	… hacer metas para el año escolar
27.	… cumplir fechas de limite
28.	… participar en papeles de liderazgo en la escuela o comunidad
29.	… enfocarse en hacer lo mejor posible
30.	… desarrollar un horario para completar la tarea y los proyectos escolares a tiempo

## The scale to factor analyze
```{r}
struct <- parent_w1 %>%
  select(q23, q24, q25, q26, q27, q28, q29, q30) 
```

## sampling adequacy
```{r}
KMO(struct) # 0.76

poly_corr <- polychoric(hw_inv_1)$rho
cortest.bartlett(poly_corr, n = nrow(hw_inv_1))
```

## Poly corr matrix
```{r}
poly <- polychoric(struct)
cor.plot(poly$rho, numbers = T, upper = F, main = "Polychoric", show.legend = F) # healthy corrs.
```

## scree plot
```{r}
scree(struct, factors = TRUE, pc = FALSE, main = "Scree plot", hline = NULL, add = FALSE)
# Scree plot clearly shows 1 factor solution
```

## parallel test
```{r}
fa.parallel(struct, n.obs=NULL, fm="uls", fa="fa", main="Parallel Analysis Scree Plots", cor="poly", use="pairwise", plot=TRUE)
# Parallel analysis suggests that the number of factors =  4, multiple warnings
```

## EFA using ULS - 2 factors
```{r}
factor_test_uls <- fa(struct, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 2) 
factor_test_uls
# warnings. Too many crossloadings. one heywood case.
```

## EFA using ULS - 1 factor
```{r}
factor_test_uls <- fa(struct, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 1) 
factor_test_uls
fa.diagram(factor_test_uls)
# no warnings. all loading above .30
```

## alpha
```{r}
alpha(struct) # 0.82
```

# convos with youth
en los últimos tres meses, usted ha tenido una conversación con su joven sobre…
68.	… sus amistades en la escuela. ___
69.	… su participación en actividades escolares. ____
70.	… eventos que van a pasar en la escuela. ___
71.	… sus profesores. ___
72.	… las cosas que aprende en clase. ___
73.	… cómo contribuye al salón de clase. ___
74.	… cómo va en sus clases. ___
75.	… sus futuros objetivos de carrera y educación (incluso los objetivos a corto plazo). ____
76.	… desafíos que surgen en la escuela (como intimidación, agresión, acoso, racismo, peleas, etc.) ___
77.	… sus actitudes sobre la escuela ____
78.	… su comportamiento en la escuela ____
79.	… las cosas que suceden en la escuela (el ambiente escolar en general) ____
80.	… la importancia de prepararse para futuras trayectorias educativas o de carrera, como la planificación universitaria y / o pensando acerca de la escuela técnica o comercial, etc. _____
81.	… su involucro en actividades extracurriculares en la escuela y en la comunidad como en papeles de liderazgo, haciendo trabajo voluntario en la comunidad, deportes, etc._____

## The scale to factor analyze
```{r}
convos <- parent_w1 %>%
  select(q68, q69, q70, q71, q72, q73, q74, q75, q76, q77, q78, q79, q80, q81) 
```

## kmo Kaiser-Meyer-Olkin factor
```{r}
KMO(convos) # 0.86
```

## Poly corr matrix
```{r}
poly <- polychoric(convos)
cor.plot(poly$rho, numbers = T, upper = F, main = "Polychoric", show.legend = F) # looks good, lower corr among items 80, 81, 73
```

## scree plot
```{r}
scree(convos, factors = TRUE, pc = FALSE, main = "Scree plot", hline = NULL, add = FALSE)
# Scree plot shows an almost 2 factor solution, but the second factor is a little less than 1.0 eigenvalue
```

## parallel test
```{r}
fa.parallel(convos, n.obs=NULL, fm="uls", fa="fa", main="Parallel Analysis Scree Plots", cor="poly", use="pairwise", plot=TRUE)
# Parallel analysis suggests that the number of factors =  3
# warnings.
```

## EFA using ULS 
```{r}
factor_test_uls <- fa(convos, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 3) 
factor_test_uls
# no warnings. some crossloading items
```

## removing crossloading items
72.	… las cosas que aprende en clase.
73.	… cómo contribuye al salón de clase.
74.	… cómo va en sus clases.
```{r}
convos <- parent_w1 %>%
  select(q68, q69, q70, q71, q75, q76, q77, q78, q79, q80, q81) 
```

## EFA using ULS 
```{r}
factor_test_uls <- fa(convos, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 3) 
factor_test_uls
fa.diagram(factor_test_uls)
# no warnings. all loading above .30
```

## factor 1: school climate & behavior (not included in multiple regression analyses)
76.	… desafíos que surgen en la escuela (como intimidación, agresión, acoso, racismo, peleas, etc.) ___
77.	… sus actitudes sobre la escuela ____
78.	… su comportamiento en la escuela ____
79.	… las cosas que suceden en la escuela (el ambiente escolar en general) ____

```{r}
convos_climate <- parent_w1 %>%
  select(q76, q77, q78, q79) 
# alpha
alpha(convos_climate) # 0.84
```

## factor 2: socialization/social support (not included in multiple regression analyses)
68.	… sus amistades en la escuela. ___
69.	… su participación en actividades escolares. ____
70.	… eventos que van a pasar en la escuela. ___
71.	… sus profesores. ___
```{r}
convos_social <- parent_w1 %>%
  select(q68, q69, q70, q71) 
# alpha
alpha(convos_social) # 0.79
```

## factor 3: future, prepping
75.	… sus futuros objetivos de carrera y educación (incluso los objetivos a corto plazo). ____
80.	… la importancia de prepararse para futuras trayectorias educativas o de carrera, como la planificación universitaria y / o pensando acerca de la escuela técnica o comercial, etc. _____
81.	… su involucro en actividades extracurriculares en la escuela y en la comunidad como en papeles de liderazgo, haciendo trabajo voluntario en la comunidad, deportes, etc._____
```{r}
convos_college <- parent_w1 %>%
  select(q75, q80, q81) 
# alpha
alpha(convos_college) # 0.77, removing item 81 would increase alpha, but scale would be only of 2 items. 
```

## Omega 
```{r}
omega(convos, nfactors = 3, n.obs = 95, flip = T, plot = T) # .86 looks really beautiful.
```

# school involvement & others 
En general, hago un esfuerzo para...
31.	… conocer el personal y la administración de la escuela  ___
32.	… conocer al menos uno de los maestros de mi joven ____
33.	… entender las reglas y pólizas de la escuela  ____
34.	… informarme sobre mis derechos como padre ____
35.	… aprender sobre el sistema educativo en este estado ____
36.	… entender la diferencia entre obtener un GED, graduarse con un diploma estándar de la escuela secundaria, o con un diploma de una secundaria internacional o con un diploma de Bachillerato Internacional.____
37.	… involucrarse en las actividades escolares, en el salón de clase, y/u otras maneras (por ejemplo, organizaciones de padres, trabajo voluntario, etc.)
38.	… tener conversaciones con los otros padres para obtener información o aprender acerca de los recursos en la escuela. ____
39.	…contactar los otros padres para obtener apoyo. ____
40.	…entender la trayectoria hacia la preparación a la universidad y para una carrera ____
41.	… asistir a la conferencia de padres y maestros cuando esté disponible. _____

## The scale to factor analyze
```{r}
scho_inv <- parent_w1 %>%
  select(q31, q32, q33, q34, q35, q36, q37, q38, q39, q40, q41)
```

## sampling adequacy
```{r}
KMO(scho_inv) # 0.83

poly_corr <- polychoric(scho_inv)$rho
cortest.bartlett(poly_corr, n = nrow(scho_inv))
```

## Poly corr matrix
```{r}
poly <- polychoric(scho_inv)
cor.plot(poly$rho, numbers = T, upper = F, main = "Polychoric", show.legend = F) # item 41 with 37-39 neg corrs.
```

## scree plot
```{r}
scree(scho_inv, factors = TRUE, pc = FALSE, main = "Scree plot", hline = NULL, add = FALSE)
# Scree plot  shows 2 factor solution
```

## The scale to factor analyze
```{r}
scho_inv <- parent_w1 %>%
  select(q31, q32, q33, q34, q35, q36, q37, q38, q39, q40) # removed 41
```

## sampling adequacy
```{r}
KMO(scho_inv) # 0.84

poly_corr <- polychoric(scho_inv)$rho
cortest.bartlett(poly_corr, n = nrow(scho_inv))
```

## Poly corr matrix
```{r}
poly <- polychoric(scho_inv)
cor.plot(poly$rho, numbers = T, upper = F, main = "Polychoric", show.legend = F) 
```

## parallel test
```{r}
fa.parallel(scho_inv, n.obs=NULL, fm="uls", fa="fa", main="Parallel Analysis Scree Plots", cor="poly", use="pairwise", plot=TRUE)
# Parallel analysis suggests that the number of factors =  3
```

## EFA using ULS 
```{r}
factor_test_uls <- fa(scho_inv, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 2) 
factor_test_uls
fa.diagram(factor_test_uls)
# warnings
# crossloading items:
# 35 = 0.58	0.34 (flagged for removal) diff. betw. primary and secondary loadings is .24
# 36 = 0.23	0.58 (not too problematic) diff. betw. primary and secondary loadings is .35
# 40 = 0.22	0.66 (not too problematic) diff. betw. primary and secondary loadings is .44
```
-------------------------------------- removing 35 --------------------------------

## Removing crossloading item 35
35.	… aprender sobre el sistema educativo en este estado (to be removed)
```{r}
scho_inv <- parent_w1 %>%
  select(q31, q32, q33, q34, q36, q37, q38, q39, q40) # removed 35
```

## new EFA iteration 
```{r}
factor_test_uls <- fa(scho_inv, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 2) 
factor_test_uls
fa.diagram(factor_test_uls)
# no warnings
```

## Factor 1: knowing school personnel and rules
31.	… conocer el personal y la administración de la escuela  ___
32.	… conocer al menos uno de los maestros de mi joven ____
33.	… entender las reglas y pólizas de la escuela  ____
34.	… informarme sobre mis derechos como padre ____
```{r}
scho_inv_fa1 <- parent_w1 %>%
  select(q31, q32, q33, q34)

# alpha
alpha(scho_inv_fa1) # 0.85 (familiarizing with school)
```

## Factor 2: networking and...
36.	… entender la diferencia entre obtener un GED, graduarse con un diploma estándar de la escuela secundaria, o con un diploma de una secundaria internacional o con un diploma de Bachillerato Internacional.____
37.	… involucrarse en las actividades escolares, en el salón de clase, y/u otras maneras (por ejemplo, organizaciones de padres, trabajo voluntario, etc.)
38.	… tener conversaciones con los otros padres para obtener información o aprender acerca de los recursos en la escuela. ____
39.	…contactar los otros padres para obtener apoyo. ____
40.	…entender la trayectoria hacia la preparación a la universidad y para una carrera ____
```{r}
scho_inv_fa2 <- parent_w1 %>%
  select(q36, q37, q38, q39, q40)

# alpha
alpha(scho_inv_fa2) # .81 (strategic sbi)
```

-------------------------------------- removing 35, 36, 40 --------------------------------

## Removing crossloading item 35
35.	… aprender sobre el sistema educativo en este estado (to be removed)
36.	… entender la diferencia entre obtener un GED, graduarse con un diploma estándar de la escuela secundaria, o con un diploma de una secundaria internacional o con un diploma de Bachillerato Internacional.
40.	…entender la trayectoria hacia la preparación a la universidad y para una carrera
```{r}
# scho_inv <- parent_w1 %>%
#   select(q31, q32, q33, q34, q37, q38, q39)
```

## new EFA iteration 
```{r}
# factor_test_uls <- fa(scho_inv, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 2) 
# factor_test_uls
# fa.diagram(factor_test_uls)
# # no warnings
```

## Factor 1: knowing school personnel and rules
31.	… conocer el personal y la administración de la escuela  ___
32.	… conocer al menos uno de los maestros de mi joven ____
33.	… entender las reglas y pólizas de la escuela  ____
34.	… informarme sobre mis derechos como padre ____
```{r}
# scho_inv_fa1 <- parent_w1 %>%
#   select(q31, q32, q33, q34)
# 
# # alpha
# alpha(scho_inv_fa1) # 0.85 (familiarizing with school)
```

## Factor 2: volunteering and networking
37.	… involucrarse en las actividades escolares, en el salón de clase, y/u otras maneras (por ejemplo, organizaciones de padres, trabajo voluntario, etc.)
38.	… tener conversaciones con los otros padres para obtener información o aprender acerca de los recursos en la escuela. ____
39.	…contactar los otros padres para obtener apoyo. ____

```{r}
# scho_inv_fa2 <- parent_w1 %>%
#   select(q37, q38, q39)
# 
# # alpha
# alpha(scho_inv_fa2) # .79
```

## Omega 
```{r}
# scho_inv <- parent_w1 %>%
#   select(q31, q32, q33, q34, q37, q38, q39)
# omega(scho_inv, nfactors = 2, n.obs = 95, flip = T, plot = T) # alpha 0.82
# # warning : 3 factors are required fo identification
```

# comfort 
Como madre/padre de esta escuela, siento que soy…
47.	…parte de una comunidad con el personal de la escuela y los otros padres. ____
48.	…tratado/a con respeto, sabiendo que mis opiniones son importantes. ____
49.	…cómoda/o preguntando por un traductor y por materiales en español si los necesito. ____
50.	… capaz de hablar con maestros o administradores sobre grandes preocupaciones relacionados con mi joven. ___
51.	…dedicada/o en creando un ambiente exitoso para todos los jóvenes. ____
52.	 …feliz de que mi joven asista a esta escuela. ___
53.	…bienvenido/a en la escuela de mi joven. ____

taps on parent comfort at school, except maybe item 51 about parent contributing

## The scale to factor analyze
```{r}
comfort <- parent_w1 %>%
  select(q47, q48, q49, q50, q51, q52, q53)
```

## sampling adequacy
```{r}
KMO(comfort) # 0.83

poly_corr <- polychoric(comfort)$rho
cortest.bartlett(poly_corr, n = nrow(comfort))
```
## Poly corr matrix
```{r}
poly <- polychoric(comfort)
cor.plot(poly$rho, numbers = T, upper = F, main = "Polychoric", show.legend = F) # all healthy corrs
```

## scree plot
```{r}
scree(comfort, factors = TRUE, pc = FALSE, main = "Scree plot", hline = NULL, add = FALSE)
# Scree plot clearly shows 1 factor solution
```

## parallel test
```{r}
fa.parallel(comfort, n.obs=NULL, fm="uls", fa="fa", main="Parallel Analysis Scree Plots", cor="poly", use="pairwise", plot=TRUE)
# Parallel analysis suggests that the number of factors =  3
# warnings
```
## EFA using ULS - 1 factor
```{r}
factor_test_uls <- fa(comfort, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 1) 
factor_test_uls
#going for 2 factors, because at least one factor would have 2 items
# no warnings
```

## EFA using ULS 
```{r}
factor_test_uls <- fa(comfort, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 2) 
factor_test_uls
#going for 2 factors, because at least one factor would have 2 items
# no warnings
```

## Factor 1: belong
48.	…tratado/a con respeto, sabiendo que mis opiniones son importantes. ____
49.	…cómoda/o preguntando por un traductor y por materiales en español si los necesito. ____
52.	 …feliz de que mi joven asista a esta escuela. ___
53.	…bienvenido/a en la escuela de mi joven. ____
```{r}
comfort_fa1 <- parent_w1 %>%
  select(q48, q49, q52, q53)
# alpha
alpha(comfort_fa1) # 0.86
```

## Factor 2: included (not included in multiple regression analyses)
47.	…parte de una comunidad con el personal de la escuela y los otros padres. ____
50.	… capaz de hablar con maestros o administradores sobre grandes preocupaciones relacionados con mi joven. ___
51.	…dedicada/o en creando un ambiente exitoso para todos los jóvenes. ____

```{r}
comfort_fa2 <- parent_w1 %>%
  select(q47, q50, q51)
# alpha
alpha(comfort_fa2) # 0.70
```

## Omega 
### says 3 factors required for identification, just to test it, it would be:

49.	…cómoda/o preguntando por un traductor y por materiales en español si los necesito. ____
52.	…feliz de que mi joven asista a esta escuela. ___
53.	…bienvenido/a en la escuela de mi joven. ____

47.	…parte de una comunidad con el personal de la escuela y los otros padres. ____
50.	… capaz de hablar con maestros o administradores sobre grandes preocupaciones relacionados con mi joven. ___
51.	…dedicada/o en creando un ambiente exitoso para todos los jóvenes. ____

Crossloading 
47
48.	…tratado/a con respeto, sabiendo que mis opiniones son importantes. ____
53

```{r}
omega(comfort, nfactors = 3, n.obs = 95, flip = T, plot = T) # 0.84
```

# endorse & rels 
Como madre/padre en esta escuela, estoy segura/o de que…
54.	… esta escuela es un buen lugar para mi joven. _____
55.	… el personal de la escuela de mi joven está haciendo cosas buenas por ella / el. _____
56.	… la gente en la escuela de mi joven es confiable. _____
57.	… la escuela de mi joven hace un buen trabajo preparando a los jóvenes para sus futuros. _____
58.	… puedo encontrar ayuda para mi joven si él / ella está batallando en una clase. ___
59.	… los maestros de mi joven se preocupan por ella/el. ____
60.	… maestros y administradores trabajan juntos para crear un ambiente seguro y acogedor para todos. ____
61.	… hay muchas oportunidades para la involucración de los padres. ____
62.	… puedo tener una conversación honesta y respetuosa sobre mi joven con su maestro. ___
63.	…puedo trabajar con un maestro para resolver cualquier problema que tenga mi joven en la escuela. _____


## The scale to factor analyze
```{r}
rels <- parent_w1 %>%
  select(q54, q55, q56, q57, q58, q59, q60, q61, q62, q63)
```

## kmo Kaiser-Meyer-Olkin factor
```{r}
KMO(rels) # 0.88
```

## Poly corr matrix
```{r}
poly <- polychoric(rels)
cor.plot(poly$rho, numbers = T, upper = F, main = "Polychoric", show.legend = F) # all healthy corrs
```

## scree plot
```{r}
scree(rels, factors = TRUE, pc = FALSE, main = "Scree plot", hline = NULL, add = FALSE)
# Scree plot clearly shows 1 factor solution
```

## parallel test
```{r}
fa.parallel(rels, n.obs=NULL, fm="uls", fa="fa", main="Parallel Analysis Scree Plots", cor="poly", use="pairwise", plot=TRUE)
# Parallel analysis suggests that the number of factors =  2
# warnings
```

## EFA using ULS with 1 factor - iteration 1
```{r}
factor_test_uls <- fa(rels, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 1) 
factor_test_uls
# matrix nonpositive definite, items 58, 60 and 63 
```

## removing items 58, 60 and 63 
```{r}
rels <- parent_w1 %>%
  select(q54, q55, q56, q57, q59, q61, q62)
```

## EFA using ULS - iteration 2 (with 1 and 2 factors)
```{r}
factor_test_uls <- fa(rels, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 2) 
factor_test_uls
# matrix nonpositive definite, items loading >1 and crossloading items.

factor_test_uls <- fa(rels, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 1) 
factor_test_uls
# matrix nonpositive definite
```
######################### re-testing again on 06/11/25, removing problematic items ############################### START ###############################

## removing item 55
```{r}
rels <- parent_w1 %>%
  select(q54, q56, q57, q59, q61, q62)
```

## EFA using ULS - iteration 3 (with 1 and 2 factors)
```{r}
factor_test_uls <- fa(rels, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 2) 
factor_test_uls
# ultraheywood case, items 61 crossloading

factor_test_uls <- fa(rels, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 1) 
factor_test_uls
# error message indicating to use another estimation method
```

## removing item 59
```{r}
rels <- parent_w1 %>%
  select(q54, q56, q57, q61, q62)
```

## EFA using ULS - iteration 4 (with 1 factor)
```{r}
factor_test_uls <- fa(rels, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 1) 
factor_test_uls
# error message indicating to use another estimation method
```
54.	… esta escuela es un buen lugar para mi joven. _____
56.	… la gente en la escuela de mi joven es confiable. _____
57.	… la escuela de mi joven hace un buen trabajo preparando a los jóvenes para sus futuros. _____
61.	… hay muchas oportunidades para la involucración de los padres. ____
62.	… puedo tener una conversación honesta y respetuosa sobre mi joven con su maestro. ___

note: 61 and 62 don't hang well with the other items

## removing items 61 and 62 --- this is PTIQ minus items 55
```{r}
rels <- parent_w1 %>%
  select(q54, q56, q57)
```

## EFA using ULS - iteration 5 (with 1 factor)
```{r}
factor_test_uls <- fa(rels, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 1) 
factor_test_uls
# no error message, although loading for 57 is 1.00  
```
## kmo Kaiser-Meyer-Olkin factor
```{r}
KMO(rels) # 0.65 # okay but close (0.5)
```

## Poly corr matrix
```{r}
poly <- polychoric(rels)
cor.plot(poly$rho, numbers = T, upper = F, main = "Polychoric", show.legend = F) # very high corrs between 56 and 57
```
The EFA did run without error messages for the scale without 55, that is the scale I originally ran: endorse_1

######################### re-testing again on 06/11/25, removing problematic items ############################### END ###############################

## trying only PTIQ endorsement scale items
23. Your child's school is a good place for your child to be
24. The staff at your child's school is doing good things for your child
25. You have confidence in the people at your child's school
26. Your child's school is doing a good job of preparing children for their futures.


54.	… esta escuela es un buen lugar para mi joven. _____
55.	… el personal de la escuela de mi joven está haciendo cosas buenas por ella / el. _____
56.	… la gente en la escuela de mi joven es confiable. _____
57.	… la escuela de mi joven hace un buen trabajo preparando a los jóvenes para sus futuros. _____

```{r}
endorse <- parent_w1 %>%
  select(q54, q55, q56, q57)
alpha(endorse) # 0.87
```

## sampling adequacy
```{r}
KMO(endorse) # 0.81

poly_corr <- polychoric(endorse)$rho
cortest.bartlett(poly_corr, n = nrow(endorse))

poly <- polychoric(endorse)
cor.plot(poly$rho, numbers = T, upper = F, main = "Polychoric", show.legend = F) # corrs above .90 potentially multicolleniearity
```

```{r}
factor_test_uls <- fa(endorse, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 1) 
factor_test_uls
fa.diagram(factor_test_uls)
# ultraheywood case
```

## removing 55 bec too similar to 57

54.	… esta escuela es un buen lugar para mi joven. _____
56.	… la gente en la escuela de mi joven es confiable. _____
57.	… la escuela de mi joven hace un buen trabajo preparando a los jóvenes para sus futuros. _____
```{r}
endorse <- parent_w1 %>%
  select(q54, q56, q57)
alpha(endorse) # 0.79
```
## sampling adequacy
```{r}
KMO(endorse) # 0.65

poly_corr <- polychoric(endorse)$rho
cortest.bartlett(poly_corr, n = nrow(endorse))

poly <- polychoric(endorse)
cor.plot(poly$rho, numbers = T, upper = F, main = "Polychoric", show.legend = F)
```


iteration 2: no negative error variance
```{r}
factor_test_uls <- fa(endorse, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 1) 
factor_test_uls
fa.diagram(factor_test_uls)
# no warnings, but there is a factor loading of 1 (item 57)
```
conclusion for paper: going with this one. 



# teacher support 
Como madre/padre en esta escuela, estoy segura/o de que…
58.	… puedo encontrar ayuda para mi joven si él / ella está batallando en una clase. ___
59.	… los maestros de mi joven se preocupan por ella/el. ____
60.	… maestros y administradores trabajan juntos para crear un ambiente seguro y acogedor para todos. ____
61.	… hay muchas oportunidades para la involucración de los padres. ____
62.	… puedo tener una conversación honesta y respetuosa sobre mi joven con su maestro. ___
63.	…puedo trabajar con un maestro para resolver cualquier problema que tenga mi joven en la escuela. _____

## The scale to factor analyze
```{r}
support <- parent_w1 %>%
  select(q58, q59, q60, q61, q62, q63)
```

## kmo Kaiser-Meyer-Olkin
```{r}
KMO(support) # 0.81
```

## Poly corr matrix
```{r}
poly <- polychoric(support)
cor.plot(poly$rho, numbers = T, upper = F, main = "Polychoric", show.legend = F) # all healthy corrs
```

## scree plot
```{r}
scree(support, factors = TRUE, pc = FALSE, main = "Scree plot", hline = NULL, add = FALSE)
# Scree plot shows 1 factor solution
```

## parallel test
```{r}
fa.parallel(support, n.obs=NULL, fm="uls", fa="fa", main="Parallel Analysis Scree Plots", cor="poly", use="pairwise", plot=TRUE)
# Parallel analysis suggests that the number of factors =  2
# warnings
```

## EFA using ULS - 1 and 2 factors
```{r}
factor_test_uls <- fa(support, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 2) 
factor_test_uls
# matrix nonpositive definite

factor_test_uls <- fa(support, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 1) 
factor_test_uls
# matrix nonpositive definite
```

## removing 60 and 61 because doesn't appear to jive well with other items

60.	… maestros y administradores trabajan juntos para crear un ambiente seguro y acogedor para todos. ____
61.	… hay muchas oportunidades para la involucración de los padres. ____
```{r}
support <- parent_w1 %>%
  select(q58, q59, q62, q63)
```

## EFA using ULS 
```{r}
# EFA using ULS 
factor_test_uls <- fa(support, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 1) 
factor_test_uls
fa.diagram(factor_test_uls)
# no warnings!
```


```{r}
alpha(support) # 0.88
```


# p-t rel from PTIQ

Not included in battery:
11. You feel welcome to visit about your child's school
12. you enjoy talking with your child's teacher 
16. you feel your child's teacher pays attention to your suggestions

Included in battery
13. you feel your child teacher cares about your child
14. you think your child's teacher is interested in getting to know you
15. you feel comfortable talking with your child's teacher about your child
17. you ask your child's teacher questions or make suggestions about your child

En esta escuela, siento que hay por lo menos un maestro…
64.	… que se preocupa por mi joven. _____
65.	… quien está interesado en conocerme _____
66.	… con quien me siento/a cómodo/a hablando sobre mi joven.  _____
67.	… de quien puedo hacer preguntas o hacer sugerencias sobre mi joven. _____

## The scale to factor analyze
```{r}
pt_rel <- parent_w1 %>%
  select(q64, q65, q66, q67)
```

## sampling adequacy
```{r}
KMO(pt_rel) # 0.77

poly_corr <- polychoric(pt_rel)$rho
cortest.bartlett(poly_corr, n = nrow(pt_rel)) # Matrix was not positive definitive.

poly <- polychoric(pt_rel)
cor.plot(poly$rho, numbers = T, upper = F, main = "Polychoric", show.legend = F) # potentially multicolleniearity
```


## scree plot
```{r}
scree(pt_rel, factors = TRUE, pc = FALSE, main = "Scree plot", hline = NULL, add = FALSE)
# Scree plot clearly shows 1 factor solution
```

## parallel test
```{r}
fa.parallel(pt_rel, n.obs=NULL, fm="uls", fa="fa", main="Parallel Analysis Scree Plots", cor="poly", use="pairwise", plot=TRUE)
# Parallel analysis suggests that the number of factors =  1
# warnings: matrix not positive definitive
```

## EFA using ULS 
```{r}
factor_test_uls <- fa(pt_rel, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 1) 
factor_test_uls
# Matrix was not positive definite, ultraheywood case
```

## removing item 66 loading > 1 and negative error variance

66.	… con quien me siento/a cómodo/a hablando sobre mi joven.  
```{r}
pt_rel <- parent_w1 %>%
  select(q64, q65, q67)
```

## without item 66
```{r}
factor_test_uls <- fa(pt_rel, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 1) 
factor_test_uls
fa.diagram(factor_test_uls)
# Warnings: ultra heywood case, item 67 loading = 1
```

Conclusion: EFA not possible with this scale. However, the scale was included in multiple regression due to importance to assess parent-teacher relationship. Alpha is good. 

```{r}
pt_rel <- parent_w1 %>%
  select(q64, q65, q66, q67)

alpha(pt_rel) # 0.87

pt_rel_2 <- parent_w1 %>%
  select(q64, q65, q67)

alpha(pt_rel_2) # 0.79
```

# self-efficacy (this is an outcome in multiple regression analyses)
En general, estoy segura/o de que…
42.	…yo entiendo cómo funciona la escuela de mi joven. ____
43.	…puedo comunicar mis preguntas y preocupaciones con los maestros y el personal de la escuela.___
44.	…puedo trabajar con la escuela para encontrar una solución positiva si surge un conflicto o un problema que involucre a mi joven en la escuela. ____
45.	…estoy ayudando a mi joven hacer bien en la escuela. ____
46.	… estoy ayudando a mi joven prepararse para sus futuras metas de educación y carrera. ____

## The scale to factor analyze
```{r}
s_efficacy <- parent_w1 %>%
  select(q42, q43, q44, q45, q46)
```

## kmo Kaiser-Meyer-Olkin
```{r}
KMO(s_efficacy) # 0.71
```

## Poly corr matrix
```{r}
poly <- polychoric(s_efficacy)
cor.plot(poly$rho, numbers = T, upper = F, main = "Polychoric", show.legend = F) # all healthy corrs
```

## scree plot
```{r}
scree(s_efficacy, factors = TRUE, pc = FALSE, main = "Scree plot", hline = NULL, add = FALSE)
# Scree plot clearly shows 1 factor solution
```


## parallel test, anyways
```{r}
fa.parallel(s_efficacy, n.obs=NULL, fm="uls", fa="fa", main="Parallel Analysis Scree Plots", cor="poly", use="pairwise", plot=TRUE)
# Parallel analysis suggests that the number of factors =  1
# warnings
```

## EFA using ULS 
```{r}
factor_test_uls <- fa(s_efficacy, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 1) 
factor_test_uls
fa.diagram(factor_test_uls)
# no warnings
```

## alpha
```{r}
alpha(s_efficacy) # 0.8 (parent self-efficacy working with school & supporting child)
```

