---
title: "models for FSC paper"
author: "Alejandra Garcia Isaza"
date: "2023-04-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rio)
library(here)
library(tidyverse)
library(haven)
library(janitor)
library(knitr)
library(surveytoolbox)
library(sjPlot)
library(kableExtra)
library(psych)
library(sjmisc)
library(MVN)

theme_set(theme_minimal())
```


# Loading paent data
```{r}
p <- read_sav(here("outcomes paper/data_outcomes", "primary_parent_w1.sav")) 

p_1 <- p %>%
  select(-response_id, -participant_4_text, -location, -location_5_text)

names(p_1) <- paste(names(p_1), "p", sep = "_")

p_2 <- p_1 %>%
  rename(c("school_id" = "school_id_p"),
         c("condition" = "condition_p"),
         c("family_id" = "family_id_p"))
```

#Loading youth data
```{r}
y <- read_sav(here("outcomes paper/data_outcomes", "all_youth_wave1.sav"))

view_df(y)
```

# Data prep: recoding missing variables as N/A
```{r include=FALSE}
# recoding missing values as N/A with function

# vector with missing values in dataset
missing_vals <- c(77, 88, 99, -99)

# function that returns true if values in vector are equal to missing_vals. The function takes a vector x, and specified values of missing data
recode_missing <- function(x, missing_vals = c(77, 88, 99, -99)) {
  test <- x %in% missing_vals
  ifelse(test, NA, x)
}

# function that recodes missing values to NA. The function takes a dataframe with variables with missing data, and specified values of missing data
recode_missing_df <- function(df, missing_vals = c(77, 88, 99, -99)) {
  modify(df, ~recode_missing(.x, missing_vals)) # here uses the function created above
}

y_1 <- recode_missing_df(y) # the function strips out variable labels
```

# Data prep: reverse scoring
```{r}
y_2 <- y_1 %>%
  mutate(q84 = likert_reverse(q84, top = 5, bottom = 1),
         q86 = likert_reverse(q86, top = 5, bottom = 1),
         q87 = likert_reverse(q87, top = 5, bottom = 1))
```


```{r}
names(y_2) <- paste(names(y_2), "y", sep = "_")
```

```{r}
y_3 <- y_2 %>%
  select(-response_id_y, -int_location_y, -int_location_4_text_y) %>%
  rename(c("school_id" = "school_id_y"),
         c("condition" = "condition_y"),
         c("family_id" = "id_y"))
```


# Joining youth and parent wave 1
```{r}
par_you_w1 <- right_join(p_2, y_3) # 95 rows of data ... not full join bec. youth id 623 has no parent data

```

Youth outcomes:

letter grades = q4
overall GPA = q5 (this is mostly NA)
hw_practices = q6, q7, q8, q9, q10, q11, q12, q13
likelihood of dropping out = q16
y_sch_eng = q83, q84, q85, q86, q87, q88, q89, q90, q91 (alpha = 0.72)

```{r}
par_you_w2 <- par_you_w1 %>%
  rowwise() %>% 
  mutate(par_agent = mean(c(q42_p, q43_p, q44_p, q47_p, q48_p, q50_p, q51_p, q54_p, q60_p), na.rm = TRUE),
         par_teach_rel = mean(c(q63_p, q64_p, q65_p, q66_p, q67_p), na.rm = TRUE),
         endorse = mean(c(q53_p, q55_p, q56_p, q57_p, q58_p, q59_p, q62_p), na.rm = TRUE),
         hw_practic = mean(c(q6_y, q7_y, q8_y, q9_y, q10_y, q11_y, q12_y, q13_y), na.rm = TRUE),
         y_sch_eng = mean(c(q83_y, q84_y, q85_y, q86_y, q87_y, q88_y, q89_y, q90_y, q91_y), na.rm = TRUE)) %>%
  select(school_id, condition, family_id, par_agent, par_teach_rel, endorse, sex_y, q2_y, q4_y, q16_y, hw_practic, y_sch_eng)
```

# making factors and releveling
```{r}

d <- par_you_w2 %>%
  mutate(youth_gender = factor(sex_y),
         youth_gender = relevel(youth_gender, ref = "1"), # male
         youth_grlevel = factor(q2_y),
         youth_grlevel = relevel(youth_grlevel, ref = "2"), # 7th grade
         youth_letter_grd = factor(q4_y),
         youth_letter_grd = relevel(youth_letter_grd, ref = "1"), # mostly A
         youth_drop = factor(q16_y),
         youth_drop = relevel(youth_drop, ref = "1")) %>% # will not drop out
 select(-sex_y, -q2_y)
```

# descriptive stats function
```{r include=FALSE}

# function that takes a df and returns a df with only the numeric columns
only_numeric <- function(df) {
  select_numeric <- dplyr::select_if(df, is.numeric)
  return(select_numeric)
}

# a list with the functions we want
smry <- list(n = function(x) length(x),
             n_valid = function(x) sum(!is.na(x)),
             na = function(x) sum(is.na(x)),
             M = function(x) round(mean(x, na.rm = TRUE), 2), 
             SD = function(x) round(sd(x, na.rm = TRUE), 2),
             Min = function(x) round(min(x, na.rm = TRUE), 2),
             Max = function(x) round(max(x, na.rm = TRUE), 2))

# wrapping solution in a function
descriptives <- function(df) {
  select_numeric <- only_numeric(df)
  mean_sd <- map_df(select_numeric, function(col) map_df(smry, ~.x(col)),
       .id = "column")
  return(mean_sd)
}
```


```{r}
only_scales <- d %>%
  select(-school_id, -condition, -family_id)

descrip_table <- descriptives(only_scales) %>%
  rename(c("Scale" = "column"))
```


# correlation matrix 
```{r}
library(Hmisc)
rcorr(as.matrix(only_scales), type = "pearson")
```

```{r}
ggplot(d, aes(y_sch_eng)) + 
  geom_histogram() # approxiamtes normality, shows one outlier z-score greater than 3.0 or -3.0 (converting to Z-scores)
```

# parent agency predicting youth engagement
```{r}
m1 <- lm(y_sch_eng ~ par_agent, d)
summary(m1) # p = 0.32

m1_c <- lm(y_sch_eng ~ youth_gender + youth_grlevel + par_agent, d) # controling for youth's gender & grade
summary(m1_c)
```
# parent-teach rel predicting youth engagement
```{r}
m2 <- lm(y_sch_eng ~ par_teach_rel, d)
summary(m2) # p = 0.17

m2_c <- lm(y_sch_eng ~ youth_gender + youth_grlevel + par_teach_rel, d) # controling for youth's gender & grade
summary(m2_c)
```
# parent endorse of sch predicting youth engagement
```{r}
m3 <- lm(y_sch_eng ~ endorse, d)
summary(m2) # p = 0.17

m3_c <- lm(y_sch_eng ~ youth_gender + youth_grlevel + endorse, d) # controling for youth's gender & grade
summary(m3_c)
```

```{r}
ggplot(d, aes(youth_letter_grd)) + 
  geom_bar()

ggplot(d, aes(q4_y)) + 
  geom_histogram()
```

#### needs to be logistic regression to use var youth_letter_grd###

# parent agency predicting youth grades
```{r}
# m1 <- glm(youth_letter_grd ~ par_agent, d)
# summary(m1) 
# 
# m1_c <- glm(youth_letter_grd ~ youth_gender + youth_grlevel + par_agent, d) # controling for youth's gender & grade
# summary(m1_c)

m1 <- lm(q4_y ~ par_agent, d)
summary(m1) # p = 0.63

m1_c <- lm(q4_y ~ youth_gender + youth_grlevel + par_agent, d) # controling for youth's gender & grade
summary(m1_c)
```

# parent-teach rel predicting youth grades
```{r}
m2 <- lm(q4_y ~ par_teach_rel, d)
summary(m2) # p = 0.98

m2_c <- lm(q4_y ~ youth_gender + youth_grlevel + par_teach_rel, d) # controling for youth's gender & grade
summary(m2_c)
```

# parent endorse of sch predicting youth engagement
```{r}
m3 <- lm(q4_y ~ endorse, d)
summary(m2) # p = 0.98

m3_c <- lm(q4_y ~ youth_gender + youth_grlevel + endorse, d) # controling for youth's gender & grade
summary(m3_c)
```

```{r}
ggplot(d, aes(youth_drop)) + 
  geom_bar()

ggplot(d, aes(q16_y)) + 
  geom_histogram()
```

# parent agency predicting youth dropping out
```{r}
# m1 <- glm(youth_drop ~ par_agent, d)
# summary(m1) 
# 
# m1_c <- glm(youth_drop ~ youth_gender + youth_grlevel + par_agent, d) # controling for youth's gender & grade
# summary(m1_c)

m1 <- lm(q16_y ~ par_agent, d)
summary(m1) # p = 0.80

m1_c <- lm(q16_y ~ youth_gender + youth_grlevel + par_agent, d) # controling for youth's gender & grade
summary(m1_c)
```

# parent-teach rel predicting youth dropping out

significant but confusing: for every one-unit increase in parent-teacher rel, the... (?) Difficult to interpret given the categorical nature of var. 
```{r}
m2 <- lm(q16_y ~ par_teach_rel, d)
summary(m2) # p = 0.04 !!

m2_c <- lm(q16_y ~ youth_gender + youth_grlevel + par_teach_rel, d) # controling for youth's gender & grade
summary(m2_c)
```

# parent endorse of sch predicting youth dropping out

significant but confusing: for every one-unit increase in parent endorse, the... (?) Difficult to interpret given the categorical nature of var. 
```{r}
m3 <- lm(q16_y ~ endorse, d)
summary(m2) # p = 0.04

m3_c <- lm(q16_y ~ youth_gender + youth_grlevel + endorse, d) # controling for youth's gender & grade
summary(m3_c)
```


```{r}
ggplot(d, aes(hw_practic)) + 
  geom_histogram()
```

# parent agency predicting youth positive hw practices
```{r}
m1 <- lm(hw_practic ~ par_agent, d)
summary(m1) # p = 0.01

m1_c <- lm(hw_practic ~ youth_gender + youth_grlevel + par_agent, d) # controling for youth's gender & grade
summary(m1_c)
```

# parent-teach rel predicting youth positive hw practices
```{r}
m2 <- lm(hw_practic ~ par_teach_rel, d)
summary(m2) # p = 0.43

m2_c <- lm(hw_practic ~ youth_gender + youth_grlevel + par_teach_rel, d) # controling for youth's gender & grade
summary(m2_c)
```

# parent endorse of sch predicting youth positive hw practices
```{r}
m3 <- lm(hw_practic ~ endorse, d)
summary(m2) # p = 0.43

m3_c <- lm(hw_practic ~ youth_gender + youth_grlevel + endorse, d) # controling for youth's gender & grade
summary(m3_c)
```












