---
title: "relations paper"
output: html_document
date: "2025-01-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rio)
library(here)
library(tidyverse)
library(haven)
library(janitor)
library(knitr)
library(surveytoolbox)
library(sjPlot)
library(kableExtra)
library(psych)
library(sjmisc)
library(MVN)
library(stats)
library(AICcmodavg)

theme_set(theme_minimal())
```

### The dataset ###

```{r}
d <- read_sav(here("outcomes paper/data_outcomes", "p_y_w1_for_analysesjan25.sav")) # 94 rows

view_df(d)

data.frame(table(d$family_id)) # no dupped ids
```

CM suggested to standardize the betas, however: 

* all structural and relational predictor variables are on a scale of 1-4

* outcomes are on a:
- 4-point scale (selfeffic_1, y_parsupp, y_parinv, y_hw)
- 5-point scale (y_sch_eng)
- 8-point scale (gen_grades)
- 8-point scale (grades)

* covariates are  
- parent_eng_comf = 5-point scale 
- youth_gender = categorical 2 levels
- youth_age = range 10-13

```{r}
d1 <- d %>%
  mutate(parent_eng_comf = as.numeric(parent_eng_comf), # average comfort
         youth_gender = factor(youth_gender),
         youth_gender = relevel(youth_gender, ref = "1"), # male
         youth_age = as.numeric(youth_age)) # average youth age

view_df(d1)
```


```{r}
only_scales <- d1 %>%
  select(acadhwi_1, checkhwi_1, checkhwi_1, struct_1, convfut_1, knowschool_1, selfeffic_1, parnets_1, welcome_1, endorse_1, supported_1, ptrel_1)

# descrip_table <- descriptives(only_scales) %>%
#   rename(c("Scale" = "column"))
```


```{r}
library(Hmisc)
rcorr(as.matrix(only_scales), type = "pearson")
```

The expected school eng value is 2.46, when we consider average values of all predictors in the model and student is male.

For every one-point increase in checking hw, school eng increases, on average, 0.23 points

```{r}
# school engagement: youth report
m <- lm(y_sch_eng ~ acadhwi_1 + checkhwi_1 + struct_1 + convfut_1 + knowschool_1 + parnets_1 + welcome_1 + endorse_1 + supported_1 + ptrel_1, d1)
summary(m) # only checking hw came out significant 0.01

m_cov <- lm(y_sch_eng ~ parent_eng_comf + youth_age + youth_gender + acadhwi_1 + checkhwi_1 + struct_1 + convfut_1 + knowschool_1 + parnets_1 + welcome_1 + endorse_1 + supported_1 + ptrel_1, d1)
summary(m_cov) # only checking hw came out significant 0.04

```

The expected grades value is 4.50, when we consider average values of all predictors in the model and student is male, but is not significant.

no predictors in the model are significantly related to youth grades.

```{r}
# grades: youth report
m2 <- lm(grades ~ acadhwi_1 + checkhwi_1 + struct_1 + convfut_1 + knowschool_1 + parnets_1 + welcome_1 + endorse_1 + supported_1 + ptrel_1, d1)
summary(m2) # only conversation about future came out significant 0.02

m2_cov <- lm(grades ~ parent_eng_comf + youth_age + youth_gender + acadhwi_1 + checkhwi_1 + struct_1 + convfut_1 + knowschool_1 + parnets_1 + welcome_1 + endorse_1 + supported_1 + ptrel_1, d1)
summary(m2_cov) # no significant relations
```

The expected youth hw value is 1.71, when we consider average values of all predictors in the model and student is male.

no predictors in the model are significantly related to youth hw.

```{r}
# hw best practices: youth report
m3 <- lm(y_hw ~ acadhwi_1 + checkhwi_1 + struct_1 + convfut_1 + knowschool_1 + parnets_1 + welcome_1 + endorse_1 + supported_1 + ptrel_1, d1)
summary(m3) # no significant relations

m3_cov <- lm(y_hw ~ parent_eng_comf + youth_age + youth_gender + acadhwi_1 + checkhwi_1 + struct_1 + convfut_1 + knowschool_1 + parnets_1 + welcome_1 + endorse_1 + supported_1 + ptrel_1, d1)
summary(m3_cov) # no significant relations
```

The expected youth-reported parent support value is 2.60, when we consider average values of all predictors in the model and student is male.

For every one-point increase in parents' knowing school, youth-reported parent support increases, on average, 0.43 points

```{r}
# parent's support: youth report
m4 <- lm(y_parsupp ~ acadhwi_1 + checkhwi_1 + struct_1 + convfut_1 + knowschool_1 + parnets_1 + welcome_1 + endorse_1 + supported_1 + ptrel_1, d1)
summary(m4) # knowing school positive 0.02

m4_cov <- lm(y_parsupp ~ parent_eng_comf + youth_age + youth_gender + acadhwi_1 + checkhwi_1 + struct_1 + convfut_1 + knowschool_1 + parnets_1 + welcome_1 + endorse_1 + supported_1 + ptrel_1, d1)
summary(m4_cov) # knowing school positive 0.03
```

1 Nada cómodo
2 Un poco cómodo
3 Moderadamente cómodo
4 Cómodo
5 Muy cómodo

```{r}
ggplot(d1, aes(parent_eng_comf)) +
geom_bar(alpha = 0.7)
```
The expected youth-reported parent involvement value is 3.29, when we consider average values of all predictors in the model and student is male.

For every one-point increase in parents' english comfort, youth-reported parent involvement decreases, on average, 0.17 and points. (is this interpretation correct?)

```{r}
# parental involvement: youth report
m5 <- lm(y_parinv ~ acadhwi_1 + checkhwi_1 + struct_1 + convfut_1 + knowschool_1 + parnets_1 + welcome_1 + endorse_1 + supported_1 + ptrel_1, d1)
summary(m5) # no relations

m5_cov <- lm(y_parinv ~ parent_eng_comf + youth_age + youth_gender + acadhwi_1 + checkhwi_1 + struct_1 + convfut_1 + knowschool_1 + parnets_1 + welcome_1 + endorse_1 + supported_1 + ptrel_1, d1)
summary(m5_cov) # parent eng conf negative and significant < 0.01, trends: knowschool < 0.10, welcome < 0.10
```

The expected parent-reported self-efficacy value is 0.39, when we consider average values of all predictors in the model and student is male.

For every one-point increase in parents' feeling welcome and endorsing school, parent self efficacy increases, on average, 0.23 and 0.45 points, respectively. 

Also, For every one-point increase in parents' comfort with speaking english at school, parent's self efficacy increases by 0.05 points

```{r}
# self efficacy: parent report
m6 <- lm(selfeffic_1 ~ acadhwi_1 + checkhwi_1 + struct_1 + convfut_1 + knowschool_1 + parnets_1 + welcome_1 + endorse_1 + supported_1 + ptrel_1, d1)
summary(m6) # know school, welcome, endorse, ptrel trending

m6_cov <- lm(selfeffic_1 ~ parent_eng_comf + youth_age + youth_gender + acadhwi_1 + checkhwi_1 + struct_1 + convfut_1 + knowschool_1 + parnets_1 + welcome_1 + endorse_1 + supported_1 + ptrel_1, d1)
summary(m6_cov) # par eng conf positive significant 0.03, welcome positive 0.05, endorse positive < 0.001
```

The expected parent-reported youth general grades value is 3.22, when we consider average values of all predictors in the model and student is male.

For every one-point increase in parents' knowing school, parent-reported youth general grades increases, on average, 0.76 points.

When student is female, parent-report of general grades decreases, on average, 0.66 points

```{r}
# grades in general: parent report
m7 <- lm(gen_grades ~ acadhwi_1 + checkhwi_1 + struct_1 + convfut_1 + knowschool_1 + parnets_1 + welcome_1 + endorse_1 + supported_1 + ptrel_1, d1)
summary(m7) # know school trending, conv fut negative and significant 0.01

m7_cov <- lm(gen_grades ~ parent_eng_comf + youth_age + youth_gender + acadhwi_1 + checkhwi_1 + struct_1 + convfut_1 + knowschool_1 + parnets_1 + welcome_1 + endorse_1 + supported_1 + ptrel_1, d1)
summary(m7_cov) # know school positive and significant 0.02, gender negative and significant 0.001
```

Other parent report outcomes: 
help_extra		range: 0-1 (binary)
attend_pta		range: 0-1 (binary)
sats_contact 	range: 1-4 (multiple)

Can a composite be made with these? 