---
title: "relations paper"
output: html_document
date: "2025-01-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rio)
library(here)
library(tidyverse)
library(haven)
library(janitor)
library(knitr)
library(surveytoolbox)
library(sjPlot)
library(kableExtra)
library(psych)
library(sjmisc)
library(MVN)

theme_set(theme_minimal())
```

### The dataset ###
This dataset has items 82, 83, 131, 133, 136 reverse scored. Look at "creating dataset.Rmd" 

```{r}
parent_w1 <- read_sav(here("outcomes paper/data_outcomes", "primary_parent_w1.sav"))
```
A relational question not included in poster

En esta escuela, siento que hay por lo menos un maestro…
64.	… que se preocupa por mi joven. 
65.	… quien está interesado en conocerme 
66.	… con quien me siento/a cómodo/a hablando sobre mi joven. 
67.	… de quien puedo hacer preguntas o hacer sugerencias sobre mi joven. 

```{r}
par_teach_rel <- parent_w1 %>%
  select(q64, q65, q66, q67)

alpha(par_teach_rel) # 0.87
```


Looking at variables that could serve as alternative outcomes

*These suggest traditional parental involvement 
q15 - help with youth extracurricular activities (binary: 0, 1)
q16 - frequency of this help (likert: 1-4.Item seems faulty, though: responses on q15 of 0 have responses in q16 from 1-2 & options 2-3 are virtually the same)
q17 - attendance to pta organized events (binary: 0, 1)
q18 - frequency of contact with teacher (likert: 1-4; could index issues)
q19 - frequency of contact with other school staff member (likert: 1-4; could even more so, index issues)
q20 - satisfaction with contact any school staff (likert: 1-4)

Could a composite variable be made using binary and likert response items? (q15, q17, q20)

*Other forms of youth school outcomes
q12_1 - how youth does in LA (likert: 1-5; higher is worse)
q12_2 - how youth does in SOC STUD (likert: 1-5; higher is worse)
q12_3 - how youth does in MATH (likert: 1-5; higher is worse)
q12_4 - how youth does in SCIENCE (likert: 1-5; higher is worse)
q14 - overall how youth does academically (likert: 1-8; higher is worse)


```{r}

```

YOUTH 

*alternative outcomes
q4 - overall grades
q7-q15 homework questions (i like q8, q9, q10)

*youth's perception of parental support (in the last 3 months...) 
q18 - find ways to take part in school activities (likert: 1-4)
q19 - find ways to do volunteer activities in the community (likert: 1-4)
q20 - take leadership roles in school or in the community (likert: 1-4)
q21 - do your best in school (likert: 1-4)
q22 - make a schedule for completing homework and school projects on time (likert: 1-4)

*youth's perception of parental involvement (in the last 3 months...) 
q36.	… visited your school for a special event (activity nights, choir performance, dances, etc.) (likert: 1-4)
q37.	… attended a parent-teacher conference (likert: 1-4)
q38.	… attended a parent group meeting (likert: 1-4)
q39.	…volunteered for a school activity, event, or project (likert: 1-4)
q40.	… had at least some type of contact with at least one of your teachers (by telephone, text, email, or an in-person meeting)? (likert: 1-4)

