---
title: "relations paper"
output: html_document
date: "2025-01-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rio)
library(here)
library(tidyverse)
library(haven)
library(janitor)
library(knitr)
library(surveytoolbox)
library(sjPlot)
library(kableExtra)
library(psych)
library(sjmisc)
library(MVN)
library(stats)
library(AICcmodavg)

theme_set(theme_minimal())
```

### The dataset ###
This dataset has items 82, 83, 131, 133, 136 reverse scored. Look at "creating dataset.Rmd" 

```{r}
parent_w1 <- read_sav(here("outcomes paper/data_outcomes", "primary_parent_w1.sav"))
```
A relational question not included in poster due to EFA solution showing difficulties

En esta escuela, siento que hay por lo menos un maestro…
64.	… que se preocupa por mi joven. 
65.	… quien está interesado en conocerme 
66.	… con quien me siento/a cómodo/a hablando sobre mi joven. 
67.	… de quien puedo hacer preguntas o hacer sugerencias sobre mi joven. 

```{r}
par_teach_rel <- parent_w1 %>%
  select(q64, q65, q66, q67)

alpha(par_teach_rel) # 0.87
```


Looking at variables that could serve as alternative outcomes

*These suggest traditional parental involvement 
q15 - help with youth extracurricular activities (binary: 0, 1)
q16 - frequency of this help (likert: 1-4.Item seems faulty, though: responses on q15 of 0 have responses in q16 from 1-2 & options 2-3 are virtually the same)
q17 - attendance to pta organized events (binary: 0, 1)
q18 - frequency of contact with teacher (likert: 1-4; could index issues)
q19 - frequency of contact with other school staff member (likert: 1-4; could even more so, index issues)
q20 - satisfaction with contact any school staff (likert: 1-4)

Could a composite variable be made using binary and likert response items? (q15, q17, q20)

*Other forms of youth school outcomes
q12_1 - how youth does in LA (likert: 1-5; higher is worse)
q12_2 - how youth does in SOC STUD (likert: 1-5; higher is worse)
q12_3 - how youth does in MATH (likert: 1-5; higher is worse)
q12_4 - how youth does in SCIENCE (likert: 1-5; higher is worse)
q14 - overall how youth does academically (likert: 1-8; higher is worse)


```{r}

```

YOUTH 

*alternative outcomes
q4 - overall grades
q7-q15 homework questions (i like q8, q9, q10)

6.	I have a specific place (i.e., after school program, my desk at home, etc.) for doing my homework ____
7.	I have a specific time for doing my homework ____
8.	When I do my homework I turn off my cell phone and other distractions ____
9.	Doing homework is an important part of my daily routine ____
10.	Doing a good job on my homework is important to me ____
11.	Doing well in school is important to me ____
12.	If I have challenges with certain parts of my homework, I take active steps to get the help I need ____
13.	I have a daily planner or other way of keeping my homework organized ____
14.	I can rely on my parents to help me make plans and take steps toward my future education and career goals ____
15.	My parents and I argue about my homework ___


*youth's perception of parental support (in the last 3 months...) 
q18 - find ways to take part in school activities (likert: 1-4)
q19 - find ways to do volunteer activities in the community (likert: 1-4)
q20 - take leadership roles in school or in the community (likert: 1-4)
q21 - do your best in school (likert: 1-4)
q22 - make a schedule for completing homework and school projects on time (likert: 1-4)

*youth's perception of parental involvement (in the last 3 months...) 
q36.	… visited your school for a special event (activity nights, choir performance, dances, etc.) (likert: 1-4)
q37.	… attended a parent-teacher conference (likert: 1-4)
q38.	… attended a parent group meeting (likert: 1-4)
q39.	…volunteered for a school activity, event, or project (likert: 1-4)
q40.	… had at least some type of contact with at least one of your teachers (by telephone, text, email, or an in-person meeting)? (likert: 1-4)


```{r}
d1 <- read_sav(here("outcomes paper/data_outcomes", "p_y_w1_for_analysesjan25.sav"))
```

```{r}
only_scales <- d1 %>%
  select(acadhwi_1, checkhwi_1, checkhwi_1, struct_1, convfut_1, knowschool_1, selfeffic_1, parnets_1, welcome_1, endorse_1, supported_1, ptrel_1)

descrip_table <- descriptives(only_scales) %>%
  rename(c("Scale" = "column"))
```


```{r}
library(Hmisc)
rcorr(as.matrix(only_scales), type = "pearson")
```

```{r}
m <- lm(y_sch_eng ~ acadhwi_1 + checkhwi_1 + struct_1 + convfut_1 + knowschool_1 + selfeffic_1 + parnets_1 + welcome_1 + endorse_1 + supported_1 + ptrel_1, d1)

summary(m) # only checking hw came out significant 0.01

```



```{r}
m2 <- lm(grades ~ acadhwi_1 + checkhwi_1 + struct_1 + convfut_1 + knowschool_1 + selfeffic_1 + parnets_1 + welcome_1 + endorse_1 + supported_1 + ptrel_1, d1)

summary(m2) # only conversation about future came out significant 0.02
```
```{r}
m3 <- lm(y_hw ~ acadhwi_1 + checkhwi_1 + struct_1 + convfut_1 + knowschool_1 + selfeffic_1 + parnets_1 + welcome_1 + endorse_1 + supported_1 + ptrel_1, d1)

summary(m3) # only selfefficacy came out significant 0.05
```
```{r}
m4 <- lm(y_parsupp ~ acadhwi_1 + checkhwi_1 + struct_1 + convfut_1 + knowschool_1 + selfeffic_1 + parnets_1 + welcome_1 + endorse_1 + supported_1 + ptrel_1, d1)

summary(m4) # knowing school positive 0.001, struct though negative 0.09 trend

```
```{r}
m5 <- lm(y_parinv ~ acadhwi_1 + checkhwi_1 + struct_1 + convfut_1 + knowschool_1 + selfeffic_1 + parnets_1 + welcome_1 + endorse_1 + supported_1 + ptrel_1, d1)

summary(m5) # know school a trend 0.06
```
```{r}
m6 <- lm(selfeffic_1 ~ acadhwi_1 + checkhwi_1 + struct_1 + convfut_1 + knowschool_1 + parnets_1 + welcome_1 + endorse_1 + supported_1 + ptrel_1, d1)

summary(m6)
```

add the following covariates: parent_eng_comf + youth_age + youth_gender

acknowledge that if kids were born in the US, that parents have been living in the US for at least 12-14 years, which is different for someone who just got here, but is oregon. 


