---
title: "sample description"
author: "Ale Garcia"
date: "2025-02-25"
output: html_document
---
This markdown is to run demographic analyses. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rio)
library(here)
library(tidyverse)
library(haven)
library(janitor)
library(knitr)
library(surveytoolbox)
library(sjPlot)
library(kableExtra)
library(psych)
library(sjmisc)

theme_set(theme_minimal())
```


### The dataset ###

```{r}
d <- read_sav(here("outcomes paper/data_outcomes", "p_y_w1_for_analysesjan25.sav")) # 94 rows

# view_df(d)

data.frame(table(d$family_id)) # no dupped rows
```

```{r}
d1 <- d %>%
  mutate(school_id = factor(school_id),
         condition = factor(condition),
         parent_role = factor(parent_role),
         country_born = factor(country_born),
         age_arrival_us = as.numeric(age_arrival_us),
         parent_ed = factor(parent_ed),
         youth_age = as.numeric(youth_age),
         youth_country_born = factor(youth_country_born),
         adults_at_home = as.numeric(adults_at_home),
         minors_at_home = as.numeric(minors_at_home),
         grade_lev = factor(grade_lev),
         youth_gender = factor(youth_gender))
```

# descriptive stats function
```{r include=FALSE}

# function that takes a df and returns a df with only the numeric columns
only_numeric <- function(df) {
  select_numeric <- dplyr::select_if(df, is.numeric)
  return(select_numeric)
}

# a list with the functions we want
smry <- list(n = function(x) length(x),
             n_valid = function(x) sum(!is.na(x)),
             na = function(x) sum(is.na(x)),
             M = function(x) round(mean(x, na.rm = TRUE), 2), 
             SD = function(x) round(sd(x, na.rm = TRUE), 2),
             Min = function(x) round(min(x, na.rm = TRUE), 2),
             Max = function(x) round(max(x, na.rm = TRUE), 2))

# wrapping solution in a function
descriptives <- function(df) {
  select_numeric <- only_numeric(df)
  mean_sd <- map_df(select_numeric, function(col) map_df(smry, ~.x(col)),
       .id = "column")
  return(mean_sd)
}
```


```{r}
descrip_table <- descriptives(d1) %>%
  rename(c("Scale" = "column"))

descrip_table
```

# function to take proportions
```{r}
prop_fun <- function(df, var) {
  df %>%
    group_by({{var}}) %>%
    summarize(n = n()) %>%
    mutate(prop = round(n/sum(n), 2))
}
```

# various
```{r}
# parents
prop_fun(d1, parent_role)
prop_fun(d1, country_born)
prop_fun(d1, parent_ed)
prop_fun(d1, white)
prop_fun(d1, latino)
prop_fun(d1, indigenous_lat_am)
prop_fun(d1, parent_eng_comf)
prop_fun(d1, h_lang_eng)
prop_fun(d1, h_lang_spa)
prop_fun(d1, h_lang_ind)
```
Educational level
1 Nunca asistí a la escuela
2 Asistí la escuela primaria/elementaría hasta el _____ grado
3 Escuela primaria/elementaría (jardín de niños hasta el quinto grado)
4 Escuela intermedia (6-8 grado)
5 Asistí la secundaria hasta el ____ grado
6 Escuela secundaria (grado 9 hasta 12)
7 GED
8 Recibí un curso de entrenamiento especializado en negocios, secretaria u otro entrenamiento técnico 
9 Título de Asociado en Artes (dos años)
10 Licenciatura
11 Licenciatura avanzada: Maestría o un Doctorado # no participant had this level of ed

re-categorized as:
1 -> 1, 2, 3:  elementary or less
2 -> 4, 5: middle school
3 -> 6, 7, 8: high school and/or certificate
4 -> 9, 10:  associate, bachelor, graduate degree - post-sec

```{r}
# Youth
prop_fun(d1, youth_country_born)
prop_fun(d1, y_latino)
prop_fun(d1, y_indigenous_americas)
prop_fun(d1, y_race_ethnicity_no_response)
prop_fun(d1, grade_lev) 
prop_fun(d1, youth_gender)
prop_fun(d1, y_homel_eng)
prop_fun(d1, y_homel_spa)
```






