---
title: ""
author: ""
date: ""
output: html_document
---

```{r setup, include=FALSE}

library(here)
library(haven)
library(tidyverse)
library(psych)

options(scipen = 999)
knitr::opts_chunk$set(echo = TRUE)
```

<h2 style="text-align: center;">Exploratory Factor Analysis - Structural and Relational Parental Practices Among Latino Immigran</h2>

```{r, include=FALSE}
parent_w1 <- read_sav(here("outcomes paper/data_outcomes", "primary_parent_w1.sav"))
```

## Homework Involvement (15 items)

Please indicate how often you have done the following to help your child with homework in the past three months...

84 I made sure they had a set time to do their homework.

85 I made sure they had a set place to do their homework.

86 I sat with my child while they did their homework.

87 I made myself available when my child had questions about their homework.

88 I helped my child when they were stuck on an assignment.

89 I helped my child prepare for exams by asking questions about the topic, etc.

90 I gave my child incentives like special privileges or gifts.

91 I used consequences, punishment, or disciplinary actions.

92 I reminded my child to do their homework.

93 I checked in with my child to make sure they did their homework.

94 I checked and corrected my child’s homework.

95 I checked with the teacher to make sure my child completed their homework.

96 I helped my child access other resources (library, art materials, computer lab, etc.).

97 I helped my child practice skills (reading, reviewing, practice problems, etc.).

98 I limited noise and/or distractions while they did their homework.

```{r}
hw_inv <- parent_w1 %>%
  select(q84, q85, q86, q87, q88, q89, q90, q91, q92, q93, q94, q95, q96, q97, q98)
```

### Kaiser-Meyer-Olkin factor sampling adequacy (> 0.5)
```{r}
KMO(hw_inv)
```
### Bartlett’s Test of Sphericity (p < .05)
```{r}
poly <- polychoric(hw_inv)$rho
cortest.bartlett(poly, n = nrow(hw_inv))
```
### Poly corr matrix
```{r}
cor.plot(poly, numbers = T, upper = F, main = "Polychoric", show.legend = F)
```

### Scree plot
```{r}
scree(hw_inv, factors = TRUE, pc = FALSE, main = "Scree plot", hline = NULL, add = FALSE)
```

### Removing items with very low correlations
```{r}
hw_inv <- parent_w1 %>%
  select(q84, q85, q86, q87, q88, q89, q90, q93, q94, q95, q96, q97, q98) # removing 91 and 92
```

### Kaiser-Meyer-Olkin factor sampling adequacy (> 0.5)
```{r}
KMO(hw_inv)
```
### Bartlett’s Test of Sphericity (p < .05)
```{r}
poly <- polychoric(hw_inv)$rho
cortest.bartlett(poly, n = nrow(hw_inv))
```
### Poly corr matrix
```{r}
cor.plot(poly, numbers = T, upper = F, main = "Polychoric", show.legend = F)
```

### Scree plot
```{r}
scree(hw_inv, factors = TRUE, pc = FALSE, main = "Scree plot", hline = NULL, add = FALSE)
```

### EFA using ULS 2 factors
```{r}
factor_test_uls <- fa(hw_inv, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 2) 
factor_test_uls
```

### Removing crossloading items  
```{r}
hw_inv <- parent_w1 %>%
  select(q84, q85, q86, q87, q88, q89, q93, q94, q96, q97) # removing 90, 95, and 98
```

### Kaiser-Meyer-Olkin factor sampling adequacy (> 0.5)
```{r}
KMO(hw_inv)
```
### Bartlett’s Test of Sphericity (p < .05)
```{r}
poly <- polychoric(hw_inv)$rho
cortest.bartlett(poly, n = nrow(hw_inv))
```
### Poly corr matrix
```{r}
cor.plot(poly, numbers = T, upper = F, main = "Polychoric", show.legend = F)
```

### Scree plot
```{r}
scree(hw_inv, factors = TRUE, pc = FALSE, main = "Scree plot", hline = NULL, add = FALSE)
```


### 2nd EFA iteration
```{r}
factor_test_uls <- fa(hw_inv, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 2) 
factor_test_uls
fa.diagram(factor_test_uls)
```


### Factor 1: Tutoring homework

86 I sat with my child while they did their homework.

87 I made myself available when my child had questions about their homework.

88 I helped my child when they were stuck on an assignment.

89 I helped my child prepare for exams by asking questions about the topic, etc.

94 I checked and corrected my child’s homework.

96 I helped my child access other resources (library, art materials, computer lab, etc.).

97 I helped my child practice skills (reading, reviewing, practice problems, etc.).
```{r}
hw_inv_fa1 <- parent_w1 %>%
  select(q86, q87, q88, q89, q94, q96, q97)

# alpha
alpha(hw_inv_fa1)
```


### Factor 2: checking homework

84 I made sure they had a set time to do their homework.

85 I made sure they had a set place to do their homework.

93 I checked in with my child to make sure they did their homework.
```{r}
hw_inv_fa2 <- parent_w1 %>%
  select(q84, q85, q93)

# alpha
alpha(hw_inv_fa2)
```


## Home Structure 

In the past three months, I have taken steps to support my child to...

23 ... develop a plan for studying and doing homework.

24 ... find ways to get involved in school activities.

25 ... find ways to get involved in volunteer activities in the community.

26 ... set goals for the school year.

27 ... meet deadlines.

28 ... take on leadership roles in school or the community.

29 ... focus on doing their best.

30 ... develop a schedule to complete homework and school projects on time.

```{r}
struct <- parent_w1 %>%
  select(q23, q24, q25, q26, q27, q28, q29, q30) 
```


### Kaiser-Meyer-Olkin factor sampling adequacy (> 0.5)
```{r}
KMO(struct)
```


### Bartlett’s Test of Sphericity (p < .05)
```{r}
poly <- polychoric(struct)$rho
cortest.bartlett(poly, n = nrow(struct))
```


### Poly corr matrix
```{r}
cor.plot(poly, numbers = T, upper = F, main = "Polychoric", show.legend = F)
```

### Scree plot
```{r}
scree(struct, factors = TRUE, pc = FALSE, main = "Scree plot", hline = NULL, add = FALSE)
```

### EFA using ULS - 1 factor
```{r}
factor_test_uls <- fa(struct, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 1) 
factor_test_uls
fa.diagram(factor_test_uls)
```

### Factor: Home Structure
```{r}
alpha(struct)
```


## School Involvement

In general, I make an effort to...

31 ... get to know school staff and administrators.

32 ... get to know at least one of my child’s teachers.

33 ... understand school rules and policies.

34 ... inform myself about my rights as a parent.

35 ... learn about the educational system in this state.

36 ... understand the difference between earning a GED, graduating with a standard high school diploma, an international high school diploma, or an International Baccalaureate (IB) diploma.

37 ... get involved in school activities, in the classroom, and/or in other ways (e.g., parent organizations, volunteer work, etc.).

38 ... have conversations with other parents to gather information or learn about school resources.

39 ... reach out to other parents for support.

40 ... understand the pathway to college and career readiness.

41 ... attend parent-teacher conferences when available.

```{r}
scho_inv <- parent_w1 %>%
  select(q31, q32, q33, q34, q35, q36, q37, q38, q39, q40, q41)
```

### Kaiser-Meyer-Olkin factor sampling adequacy (> 0.5)
```{r}
KMO(scho_inv)
```


### Bartlett’s Test of Sphericity (p < .05)
```{r}
poly <- polychoric(scho_inv)$rho
cortest.bartlett(poly, n = nrow(scho_inv))
```


### Poly corr matrix
```{r}
cor.plot(poly, numbers = T, upper = F, main = "Polychoric", show.legend = F)
```

### Scree plot
```{r}
scree(scho_inv, factors = TRUE, pc = FALSE, main = "Scree plot", hline = NULL, add = FALSE)
```

### EFA using ULS 2 factors
```{r}
factor_test_uls <- fa(scho_inv, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 2) 
factor_test_uls
```


### Removing crossloading items
```{r}
scho_inv <- parent_w1 %>%
  select(q31, q32, q33, q34, q36, q37, q38, q39) # removed 35, 40, 41
```


### Kaiser-Meyer-Olkin factor sampling adequacy (> 0.5)
```{r}
KMO(scho_inv)
```


### Bartlett’s Test of Sphericity (p < .05)
```{r}
poly <- polychoric(scho_inv)$rho
cortest.bartlett(poly, n = nrow(scho_inv))
```


### Poly corr matrix
```{r}
cor.plot(poly, numbers = T, upper = F, main = "Polychoric", show.legend = F)
```

### Scree plot
```{r}
scree(scho_inv, factors = TRUE, pc = FALSE, main = "Scree plot", hline = NULL, add = FALSE)
```

### 2nd EFA iteration 
```{r}
factor_test_uls <- fa(scho_inv, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 2) 
factor_test_uls
fa.diagram(factor_test_uls)
```


### Factor 1: School familiarity

31 ... get to know school staff and administrators.

32 ... get to know at least one of my child’s teachers.

33 ... understand school rules and policies.

34 ... inform myself about my rights as a parent.
```{r}
scho_inv_fa1 <- parent_w1 %>%
  select(q31, q32, q33, q34)

# alpha
alpha(scho_inv_fa1)
```

### Factor 2: Proactive school enagagement

36 ... understand the difference between earning a GED, graduating with a standard high school diploma, an international high school diploma, or an International Baccalaureate (IB) diploma.

37 ... get involved in school activities, in the classroom, and/or in other ways (e.g., parent organizations, volunteer work, etc.).

38 ... have conversations with other parents to gather information or learn about school resources.

39 ... reach out to other parents for support.
```{r}
scho_inv_fa2 <- parent_w1 %>%
  select(q36, q37, q38, q39)

# alpha
alpha(scho_inv_fa2)
```


## School belonging

As a parent at this school, I feel...

47 ... like part of a community with school staff and other parents.

48 ... treated with respect, knowing that my opinions matter.

49 ... comfortable asking for a translator and for materials in Spanish if I need them.

50 ... able to talk to teachers or administrators about major concerns related to my child.

51 ... committed to creating a successful environment for all children.

52 ... happy that my child attends this school.

53 ... welcomed at my child’s school.


```{r}
belong <- parent_w1 %>%
  select(q47, q48, q49, q50, q51, q52, q53)
```

### Kaiser-Meyer-Olkin factor sampling adequacy (> 0.5)
```{r}
KMO(belong)
```


### Bartlett’s Test of Sphericity (p < .05)
```{r}
poly <- polychoric(belong)$rho
cortest.bartlett(poly, n = nrow(belong))
```

### Poly corr matrix
```{r}
cor.plot(poly, numbers = T, upper = F, main = "Polychoric", show.legend = F)
```

### Scree plot
```{r}
scree(belong, factors = TRUE, pc = FALSE, main = "Scree plot", hline = NULL, add = FALSE)
```


### EFA using ULS 1 factor
```{r}
factor_test_uls <- fa(belong, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 1) 
factor_test_uls
fa.diagram(factor_test_uls)
```
### Factor: school belonging
```{r}
alpha(belong)
```

## School endorsement

As a parent at this school, I am confident that...

54 ... this school is a good place for my child.

55 ... my child’s school staff are doing good things for them.

56 ... people at my child’s school are trustworthy.

57 ... my child’s school does a good job preparing students for their futures.

```{r}
endorse <- parent_w1 %>%
  select(q54, q55, q56, q57)
```

### Kaiser-Meyer-Olkin factor sampling adequacy (> 0.5)
```{r}
KMO(endorse)
```


### Bartlett’s Test of Sphericity (p < .05)
```{r}
poly <- polychoric(endorse)$rho
cortest.bartlett(poly, n = nrow(endorse))
```


### Poly corr matrix
```{r}
cor.plot(poly, numbers = T, upper = F, main = "Polychoric", show.legend = F)
```

### Scree plot
```{r}
scree(endorse, factors = TRUE, pc = FALSE, main = "Scree plot", hline = NULL, add = FALSE)
```
### EFA using ULS 1 factor
```{r}
factor_test_uls <- fa(endorse, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 1) 
factor_test_uls
```

### Removing item Ultra-Heywood case
```{r}
endorse <- parent_w1 %>%
  select(q54, q56, q57) # removed 55
```


### Kaiser-Meyer-Olkin factor sampling adequacy (> 0.5)
```{r}
KMO(endorse)
```


### Bartlett’s Test of Sphericity (p < .05)
```{r}
poly <- polychoric(endorse)$rho
cortest.bartlett(poly, n = nrow(endorse))
```


### Poly corr matrix
```{r}
cor.plot(poly, numbers = T, upper = F, main = "Polychoric", show.legend = F)
```

### Scree plot
```{r}
scree(endorse, factors = TRUE, pc = FALSE, main = "Scree plot", hline = NULL, add = FALSE)
```

### 2nd EFA iteration 
```{r}
factor_test_uls <- fa(endorse, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 1) 
factor_test_uls
fa.diagram(factor_test_uls)
```
### Factor: school endorsement
```{r}
alpha(endorse)
```

## Parent-teacher relationship

At this school, I feel there is at least one teacher...

64 ... who cares about my child.

65 ... who is interested in getting to know me.

66 ... with whom I feel comfortable talking about my child.

67 ... from whom I can ask questions or make suggestions about my child.


```{r}
pt_rel <- parent_w1 %>%
  select(q64, q65, q66, q67)
```


### Kaiser-Meyer-Olkin factor sampling adequacy (> 0.5)
```{r}
KMO(pt_rel)
```


### Bartlett’s Test of Sphericity (p < .05)
```{r}
poly <- polychoric(pt_rel)$rho
cortest.bartlett(poly, n = nrow(pt_rel))
```


### Poly corr matrix
```{r}
cor.plot(poly, numbers = T, upper = F, main = "Polychoric", show.legend = F)
```

```{r}
scree(pt_rel, factors = TRUE, pc = FALSE, main = "Scree plot", hline = NULL, add = FALSE)
```

## EFA using ULS 
```{r}
factor_test_uls <- fa(pt_rel, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 1) 
factor_test_uls
```

### Removing item Ultra-Heywood case
```{r}
pt_rel <- parent_w1 %>%
  select(q64, q65, q67) # removed 66
```


### 2nd EFA iteration 

```{r}
factor_test_uls <- fa(pt_rel, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 1) 
factor_test_uls # it only works with ml (jaime suggest removing the scale and adding it as a limitation and future directions)
```

Another Ultra-Heywood case. Removing the item would yield a factor with 2 items. Conclusion: EFA not possible with this scale.
```{r}
pt_rel <- parent_w1 %>%
  select(q64, q65, q66, q67)

alpha(pt_rel)
```

