---
title: ""
author: ""
date: ""
output: html_document
---

```{r setup, include=FALSE}

library(here)
library(haven)
library(tidyverse)
library(psych)

options(scipen = 999)
knitr::opts_chunk$set(echo = TRUE)
```

<h2 style="text-align: center;">Exploratory Factor Analysis - Structural and Relational Parental Practices Among Latino Immigran</h2>

```{r, include=FALSE}
parent_w1 <- read_sav(here("outcomes paper/data_outcomes", "primary_parent_w1.sav"))
```

## Homework Involvement (15 items)

Please indicate how often you have done the following to help your child with homework in the past three months...

84 I made sure they had a set time to do their homework.

85 I made sure they had a set place to do their homework.

86 I sat with my child while they did their homework.

87 I made myself available when my child had questions about their homework.

88 I helped my child when they were stuck on an assignment.

89 I helped my child prepare for exams by asking questions about the topic, etc.

90 I gave my child incentives like special privileges or gifts.

91 I used consequences, punishment, or disciplinary actions.

92 I reminded my child to do their homework.

93 I checked in with my child to make sure they did their homework.

94 I checked and corrected my child’s homework.

95 I checked with the teacher to make sure my child completed their homework.

96 I helped my child access other resources (library, art materials, computer lab, etc.).

97 I helped my child practice skills (reading, reviewing, practice problems, etc.).

98 I limited noise and/or distractions while they did their homework.

```{r}
hw_inv <- parent_w1 %>%
  select(q84, q85, q86, q87, q88, q89, q90, q91, q92, q93, q94, q95, q96, q97, q98)
```

### Kaiser-Meyer-Olkin factor sampling adequacy (> 0.5)
```{r}
KMO(hw_inv)
```
### Bartlett’s Test of Sphericity (p < .05)
```{r}
poly <- polychoric(hw_inv)$rho
cortest.bartlett(poly, n = nrow(hw_inv))
```
### Poly corr matrix
```{r}
cor.plot(poly, numbers = T, upper = F, main = "Polychoric", show.legend = F)
```

### Scree plot
```{r}
scree(hw_inv, factors = TRUE, pc = FALSE, main = "Scree plot", hline = NULL, add = FALSE)
```

########### testing ###########


```{r}
hw_inv_test <- parent_w1 %>%
  select(q86, q87, q88, q89, q90, q93, q94, q95, q96, q97, q98)

alpha(hw_inv_test) # .79
```

### Kaiser-Meyer-Olkin factor sampling adequacy (> 0.5)
```{r}
KMO(hw_inv_test)
```
### Bartlett’s Test of Sphericity (p < .05)
```{r}
poly <- polychoric(hw_inv_test)$rho
cortest.bartlett(poly, n = nrow(hw_inv_test))
```
### Poly corr matrix
```{r}
cor.plot(poly, numbers = T, upper = F, main = "Polychoric", show.legend = F)
```

### Scree plot
```{r}
scree(hw_inv_test, factors = TRUE, pc = FALSE, main = "Scree plot", hline = NULL, add = FALSE)
```

### EFA using ULS 2 factors
```{r}
factor_test_uls <- fa(hw_inv_test, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 2) 
factor_test_uls
```

fa 1

87. Me hice disponible para mi joven cuando tenía preguntas sobre su tarea. ___
88. Ayudé a mi joven cuando estaba atascado/a en una tarea ___
89. Ayudé a mi joven a prepararse para los exámenes haciéndole preguntas sobre el tema, etc. ___
86. Me senté con mi joven mientras él / ella hacía su tarea ___ SEMI
96. Le ayudé a mi joven acceder a otros recursos (biblioteca, materiales de arte, laboratorio de computación, etc.) ___ SEMI
97. Le ayudé a mi joven practicar habilidades (lectura, revisión, problemas de práctica, etc.) ___ SEMI
98. Limité el ruido y/o las distracciones mientras hacia su tarea ___

90. Le di a mi joven incentivos como privilegios especiales, regalos especiales, etc. ____ CROSS



fa 2

93. Comprobé con mi joven para asegurarme de que él / ella hiciera su tarea ___
95. Comprobé con la maestra/o para asegurarme que mi joven terminó su tarea ___
94. Comprobé y corregí la tarea de mi hija/o ___ SEMI



90. Le di a mi joven incentivos como privilegios especiales, regalos especiales, etc. ____ CROSS


ONE FACTOR: q86, q87, q88, q89, q90, q93, q94, q95, q96, q97, q98


86. Me senté con mi joven mientras él / ella hacía su tarea ___ SEMI
87. Me hice disponible para mi joven cuando tenía preguntas sobre su tarea. ___
88. Ayudé a mi joven cuando estaba atascado/a en una tarea ___
89. Ayudé a mi joven a prepararse para los exámenes haciéndole preguntas sobre el tema, etc. ___
90. Le di a mi joven incentivos como privilegios especiales, regalos especiales, etc. ____ CROSS
93. Comprobé con mi joven para asegurarme de que él / ella hiciera su tarea ___
94. Comprobé y corregí la tarea de mi hija/o ___ SEMI
95. Comprobé con la maestra/o para asegurarme que mi joven terminó su tarea ___
96. Le ayudé a mi joven acceder a otros recursos (biblioteca, materiales de arte, laboratorio de computación, etc.) ___ SEMI
97. Le ayudé a mi joven practicar habilidades (lectura, revisión, problemas de práctica, etc.) ___ SEMI
98. Limité el ruido y/o las distracciones mientras hacia su tarea ___


stil TBD


########### testing ###########







### EFA using ULS 2 factors
```{r}
factor_test_uls <- fa(hw_inv, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 2) 
factor_test_uls
```


### Removing items with loadings < 0.3  or cross-loading
```{r}
hw_inv_1 <- parent_w1 %>%
  select(q84, q85, q86, q87, q88, q89, q93, q94, q96, q97)
```


### kmo Kaiser-Meyer-Olkin factor sampling adequacy (> 0.5)
```{r}
KMO(hw_inv_1)
```

### Bartlett’s Test of Sphericity (P < .05)
```{r}
poly <- polychoric(hw_inv_1)$rho
cortest.bartlett(poly, n = nrow(hw_inv_1))
```

### Poly corr matrix
```{r}
cor.plot(poly, numbers = T, upper = F, main = "Polychoric", show.legend = F)
```

### Scree plot
```{r}
scree(hw_inv_1, factors = TRUE, pc = FALSE, main = "Scree plot", hline = NULL, add = FALSE)
```


### 2nd EFA iteration: 2 factors
```{r}
factor_test_uls <- fa(hw_inv_1, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 2) 
factor_test_uls

fa.diagram(factor_test_uls)
```


### Factor 1: Tutoring homework
86	Me senté con mi joven mientras él / ella hacía su tarea.  
87	Me hice disponible para mi joven cuando tenía preguntas sobre su tarea.  
88	Ayudé a mi joven cuando estaba atascado/a en una tarea.  
89	Ayudé a mi joven a prepararse para los exámenes haciéndole preguntas sobre el tema, etc.  
94	Comprobé y corregí la tarea de mi hija/o.   
96	Le ayudé a mi joven acceder a otros recursos (biblioteca, materiales de arte, laboratorio de computación, etc.).  
97	Le ayudé a mi joven practicar habilidades (lectura, revisión, problemas de práctica, etc.).  
```{r}
hw_inv_fa1 <- parent_w1 %>%
  select(q86, q87, q88, q89, q94, q96, q97)

# alpha
alpha(hw_inv_fa1)
```


### Factor # 2: checking homework
84	Me aseguré de que tenga un cierto tiempo para hacer la tarea.  
85	Me aseguré de que él / ella tenga un cierto lugar para hacer la tarea.  
93	Comprobé con mi joven para asegurarme de que él / ella hiciera su tarea.  
```{r}
hw_inv_fa2 <- parent_w1 %>%
  select(q84, q85, q93)

# alpha
alpha(hw_inv_fa2)
```


## Home Structure 

In the past three months, I have taken steps to support my child to...

23 ... develop a plan for studying and doing homework.

24 ... find ways to get involved in school activities.

25 ... find ways to get involved in volunteer activities in the community.

26 ... set goals for the school year.

27 ... meet deadlines.

28 ... take on leadership roles in school or the community.

29 ... focus on doing their best.

30 ... develop a schedule to complete homework and school projects on time.

```{r}
struct <- parent_w1 %>%
  select(q23, q24, q25, q26, q27, q28, q29, q30) 
```


### Kaiser-Meyer-Olkin factor sampling adequacy (> 0.5)
```{r}
KMO(struct)
```


### Bartlett’s Test of Sphericity (p < .05)
```{r}
poly <- polychoric(struct)$rho
cortest.bartlett(poly, n = nrow(struct))
```


### Poly corr matrix
```{r}
cor.plot(poly, numbers = T, upper = F, main = "Polychoric", show.legend = F)
```

### Scree plot
```{r}
scree(struct, factors = TRUE, pc = FALSE, main = "Scree plot", hline = NULL, add = FALSE)
```

## EFA using ULS - 1 factor
```{r}
factor_test_uls <- fa(struct, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 1) 
factor_test_uls
fa.diagram(factor_test_uls)
```

## Factor: Home Structure
```{r}
alpha(struct)
```


## School Involvement

In general, I make an effort to...

31 ... get to know school staff and administrators.

32 ... get to know at least one of my child’s teachers.

33 ... understand school rules and policies.

34 ... inform myself about my rights as a parent.

35 ... learn about the educational system in this state.

36 ... understand the difference between earning a GED, graduating with a standard high school diploma, an international high school diploma, or an International Baccalaureate (IB) diploma.

37 ... get involved in school activities, in the classroom, and/or in other ways (e.g., parent organizations, volunteer work, etc.).

38 ... have conversations with other parents to gather information or learn about school resources.

39 ... reach out to other parents for support.

40 ... understand the pathway to college and career readiness.

41 ... attend parent-teacher conferences when available.



```{r}
scho_inv <- parent_w1 %>%
  select(q31, q32, q33, q34, q35, q36, q37, q38, q39, q40, q41)
```

### Kaiser-Meyer-Olkin factor sampling adequacy (> 0.5)
```{r}
KMO(scho_inv)
```


### Bartlett’s Test of Sphericity (p < .05)
```{r}
poly <- polychoric(scho_inv)$rho
cortest.bartlett(poly, n = nrow(scho_inv))
```


### Poly corr matrix
```{r}
cor.plot(poly, numbers = T, upper = F, main = "Polychoric", show.legend = F) # item 41 with 37-39 neg corrs.
```

### Scree plot
```{r}
scree(scho_inv, factors = TRUE, pc = FALSE, main = "Scree plot", hline = NULL, add = FALSE)
```


### Removing item with correlations < 0.3
```{r}
scho_inv <- parent_w1 %>%
  select(q31, q32, q33, q34, q35, q36, q37, q38, q39, q40) # removed 41
```


### Kaiser-Meyer-Olkin factor sampling adequacy (> 0.5)
```{r}
KMO(scho_inv)
```


### Bartlett’s Test of Sphericity (p < .05)
```{r}
poly <- polychoric(scho_inv)$rho
cortest.bartlett(poly, n = nrow(scho_inv))
```


### Poly corr matrix
```{r}
cor.plot(poly, numbers = T, upper = F, main = "Polychoric", show.legend = F)
```

### Scree plot
```{r}
scree(scho_inv, factors = TRUE, pc = FALSE, main = "Scree plot", hline = NULL, add = FALSE)
```

### EFA using ULS 2 factors
```{r}
factor_test_uls <- fa(scho_inv, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 2) 
factor_test_uls
fa.diagram(factor_test_uls)
```

### Removing crossloading item  
```{r}
scho_inv <- parent_w1 %>%
  select(q31, q32, q33, q34, q36, q37, q38, q39, q40) # removed 41
```


### Kaiser-Meyer-Olkin factor sampling adequacy (> 0.5)
```{r}
KMO(scho_inv)
```


### Bartlett’s Test of Sphericity (p < .05)
```{r}
poly <- polychoric(scho_inv)$rho
cortest.bartlett(poly, n = nrow(scho_inv))
```


### Poly corr matrix
```{r}
cor.plot(poly, numbers = T, upper = F, main = "Polychoric", show.legend = F)
```


### Scree plot
```{r}
scree(scho_inv, factors = TRUE, pc = FALSE, main = "Scree plot", hline = NULL, add = FALSE)
```

### 2nd EFA iteration 
```{r}
factor_test_uls <- fa(scho_inv, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 2) 
factor_test_uls
fa.diagram(factor_test_uls)
```

### Factor 1: School familiarity
```{r}
scho_inv_fa1 <- parent_w1 %>%
  select(q31, q32, q33, q34)

# alpha
alpha(scho_inv_fa1)
```

### Factor 2: Proactive school enagagement
```{r}
scho_inv_fa2 <- parent_w1 %>%
  select(q36, q37, q38, q39, q40)

# alpha
alpha(scho_inv_fa2)
```


## School belonging

As a parent at this school, I feel...

47 ... like part of a community with school staff and other parents.

48 ... treated with respect, knowing that my opinions matter.

49 ... comfortable asking for a translator and for materials in Spanish if I need them.

50 ... able to talk to teachers or administrators about major concerns related to my child.

51 ... committed to creating a successful environment for all children.

52 ... happy that my child attends this school.

53 ... welcomed at my child’s school.


```{r}
comfort <- parent_w1 %>%
  select(q47, q48, q49, q50, q51, q52, q53)
```

### Kaiser-Meyer-Olkin factor sampling adequacy (> 0.5)
```{r}
KMO(comfort)
```


### Bartlett’s Test of Sphericity (p < .05)
```{r}
poly <- polychoric(comfort)$rho
cortest.bartlett(poly, n = nrow(comfort))
```

### Poly corr matrix
```{r}
cor.plot(poly, numbers = T, upper = F, main = "Polychoric", show.legend = F) # all healthy corrs
```

### Scree plot
```{r}
scree(comfort, factors = TRUE, pc = FALSE, main = "Scree plot", hline = NULL, add = FALSE)
```


### EFA using ULS 2 factors
```{r}
factor_test_uls <- fa(comfort, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 1) 
factor_test_uls
```

## Factor 1: belong
48.	…tratado/a con respeto, sabiendo que mis opiniones son importantes. ____
49.	…cómoda/o preguntando por un traductor y por materiales en español si los necesito. ____
52.	 …feliz de que mi joven asista a esta escuela. ___
53.	…bienvenido/a en la escuela de mi joven. ____
```{r}
comfort_fa1 <- parent_w1 %>%
  select(q48, q49, q52, q53)
# alpha
alpha(comfort_fa1) # 0.86
```

## Factor 2: included (not included in multiple regression analyses)
47.	…parte de una comunidad con el personal de la escuela y los otros padres. ____
50.	… capaz de hablar con maestros o administradores sobre grandes preocupaciones relacionados con mi joven. ___
51.	…dedicada/o en creando un ambiente exitoso para todos los jóvenes. ____

```{r}
comfort_fa2 <- parent_w1 %>%
  select(q47, q50, q51)
# alpha
alpha(comfort_fa2) # 0.70
```



Issues: 
Homework involvement -- low correlations in the first matrix. 

School involvement: given that I took out parallel analysis, I no longer have a reason to donducxt a 2-factor model because scree plot clearly show 1-factor. 


## School endorsement

As a parent at this school, I am confident that...

54 ... this school is a good place for my child.

55 ... my child’s school staff are doing good things for them.

56 ... people at my child’s school are trustworthy.

57 ... my child’s school does a good job preparing students for their futures.

```{r}
endorse <- parent_w1 %>%
  select(q54, q55, q56, q57)
```

### Kaiser-Meyer-Olkin factor sampling adequacy (> 0.5)
```{r}
KMO(endorse)
```


### Bartlett’s Test of Sphericity (p < .05)
```{r}
poly <- polychoric(endorse)$rho
cortest.bartlett(poly, n = nrow(endorse))
```


### Poly corr matrix
```{r}
cor.plot(poly, numbers = T, upper = F, main = "Polychoric", show.legend = F) # corrs above .90 potentially multicolleniearity
```

### Scree plot
```{r}
scree(endorse, factors = TRUE, pc = FALSE, main = "Scree plot", hline = NULL, add = FALSE)
```
### EFA using ULS 1 factor
```{r}
factor_test_uls <- fa(endorse, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 1) 
factor_test_uls
```

### Removing item Ultra-Heywood case
```{r}
endorse <- parent_w1 %>%
  select(q54, q56, q57) # removed 55
```


### Kaiser-Meyer-Olkin factor sampling adequacy (> 0.5)
```{r}
KMO(endorse)
```


### Bartlett’s Test of Sphericity (p < .05)
```{r}
poly <- polychoric(endorse)$rho
cortest.bartlett(poly, n = nrow(endorse))
```


### Poly corr matrix
```{r}
cor.plot(poly, numbers = T, upper = F, main = "Polychoric", show.legend = F) # corrs above .90 potentially multicolleniearity
```

### Scree plot
```{r}
scree(endorse, factors = TRUE, pc = FALSE, main = "Scree plot", hline = NULL, add = FALSE)
```

### 2nd EFA iteration 
```{r}
factor_test_uls <- fa(endorse, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 1) 
factor_test_uls
fa.diagram(factor_test_uls)
```




## Parent-teacher relationship

At this school, I feel there is at least one teacher...

64 ... who cares about my child.

65 ... who is interested in getting to know me.

66 ... with whom I feel comfortable talking about my child.

67 ... from whom I can ask questions or make suggestions about my child.


```{r}
pt_rel <- parent_w1 %>%
  select(q64, q65, q66, q67)
```


### Kaiser-Meyer-Olkin factor sampling adequacy (> 0.5)
```{r}
KMO(pt_rel)
```


### Bartlett’s Test of Sphericity (p < .05)
```{r}
poly <- polychoric(pt_rel)$rho
cortest.bartlett(poly, n = nrow(pt_rel))
```


### Poly corr matrix
```{r}
cor.plot(poly, numbers = T, upper = F, main = "Polychoric", show.legend = F)
```

```{r}
scree(pt_rel, factors = TRUE, pc = FALSE, main = "Scree plot", hline = NULL, add = FALSE)
```

## EFA using ULS 
```{r}
factor_test_uls <- fa(pt_rel, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 1) 
factor_test_uls
```

### Removing item Ultra-Heywood case
```{r}
pt_rel <- parent_w1 %>%
  select(q64, q65, q67)# removed 66
```


### 2nd EFA iteration 
```{r}
factor_test_uls <- fa(pt_rel, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 1) 
factor_test_uls
fa.diagram(factor_test_uls)
```

Conclusion: EFA not possible with this scale.

