---
title: "Notes on duplicated ids"
output: html_notebook
---

When reviewing duplicated ids in parent's w2, I used the following code to explore if id was present in both w1 and w2:

```{r}
# id 108
p_w1_clean_6 %>%
  filter(id == 108)
p_w2_raw %>%
  filter(id == 108)
```

Then I checked with Eli and he reviewed the attendance and call logs. These were the notes:

# notes on w2

id 108 is duplicated, appears as 2 moms. In w1 there is just one 108 that is a mom. 
R_37vhYjz8eRmGcWR = is id 108
R_1rCXE5pBSje0twj = is id 109

--

ids: 112 after q145 empty. Is this the same person that started twice or two different parents?  The pattern of responses is identical. Or was it duplicated for some reason? In W1 it appears as if there were a mom and a dad. This means that the dad dropped from w1 to w2? ---> which row to delete?

dad and mom appear to have dropped by w2: negative experience with assessor, no success after multiple attemps. Mom dropped in midst of w2.

about duplicative: R_2sdrLRMKZvR9fwO is the duplicative = delete this row. 

There is no id 112 in w3, so they appear to have dropped completely from study. 

Given circumstances, do we want this data? --> Heather said: yes, keep row and let the missing data be missing. 

--

id 113: no responses at all: qualtrics var progress: 3%. Participant dropped? In w1 there are mom and dad. ---> delete row?

this was a false start. Family dropped (classical attrition) no response after 50 calls. = delete row. 
--

ids 308: 4 instances = 2 false starts, 2 completed, but in attempt 3 after q_12_2 is filled with -99 throughout. Is this one parent or two parents? 	R_UMEbJn4yl8Wc0wx is the one with completed data. This id corresponds to just 1 mom in w1. 

R_1EQFk2EbJW61nhf = false start: delete row
R_XMroI3biv8jxEVb = false start: delete row
R_3PZzUFRxLeeW6GJ = false start: delete row

R_UMEbJn4yl8Wc0wx = complete data. keep only this row. 

--

ids 314 is duplicated, appears as 2 moms. In w1 there is just one 314 that is a mom.

R_1omSIb1Wdy6hIkW is the true 314 mom 
R_2EpkLxRpA7GW0oE is 312 mom

--

id 415 is a 3 parent family. 
R_1PRjrvEzZRgtKKD = stepmom
R_2b2T5ywwHkmb0d7 = dad
R_2e9e8oXgMz4MiDt = biomom
--

ids 601: is triplicated: 2 moms, 1 dad. In w1 there is 1 mom, 1 dad. 

R_2sZpDvTZRNJfl7Z = 601 mom
R_O6DQfVORSAtnWbT = 601 dad

R_vNctgxUZcfcMwFj = says is w3. Eli couldn't find it as belonging to any other participant's id in w2. (check if this response id exists in w3). 
--

ids 201, 204, 206, 207, 216, 302, 310, 408, 414, 418, 501, 503, 505, 506, 607, 608, 610, 613, 614, 615: logs confirms that these are two-parents fams. Also these are two parent fams in w1 and w2.

Need to confirm if 109 and 601 are two-parent families? They are 2-parent fams in w1, though. 

--
new variable: 217_4_text? The question in this variable was completely changed by w2, levels also changed (no 99 and added a 4 for year money).

-- 
to check how many moms/dad/other and # of families I used:
```{r}
nrow(distinct(p_w2_clean_3, family_id))

data.frame(table(p_w2_clean_3$family_id))
```
But not sure how to know the number of two-parent families. 

## previous notes (undated):
#############################
-- wave correct?
R_2qyriEL0QoD8Kln	id 604 in wave 1, says wave = 1

R_6FiHWtagxFiXgbf	id 604 in wave 2, says wave = 1. CHANGE to wave 2, assesor mistakenly put wave 1 and should have put wave 2.  

-- new variable
new variable: 217_4_text? The question in this variable was completely changed by w2, levels also changed (no 99 and added a 4 for yearly money). 
#############################

## August 26 2022 -- meeting with Eli notes --
#########################

# notes on w2 meeting with Eli on 08.26.22
id 109: in dataset appears as a 2 parent (1 mom, 1 dad) -- Eli confirms that 109 is a family of 2, mome and dad, however the qualtrics code doesn't match from logs: (SV_0i9InOoo0RH6o9T) to what we have in dataset (R_z1iP1gjIGExS3Pr)

# In wave 1 (logs):
R_1Lwe3h7XQeF5hyj	id 601  mom, wave 1
R_UG9FQ6VvanpWwwN	id 601	dad, wave 1

# In wave 2 (logs): 
R_2sZpDvTZRNJfl7Z	id 601	mom, wave 2 = real mom
R_O6DQfVORSAtnWbT	id 601	dad, wave 2 = real dad 

R_vNctgxUZcfcMwFj	id 601  mom, wave 3 = the theory was that this row leaked from w3, but id 601 in W3 already has two id associated with 601. This still needs to be figured out. ## hence action plan is delete from wave 2 ##

# In wave 3 (logs): 
R_2QRPBKfzZwFDAaT id 601	mom, confirmed
R_3LhFdFrq0SFkkKI id 601	dad, confirmed

# From Eli (08.26.22)
R_2QRPBKfzZwFDAaT W3 601 Mom from email from JCG

# From notes I had before 08.26.22 
R_3LhFdFrq0SFkkKI	601	dad, wave 3	(in wave 3 log)
R_20lyHTLoqvQ4Rpz	601 mom, wave 3 (not sure where this came from)

R_vNctgxUZcfcMwFj	id 601  mom --> mistery what this response id belongs to
R_20lyHTLoqvQ4Rpz	601 mom, wave 3  --> mistery what this response id belongs to
these 2 rows are very similar, but not identical = Eli & I think that it is the same data, but assessor error (08.26.22: not sure where we were looking, these two ids do not appear on logs) 
