---
title: "elt"
author: "Alejandra Garcia Isaza"
date: "2/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rio)
library(here)
library(tidyverse)
library(haven)

library(surveytoolbox)
library(sjPlot)
```


```{r}

w1_raw <- haven::read_sav(here("nopublish", "ELT W1 ERC 11.11.2020.sav"))

w1_raw
```

```{r}

elt_w1_clean <- w1_raw %>% 
  janitor::clean_names() %>%
  select(-1:-8, -13:-17, -202) %>% # deleting columns with metadata
  rename(c("id" = "pj")) %>%
  arrange(id) # ordering participants ids

elt_w1_clean
```

```{r}

# fixing duplicates ids
duplicated(elt_w1_clean$id) # id_number: 257 is dupplicated and there is no 254

elt_w1_clean <- elt_w1_clean %>%
  mutate(id = case_when(response_id == "R_6EELe7Uuwi9W7zX" ~ "254",
                        TRUE ~ as.character(id))) %>%
  arrange(id)

duplicated(elt_w1_clean$id) # testing that there are no more id duplicates == No duplicates!

elt_w1_clean
```

```{r}

# fixing wrong assignment
elt_w1_clean_2 <- elt_w1_clean %>%
  mutate(school_id = str_sub(id, 1, 1),
         condition = case_when(
           school_id == "1" | school_id == "3" | school_id == "5" ~ "1",
           school_id == "2" | school_id == "4" | school_id == "6" ~ "2"
         )) %>%
  select(school_id, condition, everything()) %>%
  add_column(wave = 1, .before = 10) %>%
  select(- school)

elt_w1_clean_2
```

```{r}

# making variables numeric and adding labels  
elt_w1_clean_3 <- elt_w1_clean_2 %>%
  mutate(condition = as.numeric(condition),
         condition = set_vall(condition, c("control" = 1, "intervention" = 2)),
         school_id = as.numeric(school_id),
         school_id = set_vall(school_id, c("cascade" = 1, "prairie_mountain" = 2, "kelly" = 3, "ata" = 4, "briggs" = 5, "agnes_stewart" = 6))) # this worked!!!

elt_w1_clean_3

view_df(elt_w1_clean_3) # nice plot that serves as codebook
```

```{r}

# exporting to .sav to check proper coding of condition and labels
elt_w1_clean_3 %>%
  haven::write_sav(here("nopublish", "elt_w1_clean_3.sav"))
```

Results:
- This would be the master dataset
- school_id matchs assigned condition properly!
- Labels were properly exported (both variable labels and level labels on .sav file)

```{r}

# collapsing variable english_comfort q132_1_<<choice>>

d <- elt_w1_clean_3 %>% 
    select(id, starts_with("q132_1")) %>% 
    pivot_longer(
      cols = starts_with("q132_1"),
      names_to = "item",
      values_to = "english_comfort",
      values_drop_na = TRUE) %>% 
   filter(english_comfort == 1) %>%
  mutate(english_comfort = case_when(item == "q132_1_4" ~ "4",
                                     item == "q132_1_3" ~ "3",
                        TRUE ~ as.character(english_comfort))) %>%
  select(-item)

x <- left_join(elt_w1_clean_3, d)

x_2 <- x %>%
  mutate(english_comfort = as.numeric(english_comfort),
         english_comfort = set_vall(english_comfort, c("not at all comfortable" = 1, "somewhat comfortable" = 2, "comfortable" = 3, "very comfortable" = 4, "no response" = 99)))


view_df(x_2)

# x_2 %>%
#   haven::write_sav(here("nopublish", "x_2.sav")) # it worked, the .sav file has all 43 participants with the proper scale and the proper level labels. 
```


```{r}
# collapsing language variables q132_<<language>>_<<choice>> 

# collapsing english variables
eng <- elt_w1_clean_3 %>% 
  select(id, starts_with("q132_1")) %>% 
    pivot_longer(
      cols = starts_with("q132_1"),
      names_to = "item_1",
      values_to = "english_comfort",
      values_drop_na = TRUE) %>% 
   filter(english_comfort == 1) %>%
  mutate(english_comfort = case_when(item_1 == "q132_1_1" ~ "1",
                                     item_1 == "q132_1_2" ~ "2",
                                     item_1 == "q132_1_3" ~ "3",
                                     item_1 == "q132_1_4" ~ "4",
                                     item_1 == "q132_1_99" ~ "99",
                        TRUE ~ as.character(english_comfort)))  %>%
  select(-item_1)

duplicated(eng$id) # no duplicates

# collapsing spanish variables
spa <- elt_w1_clean_3 %>%
  select(id, starts_with("q132_2")) %>% 
    pivot_longer(
      cols = starts_with("q132_2"),
      names_to = "item_2",
      values_to = "spanish_comfort",
      values_drop_na = TRUE) %>% 
   filter(spanish_comfort == 1) %>%
  mutate(spanish_comfort = case_when(item_2 == "q132_2_1" ~ "1",
                                     item_2 == "q132_2_2" ~ "2",
                                     item_2 == "q132_2_3" ~ "3",
                                     item_2 == "q132_2_4" ~ "4",
                                     item_2 == "q132_2_99" ~ "99",
                        TRUE ~ as.character(spanish_comfort))) %>%
  select(-item_2) 

duplicated(spa$id) # id 454 is duplicated. In w1_raw it shows that this participant chose options 1 and 2 (not sure if they could choose more than one option in Qulatrics). 

# fising duplicate
spa_2 <- spa %>% 
  distinct(id, .keep_all = TRUE) # Option 1, "not at all comfortable" was chosen because the duplicate was the second option.  

duplicated(spa_2$id) # no duplicates

# joining english and spanish variables
langs <- left_join(eng, spa_2) 

# collapsing other language 1 variables
other_1 <- elt_w1_clean_3 %>% 
  select(id, starts_with("q132_3"), -q132_3_text) %>% 
    pivot_longer(
      cols = starts_with("q132_3"),
      names_to = "item_3",
      values_to = "other1_lang_comfort",
      values_drop_na = TRUE) %>% 
   filter(other1_lang_comfort == 1) %>%
  mutate(other1_lang_comfort = case_when(item_3 == "q132_3_1" ~ "1",
                                    item_3 == "q132_3_2" ~ "2",
                                    item_3 == "q132_3_3" ~ "3",
                                    item_3 == "q132_3_4" ~ "4",
                                    item_3 == "q132_3_99" ~ "99",
                        TRUE ~ as.character(other1_lang_comfort)))  %>%
  select(-item_3)


duplicated(other_1$id) # no duplicates

# collapsing other language 2 variables
other_2 <- elt_w1_clean_3 %>% 
  select(id, starts_with("q132_4"), -q132_4_text) %>% 
    pivot_longer(
      cols = starts_with("q132_4"),
      names_to = "item_4",
      values_to = "other2_lang_comfort",
      values_drop_na = TRUE) %>% 
   filter(other2_lang_comfort == 1) %>%
  mutate(other2_lang_comfort = case_when(item_4 == "q132_4_1" ~ "1",
                                    item_4 == "q132_4_2" ~ "2",
                                    item_4 == "q132_4_3" ~ "3",
                                    item_4 == "q132_4_4" ~ "4",
                                    item_4 == "q132_4_99" ~ "99",
                        TRUE ~ as.character(other2_lang_comfort)))  %>%
  select(-item_4)


duplicated(other_2$id) # no duplicates

# joining other 1 and 2 language variables
others <- left_join(other_1, other_2) 


# joining dataframes with all language variables
lang_vars <- left_join(langs, others)

duplicated(lang_vars$id) # no duplicates

# renaming variables with text input in master dataset (n = 43)
lang_inputs <- elt_w1_clean_3 %>%
  select(id, q132_3_text, q132_4_text) %>%
  rename(c("other1_lang" = "q132_3_text"), c("other2_lang" = "q132_4_text"))

duplicated(lang_inputs$id) # no duplicates

# joining renamed master dataset with all language variables now collapsed to "recover" id 153 data
all_lang_vars <- left_join(lang_inputs, lang_vars)

duplicated(all_lang_vars$id) # no duplicates

# recordering all language vars
all_lang_vars <- all_lang_vars %>%
  select(id, english_comfort, spanish_comfort, other1_lang, other1_lang_comfort, other2_lang, everything())

# creating new dataset that joins master dataset with reordered and collapsed lang vars
elt_w1_clean_4 <- left_join(elt_w1_clean_3, all_lang_vars) %>%
  select(-starts_with("q132_")) # deleting all previous language variables that were included in new variables # 

duplicated(elt_w1_clean_4$id) # no duplicates, n = 43

# adding value labels on language vars
elt_w1_clean_4 <- elt_w1_clean_4 %>%
  mutate(english_comfort = as.numeric(english_comfort),
         english_comfort = set_vall(english_comfort, c("not at all comfortable" = 1, "somewhat comfortable" = 2, "comfortable" = 3, "very comfortable" = 4, "no response" = 99)),
         spanish_comfort = as.numeric(spanish_comfort),
         spanish_comfort = set_vall(spanish_comfort, c("not at all comfortable" = 1, "somewhat comfortable" = 2, "comfortable" = 3, "very comfortable" = 4, "no response" = 99)),
         other1_lang_comfort = as.numeric(other1_lang_comfort),
         other1_lang_comfort = set_vall(other1_lang_comfort, c("not at all comfortable" = 1, "somewhat comfortable" = 2, "comfortable" = 3, "very comfortable" = 4, "no response" = 99)),
         other2_lang_comfort = as.numeric(other2_lang_comfort),
         other2_lang_comfort = set_vall(other2_lang_comfort, c("not at all comfortable" = 1, "somewhat comfortable" = 2, "comfortable" = 3, "very comfortable" = 4, "no response" = 99)),
         )

view_df(elt_w1_clean_4)

```

```{r}

# renaming demo vars
elt_w1_clean_5 <- elt_w1_clean_4 %>%
  rename(c("age" = "q127_1_text"), 
         c("birth_country" = "q128"),
         c("another_birth_country_text" = "q128_2_text"),
         c("age_first_moved_us" = "q129_1_text"),
         c("white" = "q130_1"),
         c("hispanic_latino_spanish" = "q130_2"),
         c("black_african_american" = "q130_3"),
         c("asian" = "q130_4"),
         c("american_indian_alaska_native" = "q130_5"),
         c("indigenous_americas" = "q130_6"),
         c("middle_eastern_north_african" = "q130_7"),
         c("native_hawaiian_pacific_islander" = "q130_8"),
         c("race_ethnicity_other" = "q130_9"),
         c("race_ethnicity_no_response" = "q130_99"),
         c("indigenous_americas_text" = "q130_6_text"),
         c("race_ethnicity_other_text" = "q130_9_text"),
         c("gender_id" = "q131"),
         c("years_in_position" = "q133"),
         c("years_in_school" = "q134"),
         c("equity_leadership" = "q135_1"),
         c("cultural_responsiveness" = "q135_2"),
         c("restorative_practices" = "q135_3"),
         c("diversity" = "q135_4"),
         c("ell" = "q135_5"),
         c("cont_ed_other" = "q135_6"),
         c("cont_ed_na" = "q135_88"),
         c("cont_ed_no_response" = "q135_99"),
         c("cont_ed_other_text" = "q135_6_text")) %>%
  select(-q127, -q129, -q131_3_text) # deleted: q127, q129, q131_3_text because they did not have meaningful info

view_df(elt_w1_clean_5)

```

# Notes on w1 dataset

- Participant 257 was duplicated and 254 was missing. This was fixed.
- participant 153 didn't complete survey, after q50 it's empty. 
- Total participants in wave 1 n = 43

```{r}
# exporting to .sav to check proper coding of condition and labels
# elt_w1_clean_5 %>%
#   haven::write_sav(here("nopublish", "elt_w1_clean_5.sav")) # it worked, variable and value labels were succesfully exported.
```

# Notes on Missing values

88 = "not applicable"
99 = "no response"
-99 = seen but unanswered questions -- (e.g questions that ask participants to respond if they have knowledge about classroom)
NA = unseen question, either because participant didn't complete survey (e.g. id 153) or because of skip logic (they were not teachers or had classroom knowledge) --> NA appear in SPSS as (.)
blank spaces = not sure, but prob. irrelevant

```{r}
# recoding missing values as N/A with function

# vector with missing values in dataset
missing_vals <- c(88, 99, -99)

# function that returns true if values in vector are equal to missing_vals. The function takes a vector x, and specified values of missing data
recode_missing <- function(x, missing_vals = c(88, 99, -99)) {
  test <- x %in% missing_vals
  ifelse(test, NA, x)
}

# function that recodes missing values to NA. The function takes a dataframe with variables with missing data, and specified values of missing data
recode_missing_df <- function(df, missing_vals = c(88, 99, -99)) {
  modify(df, ~recode_missing(.x, missing_vals)) # here uses the function created above
}

elt_w1_clean_6 <- recode_missing_df(elt_w1_clean_5) 

view_df(elt_w1_clean_6) # notice that applying this strips out labels
```

# Scales




# Common guidelines for evaluating Cronbach's Alpha
.00 to .69 = Poor
.70 to .79 = Fair 
.80 to .89 = Good 
.90 to .99 = Excellent/Strong

```{r}
####### internal consistency ######

library(psych)

climate_gen <- data.frame(elt_w1_clean_6[,12:35]) # q1 - q14
alpha(climate_gen) # alpha = .92 -->  excellent consistency. Porbably just because the sheer number of items in scale (26)

# these are trying to follow the scales of the Kettering measure
respect <- elt_w1_clean_6 %>%
  select(q1, q10, q11, q21, q24) %>%
  data.frame()
alpha(respect) # alpha = .78 -->  fair consistency

trust <- elt_w1_clean_6 %>%
  select(q5, q6, q7, q8) %>%
  data.frame()
alpha(trust) # alpha = .61 -->  poor consistency

# it's too hard to find ways to make them hang together. Ask Heather. 

equity_self_eff <- data.frame(elt_w1_clean_6[,61:66]) # q50 - q55
alpha(equity_self_eff) # alpha = .89 -->  good consistency

scho_lat_fam_rel <- data.frame(elt_w1_clean_6[,67:75]) # q56 - q64
alpha(scho_lat_fam_rel) # alpha = . --> .84 good consistency

teach_lat_fam_rel <- elt_w1_clean_6 %>%
  select(q65, q66, q67, q68, q69, q73, q74, q75, q76) %>%
  data.frame()
alpha(teach_lat_fam_rel) # q73 needs to be reverse coded!!

teach_self_eff <- data.frame(elt_w1_clean_6[,119:128]) # q108 - q117
alpha(teach_self_eff) # several need to be reverse coded!!
```












q1, q2, q3, q4, q5, q6, q7, q8, q9, q10, q11,  q12, q13, q14, q15, q16, q17, q18, q19, q20, q21, q22, q23, q24, q25, q26, q27, q28, q29, q30, q31, q32, q33, q34, q35, q36, q37, q38, q39, q40, q41, q42, q43, q44, q45, q46, q47, q48, q49, q50, q51, q52, q53, q54, q55, q56, q57, q58, q59, q60, q61, q62, q63, q64, q65, q66, q67, q68, q69, q70, q71, q72, q73, q74, q75, q76, q78, q79, q80, q81, q82, q83, q84, q85, q86, q87, q88, q89, q90, q91, q92, q93, q94, q95, q96, q97, q98, q99, q100, q101, q102, q103, q104, q105, q106
q108, q109, q110, q111, q112, q113, q114, q115, q116, q117
q118, q119, q120, q121, q122, q123, q124
