---
title: "elt_3_waves"
author: "Alejandra Garcia Isaza"
date: "5/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rio)
library(here)
library(tidyverse)
library(haven)
library(janitor)
library(knitr)
library(surveytoolbox)
library(sjPlot)
library(kableExtra)
library(psych)

```

## WAVE 1

# Loading the data
```{r}
w1_raw_elt <- read_sav(here("nopublish", "ELT W1 ERC 11.11.2020.sav"))
```

# Initial cleaning
```{r}
elt_w1_clean <- w1_raw_elt %>% 
  clean_names() %>% # function that formats variables' names
  select(-1:-8, -10:-17, -202) %>% # selecting out columns with metadata
  rename(c("id" = "pj")) %>% # renaming id variable.
  arrange(id) # ordering participants ids in descending order

# fixing wrong id
elt_w1_clean <- elt_w1_clean %>%
  mutate(id = case_when(response_id == "R_6EELe7Uuwi9W7zX" ~ "254",
                        TRUE ~ as.character(id))) %>%
  arrange(id)
```

# Fixing wrong condition assignment and school id
```{r}
elt_w1_clean_2 <- elt_w1_clean %>%
  mutate(school_id = str_sub(id, 1, 1), # new school id variable
         condition = case_when(
           school_id == "1" | school_id == "3" | school_id == "5" ~ "1",
           school_id == "2" | school_id == "4" | school_id == "6" ~ "2")) %>% # new condition variable
  select(school_id, condition, everything()) %>%
  add_column(wave = 1, .before = 9) %>% # new wave variable
  select(- school) # selecting out (i.e. deleting) original faulty school variable
```

# Adding value labels and recoding response option in q68
```{r}
elt_w1_clean_3 <- elt_w1_clean_2 %>%
  mutate(condition = as.numeric(condition),
         condition = set_vall(condition, c("control" = 1, "intervention" = 2)),
         school_id = as.numeric(school_id),
         school_id = set_vall(school_id, c("cascade" = 1, "prairie_mountain" = 2, "kelly" = 3, "ata" = 4, "briggs" = 5, "agnes_stewart" = 6)),
         q68 = ifelse(q68 == 5, 99, q68),
         q68 = set_varl(q68, "When I communicate with Latino families, I keep in mind that many Latino parents may not understand how to navigate the educational system in this
country."),
         q68 = set_vall(q68, c("Strongly Disagree" = 1, "Disagree" = 2, "Agree" = 3, "Strongly Agree" = 4, "No response" = 99)))
```

# Collapsing spreaded out language variables
```{r}
lang_vars <- elt_w1_clean_3 %>% 
  select(id, starts_with("q132_1"), starts_with("q132_2"), starts_with("q132_3"), starts_with("q132_4"), -q132_4_text, -q132_3_text) %>% 
  pivot_longer(
    cols = starts_with("q132"),
    names_to = c("item", "language", "response"),
    names_sep = "_",
    values_to = "language_comfort",
    values_drop_na = TRUE) %>% 
  filter(language_comfort == 1) %>%
  mutate(language_comfort = as.numeric(language_comfort))

get_dupes(lang_vars, id, language)

# id 454 chose options 1 and 2.
lang_vars <- lang_vars %>% 
  distinct(id, language, .keep_all = TRUE) # Option 1, "not at all comfortable" was chosen because the duplicate was the second option.  

# renaming vars
lang_vars <- lang_vars %>% 
  pivot_wider(names_from = language, values_from = response) %>%
  rename(c("english_comfort" = "1"),
         c("spanish_comfort" = "2"),
         c("other1_lang_comfort" = "3"),
         c("other2_lang_comfort" = "4")) %>%
  select(-item, -language_comfort)

# renaming input var in elt_w1_clean_3 dataset
lang_inputs <- elt_w1_clean_3 %>%
  select(id, q132_3_text, q132_4_text) %>%
  rename(c("other1_lang" = "q132_3_text"), 
         c("other2_lang" = "q132_4_text"))

# joining renamed elt_w1_clean_3 dataset with all language variables now collapsed to "recover" id 153 data
all_lang_vars <- left_join(lang_inputs, lang_vars) %>%
select(id, english_comfort, spanish_comfort, other1_lang, other1_lang_comfort, other2_lang, everything())
```

# Adding value and var labels to joined dataset 
```{r}
# creating dataset that joins master dataset with reordered and collapsed lang vars
elt_w1_clean_4 <- left_join(elt_w1_clean_3, all_lang_vars) %>%
  mutate(english_comfort = as.numeric(english_comfort),
         english_comfort = set_vall(english_comfort, c("not at all comfortable" = 1, "somewhat comfortable" = 2, "comfortable" = 3, "very comfortable" = 4, "no response" = 99)),
         spanish_comfort = as.numeric(spanish_comfort),
         spanish_comfort = set_vall(spanish_comfort, c("not at all comfortable" = 1, "somewhat comfortable" = 2, "comfortable" = 3, "very comfortable" = 4, "no response" = 99)),
         other1_lang_comfort = as.numeric(other1_lang_comfort),
         other1_lang_comfort = set_vall(other1_lang_comfort, c("not at all comfortable" = 1, "somewhat comfortable" = 2, "comfortable" = 3, "very comfortable" = 4, "no response" = 99)),
         other2_lang_comfort = as.numeric(other2_lang_comfort),
         other2_lang_comfort = set_vall(other2_lang_comfort, c("not at all comfortable" = 1, "somewhat comfortable" = 2, "comfortable" = 3, "very comfortable" = 4, "no response" = 99))) %>%
  select(-starts_with("q132_")) # deleting all previous language variables that were included in new variables  
```

# Renaming demo vars
```{r}
elt_w1_clean_5 <- elt_w1_clean_4 %>%
  rename(c("age" = "q127_1_text"), 
         c("birth_country" = "q128"),
         c("another_birth_country_text" = "q128_2_text"),
         c("age_first_moved_us" = "q129_1_text"),
         c("white" = "q130_1"),
         c("hispanic_latino_spanish" = "q130_2"),
         c("black_african_american" = "q130_3"),
         c("asian" = "q130_4"),
         c("american_indian_alaska_native" = "q130_5"),
         c("indigenous_americas" = "q130_6"),
         c("middle_eastern_north_african" = "q130_7"),
         c("native_hawaiian_pacific_islander" = "q130_8"),
         c("race_ethnicity_other" = "q130_9"),
         c("race_ethnicity_no_response" = "q130_99"),
         c("indigenous_americas_text" = "q130_6_text"),
         c("race_ethnicity_other_text" = "q130_9_text"),
         c("gender_id" = "q131"),
         c("years_in_position" = "q133"),
         c("years_in_school" = "q134"),
         c("equity_leadership" = "q135_1"),
         c("cultural_responsiveness" = "q135_2"),
         c("restorative_practices" = "q135_3"),
         c("diversity" = "q135_4"),
         c("ell" = "q135_5"),
         c("cont_ed_other" = "q135_6"),
         c("cont_ed_na" = "q135_88"),
         c("cont_ed_no_response" = "q135_99"),
         c("cont_ed_other_text" = "q135_6_text")) %>%
  mutate(age = as.numeric(age)) %>% # making variable numeric for QUAN analyses
  select(-q127, -q129, -q131_3_text) # selecting out because they did not have meaningful info
```


```{r}
# final cleaned dataset wave 1
view_df(elt_w1_clean_5)
```


## WAVE 2

# Loading the data
```{r}
w2_raw_elt <- read_sav(here("nopublish", "ELT W2 ERC 11.11.2020.sav"))
```

# Initial cleaning
```{r}
elt_w2_clean <- w2_raw_elt %>% 
  clean_names() %>% # function that formats variables' names
  select(-1:-8, -10:-17, -202) %>% # selecting out columns with metadata
  rename(c("id" = "pj")) %>% # renaming id variable.
  arrange(id) # ordering participants ids in descending order

# checking for dupplicated ids
get_dupes(elt_w2_clean, id)
```

# Fixing wrong condition assignment and school id
```{r}
# this wave also had the wrong condition assignment
elt_w2_clean_2 <- elt_w2_clean %>%
  mutate(school_id = str_sub(id, 1, 1), # new school id variable
         condition = case_when(
           school_id == "1" | school_id == "3" | school_id == "5" ~ "1",
           school_id == "2" | school_id == "4" | school_id == "6" ~ "2")) %>% # new condition variable
  select(school_id, condition, everything()) %>%
  add_column(wave = 2, .before = 9) %>% # new wave variable
  select(- school) # selecting out (i.e. deleting) original faulty school variable
```

# Adding value labels and recoding response option in q68
```{r}
# this wave also had q68 with the wrong responce code
elt_w2_clean_3 <- elt_w2_clean_2 %>%
  mutate(condition = as.numeric(condition),
         condition = set_vall(condition, c("control" = 1, "intervention" = 2)),
         school_id = as.numeric(school_id),
         school_id = set_vall(school_id, c("cascade" = 1, "prairie_mountain" = 2, "kelly" = 3, "ata" = 4, "briggs" = 5, "agnes_stewart" = 6)),
         q68 = ifelse(q68 == 5, 99, q68),
         q68 = set_varl(q68, "When I communicate with Latino families, I keep in mind that many Latino parents may not understand how to navigate the educational system in this
country."),
         q68 = set_vall(q68, c("Strongly Disagree" = 1, "Disagree" = 2, "Agree" = 3, "Strongly Agree" = 4, "No response" = 99)))
```

# Collapsing spreaded out language variables
```{r}
lang_vars_w2 <- elt_w2_clean_3 %>% 
  select(id, starts_with("q132_1"), starts_with("q132_2"), starts_with("q132_3"), starts_with("q132_4"), -q132_4_text, -q132_3_text) %>% 
  pivot_longer(
    cols = starts_with("q132"),
    names_to = c("item", "language", "response"),
    names_sep = "_",
    values_to = "language_comfort",
    values_drop_na = TRUE) %>% 
  filter(language_comfort == 1) %>%
  mutate(language_comfort = as.numeric(language_comfort))

# checking for dupplicated ids
get_dupes(lang_vars_w2, id, language)

# id 155 chose options 1 and 99 for other 1.
lang_vars_w2 <- lang_vars_w2 %>% 
  distinct(id, language, .keep_all = TRUE) # Option 1, "not at all comfortable" was chosen because the duplicate was the second option.  

# renaming vars
lang_vars_w2 <- lang_vars_w2 %>% 
  pivot_wider(names_from = language, values_from = response) %>%
  rename(c("english_comfort" = "1"),
         c("spanish_comfort" = "2"),
         c("other1_lang_comfort" = "3"),
         c("other2_lang_comfort" = "4")) %>%
  select(-item, -language_comfort)

# renaming input var in elt_w1_clean_3 dataset
lang_inputs_w2 <- elt_w2_clean_3 %>%
  select(id, q132_3_text, q132_4_text) %>%
  rename(c("other1_lang" = "q132_3_text"), 
         c("other2_lang" = "q132_4_text"))

# joining renamed elt_w1_clean_3 dataset with all language variables now collapsed to "recover" id 153 data
all_lang_vars_w2 <- left_join(lang_inputs_w2, lang_vars_w2) %>%
select(id, english_comfort, spanish_comfort, other1_lang, other1_lang_comfort, other2_lang, everything())
```

# need to identify which variables is the one that id 155 is duplicated in. Explore in general this! 











```{r}
view_df(elt_w2_clean_3)
```

