---
title: "immediate outcomes juntos"
author: "Alejandra Garcia Isaza"
date: "2023-01-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rio)
library(here)
library(tidyverse)
library(haven)
library(janitor)
library(knitr)
library(surveytoolbox)
library(sjPlot)
library(kableExtra)
library(psych)
library(sjmisc)
```

### PARENT ###

# Loading the data (wave 1)
```{r}
d_raw_w1 <- read_sav(here("nopublish", "all_parent_wave1.sav")) %>%
  arrange(participant_id)
```

# Loading the data (wave 2)
```{r}
d_raw_w2 <- read_sav(here("nopublish", "all_parent_wave2.sav")) %>%
  arrange(participant_id)
```


# filtering primary caregiver only (wave 1)
```{r}
d_w1 <- d_raw_w1 %>%
  select(-response_id) %>%
  distinct(family_id, .keep_all = TRUE) # 95 primary caregivers
```

# filtering primary caregiver only (wave 2)
```{r}
d_w2 <- d_raw_w2 %>%
  select(-response_id) %>%
  distinct(family_id, .keep_all = TRUE) # 84 primary caregivers
```

id 110, 113, 209, 405, 407, 419, 508, 509, 612, 619, 620 not present in w2 = 11 primary caregiver participants lost to w2

```{r}
# adding "p1" to each variable name: "p" stands for parent, "1" for wave 1.
# names(d_w1) <- paste(names(d_w1), "p1", sep = "_")

d2_w1 <- d_w1 %>%
  rename(c("school_id" = "school_id_p1"),
         c("condition" = "condition_p1"),
         c("family_id" = "family_id_p1"),
         c("participant_id" = "participant_id_p1"))
```

```{r}
# adding "p2" to each variable name: "p" stands for parent, "2" for wave 2.
# names(d_w2) <- paste(names(d_w2), "p2", sep = "_")

d2_w2 <- d_w2 %>%
  rename(c("school_id" = "school_id_p2"),
         c("condition" = "condition_p2"),
         c("family_id" = "family_id_p2"),
         c("participant_id" = "participant_id_p2"))
```

# joining w1 and w2 (PARENT)
```{r}
parent_w1_w2 <- left_join(d2_w1, d2_w2)
```

# creating dataset with w1 and w2 for SPSS analysis (PARENT)
```{r}
# parent_w1_w2 %>%
#   haven::write_sav(here("nopublish", "parent_w1_w2.sav"))
```

# loading parent W1 & W2 dataset
```{r}
pa_w1_w2 <- read_sav(here("nopublish", "parent_w1_w2.sav"))
```


TO DO NEXT
- use code to make -99, 99, 88, 77 to NA
- create scales?

### YOUTH ###

# Loading the data (wave 1)
```{r}
y_raw_w1 <- read_sav(here("nopublish", "all_youth_wave1.sav")) %>%
  rename(c("family_id" = "id")) %>%
  arrange(family_id)
```

# Loading the data (wave 2)
```{r}
y_raw_w2 <- read_sav(here("nopublish", "all_youth_wave2.sav")) %>%
  rename(c("family_id" = "id")) %>%
  arrange(family_id)
```
id 110, 112, 113, 209, 210, 407, 410, 419, 508, 509, 606, 612, 620, 621, 623 not present in w2 = 15 participants lost to w2

```{r}
# adding "y1" to each variable name: "y" stands for youth, "1" for wave 1.
# names(y_raw_w1) <- paste(names(y_raw_w1), "y1", sep = "_")

y_w1 <- y_raw_w1 %>%
  select(-response_id_y1) %>%
  rename(c("school_id" = "school_id_y1"),
         c("condition" = "condition_y1"),
         c("family_id" = "family_id_y1")) 
```
```{r}
# adding "y1" to each variable name: "y" stands for youth, "1" for wave 1.
# names(y_raw_w2) <- paste(names(y_raw_w2), "y2", sep = "_")

y_w2 <- y_raw_w2 %>%
  select(-response_id_y2) %>%
  rename(c("school_id" = "school_id_y2"),
         c("condition" = "condition_y2"),
         c("family_id" = "family_id_y2")) 
```

# joining w1 and w2 (YOUTH)
```{r}
youth_w1_w2 <- left_join(y_w1, y_w2)
```

# creating dataset with w1 and w2 for SPSS analysis (YOUTH)
```{r}
# youth_w1_w2 %>%
#   haven::write_sav(here("nopublish", "youth_w1_w2.sav"))
```

# loading YOUTH W1 & W2 dataset
```{r}
y_w1_w2 <- read_sav(here("nopublish", "youth_w1_w2.sav"))
```

# joining w1 and w2 parent & youth (ALL)
```{r}
fam_w1_w2 <- inner_join(pa_w1_w2, y_w1_w2) # id 504 was not included 
```