---
title: "parent_3_waves"
author: "Alejandra Garcia Isaza"
date: "6/30/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rio)
library(here)
library(tidyverse)
library(haven)
library(janitor)
library(knitr)
library(surveytoolbox)
library(sjPlot)
library(kableExtra)
library(psych)
library(sjmisc)
```

################## WAVE 1 ##################

# Loading the data & initial cleaning
```{r}
p_w1_raw <- read_sav(here("nopublish", "7b. Juntos Parent W1-SPAN_February 10, 2021_10.24.sav"))

p_w1_raw <- p_w1_raw %>% 
  arrange(ID_1_TEXT) %>% # ordering participants ids
  janitor::clean_names() %>% # cleaning names
  select(-id) %>% 
  rename(c("id" = "id_1_text"))
```

# Checking duplicated ids
```{r}
data.frame(table(p_w1_raw$id)) # 29 ids have 2 or more instances each (ids 415 and 418 have 3 instances each)

dupes <- p_w1_raw %>%
  select(id, response_id) %>%
  arrange(id)
```

# Further cleaning & dealing with duplicates
```{r}
p_w1_clean <- p_w1_raw %>% 
  select(-1:-8, -10:-17, -19, -20) %>% # selecting out columns with metadata
  filter(id != "999") %>% # deleting 2 test rows
  filter(response_id != "R_31L6rXjsdcuBaz0") %>% # deleting 1 duped row "false start" (id 312)
  mutate(id = case_when(response_id == "R_12ch5JnOHKHRgre" ~ "408", # id 418 (top one) recoded as 408
                        response_id == "R_uaijWwHdl8sIOmB" ~ "317", # had a pj before id #
                        TRUE ~ as.character(id))) %>%
  arrange(id)

data.frame(table(p_w1_clean$id)) # duplicated ids match two-parent families, id 415 is a family of 3
```

# Fixing wrong condition assignment and school id
```{r}
# fixing wrong assignment
p_w1_clean_2 <- p_w1_clean %>%
  mutate(school_id = str_sub(id, 1, 1),
         condition = case_when(
           school_id == "1" | school_id == "3" | school_id == "5" ~ "1",
           school_id == "2" | school_id == "4" | school_id == "6" ~ "2"
         )) %>%
  select(school_id, condition, everything()) %>%
  select(- school) # selecting out (i.e. deleting) original faulty school variable

```

# Adding value labels and making vars numeric 
```{r}
p_w1_clean_3 <- p_w1_clean_2 %>%
  mutate(condition = as.numeric(condition),
         condition = set_vall(condition, c("control" = 1, "intervention" = 2)),
         school_id = as.numeric(school_id),
         school_id = set_vall(school_id, c("cascade" = 1, "prairie_mountain" = 2, "kelly" = 3, "ata" = 4, "briggs" = 5, "agnes_stewart" = 6)))
```

# Creating individual id variable
```{r}
p_w1_clean_3 <- p_w1_clean_3 %>% 
  mutate(participant = ifelse(response_id == "R_1ON5POXTzIXrIeb", 4, participant), # changing partic. type to correspond to records
         participant_4_text = ifelse(response_id == "R_1ON5POXTzIXrIeb", "stepmom", participant_4_text), 
         participant = as.numeric(participant),
         participant = set_vall(participant, c("mom" = 1, "dad" = 2, "tutor legal" = 3, "other, specify" = 4)),
         participant_id = paste0(id, "_", participant), .before = 5) %>% # new var as unique identifier
  rename(c("family_id" = "id")) # 125 participants, 92 moms, 32 dads, 1 grandparent. 28 families
```

# renaming vars to correspond to value labels and collapsing spreaded out variable
```{r}
# item that asks about # of adults in home was spreaded out and had an unusual coding.
p_w1_clean_3 <- p_w1_clean_3 %>%
  rename(c("q211_1" = "q211_2"),
         c("q211_2" = "q211_99"),
         c("q211_3" = "q211_4"),
         c("q211_4" = "q211_5"),
         c("q211_5" = "q211_6"),
         c("q211_99" = "q211_7"),
         c("adults_home_text" = "q211_6_text"))

# collapsing variable
num_adults <- p_w1_clean_3 %>% 
  select(participant_id, starts_with("q211_"), -adults_home_text) %>%
  pivot_longer(
    cols = starts_with("q211"),
    names_to = c("item", "adults_home"),
    names_sep = "_",
    values_to = "response",
    values_drop_na = TRUE) %>% 
  filter(response == 1) %>%
  mutate(response = as.numeric(response))

get_dupes(num_adults, participant_id)

# id 114_1 chose options 2 and 4
# id 214_1 chose options 2 and 5
# Option 2, "2" will be chosen because the {distinct} function keeps the first row. Also it can be argued that there are 2 adults in each household and not 4 and 5, respectively. 

num_adults <- num_adults %>% 
  distinct(participant_id, .keep_all = TRUE) %>% 
  select(-item, -response) 

p_w1_clean_4 <- left_join(p_w1_clean_3, num_adults) %>%
  select(-291:-296) %>%
  mutate(adults_home = as.numeric(adults_home),
         adults_home = set_vall(adults_home, c("1" = 1, "2" = 2, "3" = 3, "4" = 4, "mas de 5, ¿cuántas?" = 5, "refuse to answer" = 99)),
         adults_home = set_varl(adults_home, "Número de adultos (más de 18 años): - Selected Choice")) %>%
  move_columns(adults_home, .before = "adults_home_text")
```

# renaming demo vars
```{r}
p_w1_clean_5 <- p_w1_clean_4 %>%
  rename(c("lang_english" = "q170_1"),
         c("lang_spanish" = "q170_2"),
         c("lang_indigenous" = "q170_3"),
         c("lang_other" = "q170_4"),
         c("lang_refuse_answer" = "q170_99"),
         c("lang_indigenous_text" = "q170_3_text"),
         c("lang_other_text" = "q170_4_text"),
         c("youth_lang_broker" = "q171"),
         c("spanish_comfort_home" = "q172_1"),
         c("spanish_comfort_school" = "q172_2"),
         c("spanish_comfort_youth_school" = "q172_3"),
         c("spanish_comfort_work" = "q172_4"),
         c("spanish_comfort_friends" = "q172_5"),
         c("spanish_comfort_general" = "q172_6"),
         c("english_comfort_home" = "q173_1"),
         c("english_comfort_school" = "q173_2"),
         c("english_comfort_youth_school" = "q173_3"),
         c("english_comfort_work" = "q173_4"),
         c("english_comfort_friends" = "q173_5"),
         c("english_comfort_general" = "q173_6"),
         c("minors_home" = "q211a"),
         c("minors_home_text" = "q211a_6_text"),
         c("pub_assist" = "q212"),
         c("pub_assist_snap" = "q212a_1"),
         c("pub_assist_frpl" = "q212a_2"),
         c("pub_assist_tanf" = "q212a_3"),
         c("pub_assist_unemployment" = "q212a_4"),
         c("pub_assist_disability" = "q212a_5"),
         c("pub_assist_social_sec" = "q212a_6"),
         c("pub_assist_health_insu" = "q212a_7"),
         c("pub_assist_wic" = "q212a_18"),
         c("pub_assist_food_box" = "q212a_19"),
         c("pub_assist_other" = "q212a_10"),
         c("pub_assist_refuse_answer" = "q212a_99")) %>%
  select(-response_id) # response_id no longer needed

```

# ------ Notes on W1 dataset -------
# participants ids 415 and 418 had 3 records each. 
# participant id 415 was a family of three: 2 moms, 1 dad. One of the moms was recoded as "other, specify" in the `participant` variable and "stepmom" in the `participant_4_text` variable. 
# Participant id 418 was "triplicated" and there was no "mom" for id 408 that was only a "dad". This was fixed.
# # Total participants in wave n = 125
# ------------------------------------

```{r}
# adding "p1" to each variable name: "p" stands for parent, "1" for wave 1. 
names(p_w1_clean_5) <- paste(names(p_w1_clean_5), "p1", sep = "_")

p_w1_clean_6 <- p_w1_clean_5 %>%
  rename(c("school_id" = "school_id_p1"), 
         c("condition" = "condition_p1"),
         c("family_id" = "family_id_p1"),
         c("participant_id" = "participant_id_p1")) # renaming these variables to be able to join later
```

```{r}
view_df(p_w1_clean_6) # it appears that there are no questions about age, gender id, race/ethnicity. 
```

################## WAVE 2 ##################

# Loading the data & initial cleaning
```{r}
p_w2_raw <- read_sav(here("nopublish", "7bb. Juntos Parent W2-SPAN_February 10, 2021_10.25.sav"))

p_w2_raw <- p_w2_raw %>% 
  arrange(ID_1_TEXT) %>% # ordering participants ids
  janitor::clean_names() %>% # cleaning names
  select(-id) %>% 
  rename(c("id" = "id_1_text"))

p_w2_raw %>%
  filter(finished == 0)

# id 112
p_w2_raw %>%
  filter(id == 112)

p_w1_clean_6 %>%
  filter(family_id == 112)

# id 113
p_w2_raw %>%
  filter(id == 113)

p_w1_clean_6 %>%
  filter(family_id == 113)

# id 308
p_w2_raw %>%
  filter(id == 308)

p_w1_clean_6 %>%
  filter(family_id == 308)
```

# Checking duplicated ids
```{r}
data.frame(table(p_w2_raw$id)) # 26 ids have 2 or more instances each (ids 415 and 418 have 3 instances each)

dupes <- p_w2_raw %>%
  select(id, response_id) %>%
  arrange(id)
```

# notes on w2

id 108 is duplicated, appears as 2 moms. In w1 there is just one 108 that is a mom.

ids: 112 after q145 empty. Is this the same person that started twice or two different parents?  The pattern of responses is identical. Or was it duplicated for some reason? In W1 it appears as if there were a mom and a dad. This means that the dad dropped from w1 to w2? ---> which row to delete?

id 113: no responses at all: qualtrics var progress: 3%. Participant dropped? In w1 there are mom and dad. ---> delete row?

ids 308: 4 instances = 2 false starts, 2 completed, but in attempt 3 after q_12_2 is filled with -99 throughout. Is this one parent or two parents? 	R_UMEbJn4yl8Wc0wx is the one with completed data. This id corresponds to just 1 mom in w1. 

ids 314 is duplicated, appears as 2 moms. In w1 there is just one 314 that is a mom.

id 415 is a 3 parent family. 

ids 601: is triplicated: 2 moms, 1 dad. In w1 there is 1 mom, 1 dad. 

id 201, 204, 206, 207, 216, 302, 310, 408, 414, 418, 501, 503, 505, 506, 607, 608, 610, 613, 614, 615, : two parent fams in w1 and w2.

new variable: 217_4_text? 

-------------------------------------------------------------------------------------------------------------------------------
--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------

################## WAVE 3 ##################

# Loading the data & initial cleaning
```{r}
# wave 3 was separated in 2 different datasets.

p_w3_raw_a <- read_sav(here("nopublish", "7bb. Juntos Parent W3-SPAN_February 10, 2021_10.26.sav"))

p_w3_raw_a <- p_w3_raw_a %>% 
  arrange(ID_1_TEXT) %>% # ordering participants ids
  janitor::clean_names() %>% # cleaning names
  select(-id) %>% 
  rename(c("id" = "id_1_text"))

p_w3_raw_b <- read_sav(here("nopublish", "7bb. Juntos Parent W3-SPAN_February 10, 2021_10.27.sav"))

p_w3_raw_b <- p_w3_raw_b %>% 
  arrange(ID_1_TEXT) %>% # ordering participants ids
  janitor::clean_names() %>% # cleaning names
  select(-id) %>% 
  rename(c("id" = "id_1_text"))

p_w3_raw <- full_join(p_w3_raw_a, p_w3_raw_b) %>% 
  arrange(id)

```
```{r}
p_w3_raw %>%
  filter(finished == 0) # 310 completely empty (no other with complete) & 613 completely empty with 2 other completes 
```

