---
title: "parent_all_waves"
author: "Alejandra Garcia Isaza"
date: "5/18/2022"
output: html_document
---

Note on May 18, 2022: I copied the code below from "parent_wave_1.Rmd" to provide CEQP with fully cleaned all waves datasets.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rio)
library(here)
library(tidyverse)
library(haven)
library(janitor)
library(knitr)
library(surveytoolbox)
library(sjPlot)
library(kableExtra)
library(psych)
library(sjmisc)
```

################## WAVE 1 ##################

# Loading the data & initial cleaning
```{r}
p_w1_raw <- read_sav(here("nopublish", "7b. Juntos Parent W1-SPAN_February 10, 2021_10.24.sav"))

p_w1_raw <- p_w1_raw %>% 
  arrange(ID_1_TEXT) %>% # ordering participants ids
  janitor::clean_names() %>% # cleaning names
  select(-id) %>% 
  rename(c("id" = "id_1_text"))
```

# Checking duplicated ids
```{r}
data.frame(table(p_w1_raw$id)) # 29 ids have 2 or more instances each (ids 415 and 418 have 3 instances each)

dupes <- p_w1_raw %>%
  select(id, response_id) %>%
  arrange(id)
```

# Further cleaning & dealing with duplicates
```{r}
p_w1_clean <- p_w1_raw %>% 
  select(-1:-8, -10:-17, -19, -20) %>% # selecting out columns with metadata
  filter(id != "999") %>% # deleting 2 test rows
  filter(response_id != "R_31L6rXjsdcuBaz0") %>% # deleting 1 duped row "false start" (id 312)
  mutate(id = case_when(response_id == "R_12ch5JnOHKHRgre" ~ "408", # id 418 (top one) recoded as 408
                        response_id == "R_uaijWwHdl8sIOmB" ~ "317", # had a pj before id #
                        TRUE ~ as.character(id))) %>%
  arrange(id)

data.frame(table(p_w1_clean$id)) # duplicated ids match two-parent families, id 415 is a family of 3
```

# Fixing wrong condition assignment and school id
```{r}
# fixing wrong assignment
p_w1_clean_2 <- p_w1_clean %>%
  mutate(school_id = str_sub(id, 1, 1),
         condition = case_when(
           school_id == "1" | school_id == "3" | school_id == "5" ~ "1",
           school_id == "2" | school_id == "4" | school_id == "6" ~ "2"
         )) %>%
  select(school_id, condition, everything()) %>%
  select(- school) # selecting out (i.e. deleting) original faulty school variable
```

# Adding value labels and making vars numeric 
```{r}
p_w1_clean_3 <- p_w1_clean_2 %>%
  mutate(condition = as.numeric(condition),
         condition = set_vall(condition, c("control" = 1, "intervention" = 2)),
         school_id = as.numeric(school_id),
         school_id = set_vall(school_id, c("cascade" = 1, "prairie_mountain" = 2, "kelly" = 3, "ata" = 4, "briggs" = 5, "agnes_stewart" = 6)))
```

# Creating individual id variable
```{r}
p_w1_clean_3 <- p_w1_clean_3 %>% 
  mutate(participant = ifelse(response_id == "R_1ON5POXTzIXrIeb", 4, participant), # changing partic. type to correspond to records
         participant_4_text = ifelse(response_id == "R_1ON5POXTzIXrIeb", "stepmom", participant_4_text), 
         participant = as.numeric(participant),
         participant = set_vall(participant, c("mom" = 1, "dad" = 2, "tutor legal" = 3, "other, specify" = 4)),
         participant_id = paste0(id, "_", participant), .before = 5) %>% # new var as unique identifier
  rename(c("family_id" = "id")) 
```

# renaming vars to correspond to value labels 
```{r}
# item that asks about # of adults in home is spreaded out and had an unusual coding.
p_w1_clean_3 <- p_w1_clean_3 %>%
  rename(c("q211_1" = "q211_2"),
         c("q211_2" = "q211_99"),
         c("q211_3" = "q211_4"),
         c("q211_4" = "q211_5"),
         c("q211_5" = "q211_6"),
         c("q211_99" = "q211_7"),
         c("q211_6_text" = "q211_6_text"))
```

```{r}
view_df(p_w1_clean_3)
```

# Number of participants in Wave 1
```{r}
# 125 total participants in w1, from which...

p_w1_clean_3 %>%
  filter(participant == 1) # 91 were moms

p_w1_clean_3 %>%
  filter(participant == 2) # 32 were dads

p_w1_clean_3 %>%
  filter(participant == 4) # 2 were "other" (1 stepmom, 1 grandparent)

unique(p_w1_clean_3$family_id) # 96 unique id values, thus 96 families in W1. 
```

# complete parent data to join with other waves 
```{r}
p_w1_clean_4 <- p_w1_clean_3 %>%
  filter(family_id != 301) %>% # deleted bec participant dropped from study and is -99 throughout. Id not present in w2 and w3.
  arrange(participant_id)
```

# Number of participants in Wave 1
```{r}
# 124 total participants in w1, from which...

p_w1_clean_4 %>%
  filter(participant == 1) # 91 were moms

p_w1_clean_4 %>%
  filter(participant == 2) # 31 were dads

p_w1_clean_4 %>%
  filter(participant == 4) # 2 were "other" (1 stepmom, 1 grandparent)

unique(p_w1_clean_4$family_id) # 95 unique id values, thus 94 families in W1. 
```



################## WAVE 2 ##################

# Loading the data & initial cleaning
```{r}
p_w2_raw <- read_sav(here("nopublish", "7bb. Juntos Parent W2-SPAN_February 10, 2021_10.25.sav"))

p_w2_raw <- p_w2_raw %>% 
  arrange(ID_1_TEXT) %>% # ordering participants ids
  janitor::clean_names() %>% # cleaning names
  select(-id) %>% 
  rename(c("id" = "id_1_text"))
```

# Checking incomplete rows & duplicated ids
```{r}
p_w2_raw %>%
  filter(finished == 0)

p_w2_raw %>%
  filter(finished == 1)

data.frame(table(p_w2_raw$id)) # 26 ids have 2 or more instances each (ids 415 and 601 have 3 instances each, 308 has 4 instances)
```

# Further cleaning & dealing with duplicates
```{r}
p_w2_clean <- p_w2_raw %>% 
  select(-1:-8, -10:-17, -19, -20) %>% # selecting out columns with metadata
  filter(response_id != "R_2sdrLRMKZvR9fwO", # deleting duped row "false start" (id 112) - this family dropped amid w2
         response_id != "R_2zoPGHCXUdvsDgd", # deleting duped row "false start" (id 113)
         response_id != "R_1EQFk2EbJW61nhf", # deleting duped row "false start" (id 308)
         response_id != "R_XMroI3biv8jxEVb", # deleting duped row "false start" (id 308)
         response_id != "R_3PZzUFRxLeeW6GJ") %>% # deleting duped row "false start" (id 308) 
  mutate(id = case_when(response_id == "R_1rCXE5pBSje0twj" ~ "109", # id 108 (top one) recoded as 109
                        response_id == "R_2EpkLxRpA7GW0oE" ~ "312", # id 314 (bottom one) recoded as 312
                        TRUE ~ as.character(id))) %>%
  arrange(id)

data.frame(table(p_w2_clean$id)) # duplicated ids match two-parent families, id 415 is a family of 3. Get confirmation about 109 and 601, they are 2-parent fams in w1, though. 
```

# ------ Notes on W2 dataset -------
# duplicated ids match two-parent families, id 415 is a family of 3. Get confirmation about 109 and 601, they are 2-parent fams in w1, though. 

# ------------------------------------

```{r}


p_w1_clean %>%
  filter(id == 601)

p_w2_clean %>%
  filter(id == 601)# says id 601 from wave 3, but participants with id 601 in wave 3 dataset are verified 

p_w1_raw %>%
  filter(id == 601) 

p_w2_raw %>%
  filter(id == 601) 

# response id that appears as belonging to id 601-mom in logs from wave 3
p_w1_raw %>%
  filter(response_id == "R_2QRPBKfzZwFDAaT") # not in wave 1

p_w2_raw %>%
  filter(response_id == "R_2QRPBKfzZwFDAaT") # not in wave 2

# and is also not present in wave 3 ---> no clue where this came out from
```

Notes with Eli
In wave 1:
R_UG9FQ6VvanpWwwN	id 601	dad, wave 1
R_1Lwe3h7XQeF5hyj	id 601  mom, wave 1

In wave 2: 
R_2sZpDvTZRNJfl7Z	id 601	mom, wave 2 = real mom
R_O6DQfVORSAtnWbT	id 601	dad, wave 2 = real dad

R_vNctgxUZcfcMwFj	id 601  mom, wave 3 = leaked from w3

In wave 3: 
R_3LhFdFrq0SFkkKI	601	dad, wave 3	
R_20lyHTLoqvQ4Rpz	601 mom, wave 3

R_vNctgxUZcfcMwFj	id 601  mom, wave 3
R_20lyHTLoqvQ4Rpz	601 mom, wave 3 
these 2 rows are very similar, but not identical = Eli & I think that it is the same data, but assessor error. 
#####


in dataset (w1)
R_UG9FQ6VvanpWwwN	601 dad wave1
R_1Lwe3h7XQeF5hyj	601 mom wave1

in dataset (w2)
R_2sZpDvTZRNJfl7Z	601	mom wave2
R_O6DQfVORSAtnWbT	601 dad wave2
R_vNctgxUZcfcMwFj	601 mom wave3

in dataset (w3)
R_20lyHTLoqvQ4Rpz 601 mom wave3 
R_3LhFdFrq0SFkkKI 601 dad wave3


in logs (w1)
R_1Lwe3h7XQeF5hyj 601 mom wave1
R_UG9FQ6VvanpWwwN 601 dad wave1

in logs (w2)
R_2sZpDvTZRNJfl7Z 601 mom wave2
R_O6DQfVORSAtnWbT 601 dad wave2

in logs (w3)
R_2QRPBKfzZwFDAaT 601 mom wave3
R_3LhFdFrq0SFkkKI 601 dad wave3

now I don't know who is the true id 601 mom:
R_20lyHTLoqvQ4Rpz (appears in dataset)
R_2QRPBKfzZwFDAaT (appears in logs but anywhere in datasets)

############ HERE ########

# Fixing wrong condition assignment and school id
```{r}
# fixing wrong assignment
p_w2_clean_2 <- p_w2_clean %>%
  mutate(school_id = str_sub(id, 1, 1),
         condition = case_when(
           school_id == "1" | school_id == "3" | school_id == "5" ~ "1",
           school_id == "2" | school_id == "4" | school_id == "6" ~ "2"
         )) %>%
  select(school_id, condition, everything()) %>%
  select(- school) # selecting out (i.e. deleting) original faulty school variable
```

# Adding value labels and making vars numeric 
```{r}
p_w2_clean_3 <- p_w2_clean_2 %>%
  mutate(condition = as.numeric(condition),
         condition = set_vall(condition, c("control" = 1, "intervention" = 2)),
         school_id = as.numeric(school_id),
         school_id = set_vall(school_id, c("cascade" = 1, "prairie_mountain" = 2, "kelly" = 3, "ata" = 4, "briggs" = 5, "agnes_stewart" = 6)))
```

# Creating individual id variable
```{r}
p_w2_clean_3 <- p_w2_clean_3 %>% 
  mutate(participant = ifelse(response_id == "R_1PRjrvEzZRgtKKD", 4, participant), # changing partic. type to correspond to records
         participant_4_text = ifelse(response_id == "R_1PRjrvEzZRgtKKD", "stepmom", participant_4_text), 
         participant = as.numeric(participant),
         participant = set_vall(participant, c("mom" = 1, "dad" = 2, "tutor legal" = 3, "other, specify" = 4)),
         participant_id = paste0(id, "_", participant), .before = 5) %>% # new var as unique identifier
  rename(c("family_id" = "id")) # 108 participants, 82 moms, 25 dads, 1 stepmom. 84 families
```

# renaming vars to correspond to value labels and collapsing spreaded out variable
```{r}
# item that asks about # of adults in home was spreaded out and had an unusual coding.
p_w2_clean_3 <- p_w2_clean_3 %>%
  rename(c("q211_1" = "q211_2"),
         c("q211_2" = "q211_99"),
         c("q211_3" = "q211_4"),
         c("q211_4" = "q211_5"),
         c("q211_5" = "q211_6"),
         c("q211_99" = "q211_7"),
         c("adults_home_text" = "q211_6_text"))

# collapsing variable
num_adults_w2 <- p_w2_clean_3 %>% 
  select(participant_id, starts_with("q211_"), -adults_home_text) %>%
  pivot_longer(
    cols = starts_with("q211"),
    names_to = c("item", "adults_home"),
    names_sep = "_",
    values_to = "response",
    values_drop_na = TRUE) %>% 
  filter(response == 1) %>%
  mutate(response = as.numeric(response))

get_dupes(num_adults_w2, participant_id)

# id 502_1 chose options 2 and 4

# Option 2, "2" will be chosen because the {distinct} function keeps the first row. Also it can be argued that there are 2 adults in each household and not 4 and 5, respectively. 

num_adults_w2 <- num_adults_w2 %>% 
  distinct(participant_id, .keep_all = TRUE) %>% 
  select(-item, -response) 

# not completely sure why it goes down to 124 in the num_adults_w2? 
```

```{r}
view_df(p_w2_clean_3)
```


