---
title: "parent_3_waves"
author: "Alejandra Garcia Isaza"
date: "6/30/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rio)
library(here)
library(tidyverse)
library(haven)
library(janitor)
library(knitr)
library(surveytoolbox)
library(sjPlot)
library(kableExtra)
library(psych)
library(sjmisc)
```

################## WAVE 1 ##################

# Loading the data & initial cleaning
```{r}
p_w1_raw <- read_sav(here("nopublish", "7b. Juntos Parent W1-SPAN_February 10, 2021_10.24.sav"))

p_w1_raw <- p_w1_raw %>% 
  arrange(ID_1_TEXT) %>% # ordering participants ids
  janitor::clean_names() %>% # cleaning names
  select(-id) %>% 
  rename(c("id" = "id_1_text"))
```

# Checking duplicated ids
```{r}
data.frame(table(p_w1_raw$id)) # 29 ids have 2 or more instances each (ids 415 and 418 have 3 instances each)

dupes <- p_w1_raw %>%
  select(id, response_id) %>%
  arrange(id)
```

# Further cleaning & dealing with duplicates
```{r}
p_w1_clean <- p_w1_raw %>% 
  select(-1:-8, -10:-17, -19, -20) %>% # selecting out columns with metadata
  filter(id != "999") %>% # deleting 2 test rows
  filter(response_id != "R_31L6rXjsdcuBaz0") %>% # deleting 1 duped row "false start" (id 312)
  mutate(id = case_when(response_id == "R_12ch5JnOHKHRgre" ~ "408", # id 418 (top one) recoded as 408
                        response_id == "R_uaijWwHdl8sIOmB" ~ "317", # had a pj before id #
                        TRUE ~ as.character(id))) %>%
  arrange(id)

data.frame(table(p_w1_clean$id)) # duplicated ids match two-parent families, id 415 is a family of 3
```

# Fixing wrong condition assignment and school id
```{r}
# fixing wrong assignment
p_w1_clean_2 <- p_w1_clean %>%
  mutate(school_id = str_sub(id, 1, 1),
         condition = case_when(
           school_id == "1" | school_id == "3" | school_id == "5" ~ "1",
           school_id == "2" | school_id == "4" | school_id == "6" ~ "2"
         )) %>%
  select(school_id, condition, everything()) %>%
  select(- school) # selecting out (i.e. deleting) original faulty school variable
```

# Adding value labels and making vars numeric 
```{r}
p_w1_clean_3 <- p_w1_clean_2 %>%
  mutate(condition = as.numeric(condition),
         condition = set_vall(condition, c("control" = 1, "intervention" = 2)),
         school_id = as.numeric(school_id),
         school_id = set_vall(school_id, c("cascade" = 1, "prairie_mountain" = 2, "kelly" = 3, "ata" = 4, "briggs" = 5, "agnes_stewart" = 6)))
```

# Creating individual id variable
```{r}
p_w1_clean_3 <- p_w1_clean_3 %>% 
  mutate(participant = ifelse(response_id == "R_1ON5POXTzIXrIeb", 4, participant), # changing partic. type to correspond to records
         participant_4_text = ifelse(response_id == "R_1ON5POXTzIXrIeb", "stepmom", participant_4_text), 
         participant = as.numeric(participant),
         participant = set_vall(participant, c("mom" = 1, "dad" = 2, "tutor legal" = 3, "other, specify" = 4)),
         participant_id = paste0(id, "_", participant), .before = 5) %>% # new var as unique identifier
  rename(c("family_id" = "id")) 

view_df(p_w1_clean_3)
```

# renaming vars to correspond to value labels and collapsing spreaded out variable
```{r}
# item that asks about # of adults in home was spreaded out and had an unusual coding.
p_w1_clean_3 <- p_w1_clean_3 %>%
  rename(c("q211_1" = "q211_2"),
         c("q211_2" = "q211_99"),
         c("q211_3" = "q211_4"),
         c("q211_4" = "q211_5"),
         c("q211_5" = "q211_6"),
         c("q211_99" = "q211_7"),
         c("adults_home_text" = "q211_6_text"))

# collapsing variable
num_adults <- p_w1_clean_3 %>% 
  select(participant_id, starts_with("q211_"), -adults_home_text) %>%
  pivot_longer(
    cols = starts_with("q211"),
    names_to = c("item", "adults_home"),
    names_sep = "_",
    values_to = "response",
    values_drop_na = TRUE) %>% 
  filter(response == 1) %>%
  mutate(response = as.numeric(response))

get_dupes(num_adults, participant_id)

# id 114_1 chose options 2 and 4
# id 214_1 chose options 2 and 5

# Option 2, "2" will be chosen because the {distinct} function keeps the first row. Also it can be argued that there are 2 adults in each household and not 4 and 5, respectively. 

num_adults <- num_adults %>% 
  distinct(participant_id, .keep_all = TRUE) %>% 
  select(-item, -response) 

p_w1_clean_4 <- left_join(p_w1_clean_3, num_adults) %>%
  select(-291:-296) %>%
  mutate(adults_home = as.numeric(adults_home),
         adults_home = set_vall(adults_home, c("1" = 1, "2" = 2, "3" = 3, "4" = 4, "mas de 5, ¿cuántas?" = 5, "refuse to answer" = 99)),
         adults_home = set_varl(adults_home, "Número de adultos (más de 18 años): - Selected Choice")) %>%
  move_columns(adults_home, .before = "adults_home_text")
```

# renaming demo vars
```{r}
p_w1_clean_5 <- p_w1_clean_4 %>%
  rename(c("lang_english" = "q170_1"),
         c("lang_spanish" = "q170_2"),
         c("lang_indigenous" = "q170_3"),
         c("lang_other" = "q170_4"),
         c("lang_refuse_answer" = "q170_99"),
         c("lang_indigenous_text" = "q170_3_text"),
         c("lang_other_text" = "q170_4_text"),
         c("youth_lang_broker" = "q171"),
         c("spanish_comfort_home" = "q172_1"),
         c("spanish_comfort_school" = "q172_2"),
         c("spanish_comfort_youth_school" = "q172_3"),
         c("spanish_comfort_work" = "q172_4"),
         c("spanish_comfort_friends" = "q172_5"),
         c("spanish_comfort_general" = "q172_6"),
         c("english_comfort_home" = "q173_1"),
         c("english_comfort_school" = "q173_2"),
         c("english_comfort_youth_school" = "q173_3"),
         c("english_comfort_work" = "q173_4"),
         c("english_comfort_friends" = "q173_5"),
         c("english_comfort_general" = "q173_6"),
         c("minors_home" = "q211a"),
         c("minors_home_text" = "q211a_6_text"),
         c("pub_assist" = "q212"),
         c("pub_assist_snap" = "q212a_1"),
         c("pub_assist_frpl" = "q212a_2"),
         c("pub_assist_tanf" = "q212a_3"),
         c("pub_assist_unemployment" = "q212a_4"),
         c("pub_assist_disability" = "q212a_5"),
         c("pub_assist_social_sec" = "q212a_6"),
         c("pub_assist_health_insu" = "q212a_7"),
         c("pub_assist_wic" = "q212a_18"),
         c("pub_assist_food_box" = "q212a_19"),
         c("pub_assist_other" = "q212a_10"),
         c("pub_assist_refuse_answer" = "q212a_99")) %>%
  select(-response_id) # response_id no longer needed
```

# Number of participants in Wave 1
```{r}
# 125 total participants in w1, from which...

p_w1_clean_5 %>%
  filter(participant == 1) # 91 were moms

p_w1_clean_5 %>%
  filter(participant == 2) # 32 were dads

p_w1_clean_5 %>%
  filter(participant == 4) # 2 were "other" (1 stepmom, 1 grandparent)

unique(p_w1_clean_5$family_id) # 96 unique id values, thus 96 families in W1. 
```

# ------ Notes on W1 dataset -------
# participants ids 415 and 418 had 3 records each. 
# participant id 415 was a family of three: 2 moms, 1 dad. One of the moms was recoded as "other, specify" in the `participant` variable and "stepmom" in the `participant_4_text` variable. 
# Participant id 418 was "triplicated" and there was no "mom" for id 408 that was only a "dad". This was fixed.
# # Total participants in wave n = 125
# ------------------------------------

```{r}
view_df(p_w1_clean_5) # it appears that there are no questions about age, gender id, race/ethnicity. 
```

```{r}
# adding "p1" to each variable name: "p" stands for parent, "1" for wave 1. 
# names(p_w1_clean_5) <- paste(names(p_w1_clean_5), "p1", sep = "_")
# 
# p_w1_clean_6 <- p_w1_clean_5 %>%
#   rename(c("school_id" = "school_id_p1"), 
#          c("condition" = "condition_p1"),
#          c("family_id" = "family_id_p1"),
#          c("participant_id" = "participant_id_p1")) # renaming these variables to be able to join later
# 
# view_df(p_w1_clean_6)
```

################## WAVE 2 ##################

# Loading the data & initial cleaning
```{r}
p_w2_raw <- read_sav(here("nopublish", "7bb. Juntos Parent W2-SPAN_February 10, 2021_10.25.sav"))

p_w2_raw <- p_w2_raw %>% 
  arrange(ID_1_TEXT) %>% # ordering participants ids
  janitor::clean_names() %>% # cleaning names
  select(-id) %>% 
  rename(c("id" = "id_1_text"))
```

# Checking incomplete rows & duplicated ids
```{r}
p_w2_raw %>%
  filter(finished == 0)

p_w2_raw %>%
  filter(finished == 1)

data.frame(table(p_w2_raw$id)) # 26 ids have 2 or more instances each (ids 415 and 418 have 3 instances each)
```

# Further cleaning & dealing with duplicates
```{r}
p_w2_clean <- p_w2_raw %>% 
  select(-1:-8, -10:-17, -19, -20) %>% # selecting out columns with metadata
  filter(response_id != "R_2sdrLRMKZvR9fwO", # deleting duped row "false start" (id 112) - this family dropped amid w2
         response_id != "R_2zoPGHCXUdvsDgd", # deleting duped row "false start" (id 113)
         response_id != "R_1EQFk2EbJW61nhf", # deleting duped row "false start" (id 308)
         response_id != "R_XMroI3biv8jxEVb", # deleting duped row "false start" (id 308)
         response_id != "R_3PZzUFRxLeeW6GJ") %>% # deleting duped row "false start" (id 308) 
  mutate(id = case_when(response_id == "R_1rCXE5pBSje0twj" ~ "109", # id 108 (top one) recoded as 109
                        response_id == "R_2EpkLxRpA7GW0oE" ~ "312", # id 314 (bottom one) recoded as 312
                        TRUE ~ as.character(id))) %>%
  arrange(id)

data.frame(table(p_w2_clean$id)) # duplicated ids match two-parent families, id 415 is a family of 3. Get confirmation about 109 and 601, they are 2-parent fams in w1, though. 
```

############ HERE ########

# Fixing wrong condition assignment and school id
```{r}
# fixing wrong assignment
p_w2_clean_2 <- p_w2_clean %>%
  mutate(school_id = str_sub(id, 1, 1),
         condition = case_when(
           school_id == "1" | school_id == "3" | school_id == "5" ~ "1",
           school_id == "2" | school_id == "4" | school_id == "6" ~ "2"
         )) %>%
  select(school_id, condition, everything()) %>%
  select(- school) # selecting out (i.e. deleting) original faulty school variable
```

# Adding value labels and making vars numeric 
```{r}
p_w2_clean_3 <- p_w2_clean_2 %>%
  mutate(condition = as.numeric(condition),
         condition = set_vall(condition, c("control" = 1, "intervention" = 2)),
         school_id = as.numeric(school_id),
         school_id = set_vall(school_id, c("cascade" = 1, "prairie_mountain" = 2, "kelly" = 3, "ata" = 4, "briggs" = 5, "agnes_stewart" = 6)))
```

# Creating individual id variable
```{r}
p_w2_clean_3 <- p_w2_clean_3 %>% 
  mutate(participant = ifelse(response_id == "R_1PRjrvEzZRgtKKD", 4, participant), # changing partic. type to correspond to records
         participant_4_text = ifelse(response_id == "R_1PRjrvEzZRgtKKD", "stepmom", participant_4_text), 
         participant = as.numeric(participant),
         participant = set_vall(participant, c("mom" = 1, "dad" = 2, "tutor legal" = 3, "other, specify" = 4)),
         participant_id = paste0(id, "_", participant), .before = 5) %>% # new var as unique identifier
  rename(c("family_id" = "id")) # 108 participants, 82 moms, 25 dads, 1 stepmom. 84 families
```

# renaming vars to correspond to value labels and collapsing spreaded out variable
```{r}
# item that asks about # of adults in home was spreaded out and had an unusual coding.
p_w2_clean_3 <- p_w2_clean_3 %>%
  rename(c("q211_1" = "q211_2"),
         c("q211_2" = "q211_99"),
         c("q211_3" = "q211_4"),
         c("q211_4" = "q211_5"),
         c("q211_5" = "q211_6"),
         c("q211_99" = "q211_7"),
         c("adults_home_text" = "q211_6_text"))

# collapsing variable
num_adults_w2 <- p_w2_clean_3 %>% 
  select(participant_id, starts_with("q211_"), -adults_home_text) %>%
  pivot_longer(
    cols = starts_with("q211"),
    names_to = c("item", "adults_home"),
    names_sep = "_",
    values_to = "response",
    values_drop_na = TRUE) %>% 
  filter(response == 1) %>%
  mutate(response = as.numeric(response))

get_dupes(num_adults_w2, participant_id)

# id 502_1 chose options 2 and 4

# Option 2, "2" will be chosen because the {distinct} function keeps the first row. Also it can be argued that there are 2 adults in each household and not 4 and 5, respectively. 

num_adults_w2 <- num_adults_w2 %>% 
  distinct(participant_id, .keep_all = TRUE) %>% 
  select(-item, -response) 

# not completely sure why it goes down to 124 in the num_adults_w2? 
```

```{r}
view_df(p_w2_clean_3)
```

# notes on w2

# wave correct?
```{r}
# response_id != "R_vNctgxUZcfcMwFj") %>% # deleting row bec. it belongs to w3 (id 601) & there is already a mom and dad 
  
p_w1_clean %>%
  filter(id == 601)

p_w2_clean %>%
  filter(id == 601)
```
```{r}
p_w3_raw %>%
  select(-1:-8, -10:-17, -19, -20) %>%
  filter(id == 101)
```


In wave 1:
R_UG9FQ6VvanpWwwN	id 601	dad, wave 1
R_1Lwe3h7XQeF5hyj	id 601  mom, wave 1

In wave 2: 
R_2sZpDvTZRNJfl7Z	id 601	mom, wave 2 = real mom
R_O6DQfVORSAtnWbT	id 601	dad, wave 2 = real dad

R_vNctgxUZcfcMwFj	id 601  mom, wave 3 = leaked from w3

In wave 3: 
R_3LhFdFrq0SFkkKI	601	dad, wave 3	
R_20lyHTLoqvQ4Rpz	601 mom, wave 3

R_vNctgxUZcfcMwFj	id 601  mom, wave 3
R_20lyHTLoqvQ4Rpz	601 mom, wave 3 
these 2 rows are very similar, but not identical = Eli & I think that it is the same data, but assessor error. 

# wave correct?
R_2qyriEL0QoD8Kln	id 604 in wave 1, says wave = 1

R_6FiHWtagxFiXgbf	id 604 in wave 2, says wave = 1. CHANGE to wave 2, assesor mistakenly put wave 1 and should have put wave 2.  

# new variable
new variable: 217_4_text? The question in this variable was completely changed by w2, levels also changed (no 99 and added a 4 for yearly money). 

----------------------------------------------------------------------------------------------------

################## WAVE 3 ##################

# Loading the data & initial cleaning
```{r}
# wave 3 was separated in 2 different datasets.

p_w3_raw_a <- read_sav(here("nopublish", "7bb. Juntos Parent W3-SPAN_February 10, 2021_10.26.sav"))

p_w3_raw_a <- p_w3_raw_a %>% 
  arrange(ID_1_TEXT) %>% # ordering participants ids
  janitor::clean_names() %>% # cleaning names
  select(-id) %>% 
  rename(c("id" = "id_1_text"))

p_w3_raw_b <- read_sav(here("nopublish", "7bb. Juntos Parent W3-SPAN_February 10, 2021_10.27.sav"))

p_w3_raw_b <- p_w3_raw_b %>% 
  arrange(ID_1_TEXT) %>% # ordering participants ids
  janitor::clean_names() %>% # cleaning names
  select(-id) %>% 
  rename(c("id" = "id_1_text"))

p_w3_raw <- full_join(p_w3_raw_a, p_w3_raw_b) %>% 
  arrange(id)
```

```{r}
p_w3_raw %>%
  filter(finished == 0) # 310 completely empty (no other with complete) & 613 completely empty with 2 other completes 
```

# Checking duplicated ids
```{r}
data.frame(table(p_w3_raw$id)) # 24 ids have 2 or more instances each (ids 408, 415, 613 have 3 instances each)

dupes <- p_w3_raw %>%
  select(id, response_id) %>%
  arrange(id)
```

```{r}
# id 308
p_w3_raw %>%
  filter(id == 601)
```

# notes on w3

id 107 is duplicated, appears as 2 moms. In w1 and w2 there is just one 107 that is a mom.

ids 109, 110 duplicated: both have a mom and dad, also on w1. But there is no 109 mom and any 110 parent in w2. 

id 113 duplicated: mom & dad. No dad on w2. mom & dad on w1. 

id 207 only a mom, no dad on w3. But mom & dad in w1 and w2. Dropped?

id 310 only a dad, no mom on w3. But mom & dad in w1 and w2. Dropped?

id 405 duplicated. It says "imported". Mom & dad in w1, no 405 in w2. 

id 408 triplicated: 2 dads 1 mom. One of the dads says "imported". In w1 & w2 there is mom & dad. 

id 415 three-parent family. Second mom properly coded as other and stepmom. 

id 503+ duplicated: mom & dad. doesn't say "imported". I think this one is "new" data from page 2 of w3. 

id 605 duplicated. It says "imported". In w1 and w2 there is just one 605 that is a mom.

id 610+ duplicated: mom & dad. doesn't say "imported". I think this one is "new" data from page 2 of w3. 

id 613+ triplicated: 1 mom & 2 dads. In w1 & w2 there is mom & dad.

id 614 is missing the mom. In w1 & w2 there is mom & dad. Dropped?


ids 201, 204, 206, 216, 302, 414, 418, 501, 503+, 506, 601(tripli in w3), 607, 608, 610+, 613+, 615: two parent fams in w1, w2 and w3.

