---
title: "parent_wave_2"
author: "Alejandra Garcia Isaza"
date: '2022-08-26'
output: html_document
---

Note on August 26, 2022: I copied the code below from "parent_3_waves.Rmd" which I think is the doc I was working on the latest for cleaning. Here I will only work on parents wave 2 (post-intervention).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rio)
library(here)
library(tidyverse)
library(haven)
library(janitor)
library(knitr)
library(surveytoolbox)
library(sjPlot)
library(kableExtra)
library(psych)
library(sjmisc)
```

################## WAVE 2 ##################

# Loading the data & initial cleaning
```{r}
p_w2_raw <- read_sav(here("nopublish", "7bb. Juntos Parent W2-SPAN_February 10, 2021_10.25.sav"))

p_w2_raw <- p_w2_raw %>% 
  arrange(ID_1_TEXT) %>% # ordering participants ids
  janitor::clean_names() %>% # cleaning names
  select(-id) %>% 
  rename(c("id" = "id_1_text"))
```

# Checking incomplete rows & duplicated ids
```{r}
p_w2_raw %>%
  filter(finished == 0)

p_w2_raw %>%
  filter(finished == 1)

data.frame(table(p_w2_raw$id)) # 26 ids have 2 or more instances each (ids 415 and 418 have 3 instances each)
```

# Further cleaning & dealing with duplicates
```{r}
p_w2_clean <- p_w2_raw %>% 
  select(-1:-8, -10:-17, -19, -20) %>% # selecting out columns with metadata
  filter(response_id != "R_2sdrLRMKZvR9fwO", # deleting duped row "false start" (id 112) - this family dropped amid w2
         response_id != "R_2zoPGHCXUdvsDgd", # deleting duped row "false start" (id 113)
         response_id != "R_1EQFk2EbJW61nhf", # deleting duped row "false start" (id 308)
         response_id != "R_XMroI3biv8jxEVb", # deleting duped row "false start" (id 308)
         response_id != "R_3PZzUFRxLeeW6GJ") %>% # deleting duped row "false start" (id 308) 
  mutate(id = case_when(response_id == "R_1rCXE5pBSje0twj" ~ "109", # id 108 (top one) recoded as 109
                        response_id == "R_2EpkLxRpA7GW0oE" ~ "312", # id 314 (bottom one) recoded as 312
                        TRUE ~ as.character(id))) %>%
  arrange(id)

data.frame(table(p_w2_clean$id)) # duplicated ids match two-parent families, id 415 is a family of 3. Get confirmation about 109 and 601, they are 2-parent fams in w1, though. 
```

# Fixing wrong condition assignment and school id
```{r}
# fixing wrong assignment
p_w2_clean_2 <- p_w2_clean %>%
  mutate(school_id = str_sub(id, 1, 1),
         condition = case_when(
           school_id == "1" | school_id == "3" | school_id == "5" ~ "1",
           school_id == "2" | school_id == "4" | school_id == "6" ~ "2"
         )) %>%
  select(school_id, condition, everything()) %>%
  select(- school) # selecting out (i.e. deleting) original faulty school variable
```

# Adding value labels and making vars numeric 
```{r}
p_w2_clean_3 <- p_w2_clean_2 %>%
  mutate(condition = as.numeric(condition),
         condition = set_vall(condition, c("control" = 1, "intervention" = 2)),
         school_id = as.numeric(school_id),
         school_id = set_vall(school_id, c("cascade" = 1, "prairie_mountain" = 2, "kelly" = 3, "ata" = 4, "briggs" = 5, "agnes_stewart" = 6)))
```

# Creating individual id variable
```{r}
p_w2_clean_3 <- p_w2_clean_3 %>% 
  mutate(participant = ifelse(response_id == "R_1PRjrvEzZRgtKKD", 4, participant), # changing partic. type to correspond to records
         participant_4_text = ifelse(response_id == "R_1PRjrvEzZRgtKKD", "stepmom", participant_4_text), 
         participant = ifelse(response_id == "R_2SCy44y5QaDXggq", 4, participant), # changing partic. type to abuela to correspond to W1
         participant_4_text = ifelse(response_id == "R_2SCy44y5QaDXggq", "abuela", participant_4_text),
         participant = as.numeric(participant),
         participant = set_vall(participant, c("mom" = 1, "dad" = 2, "tutor legal" = 3, "other, specify" = 4)),
         participant_id = paste0(id, "_", participant), .before = 5) %>% # new var as unique identifier
  rename(c("family_id" = "id"))
```


# renaming vars to correspond to value labels
```{r}
# item that asks about # of adults in home was spreaded out and had an unusual coding.
p_w2_clean_3 <- p_w2_clean_3 %>%
  rename(c("q211_1" = "q211_2"),
         c("q211_2" = "q211_99"),
         c("q211_3" = "q211_4"),
         c("q211_4" = "q211_5"),
         c("q211_5" = "q211_6"),
         c("q211_99" = "q211_7"),
         c("adults_home_text" = "q211_6_text"))
```

```{r}
view_df(p_w2_clean_3)
```


# Number of participants in Wave 2
```{r}
# 109 total parent participants in w2, from which...

p_w2_clean_3 %>%
  filter(participant == 1) # 82 were moms

p_w2_clean_3 %>%
  filter(participant == 2) # 25 were dads

p_w2_clean_3 %>%
  filter(participant == 4) # 2 were "other" (1 stepmom, 1 abuela) 

unique(p_w2_clean_3$family_id) # 84 unique id values, thus 84 families in W2. 
```
```{r}
data.frame(table(p_w2_clean_3$family_id)) # 415 family of 3: checks-out; 601 triplicated, still figuring it out (08.26.22)
```


# notes on w2 meeting with Eli on 08.26.22
id 109: in dataset appears as a 2 parent (1 mom, 1 dad) -- Eli confirms that 109 is a family of 2, mome and dad, however the qualtrics code doesn't match from logs: (SV_0i9InOoo0RH6o9T) to what we have in dataset (R_z1iP1gjIGExS3Pr)

# In wave 1 (logs):
R_1Lwe3h7XQeF5hyj	id 601  mom, wave 1
R_UG9FQ6VvanpWwwN	id 601	dad, wave 1

# In wave 2 (logs): 
R_2sZpDvTZRNJfl7Z	id 601	mom, wave 2 = real mom
R_O6DQfVORSAtnWbT	id 601	dad, wave 2 = real dad 

R_vNctgxUZcfcMwFj	id 601  mom, wave 3 = the theory was that this row leaked from w3, but id 601 in W3 already has two id associated with 601. This still needs to be figured out. ## hence action plan is delete from wave 2 ##

# In wave 3 (logs): 
R_2QRPBKfzZwFDAaT id 601	mom, confirmed
R_3LhFdFrq0SFkkKI id 601	dad, confirmed

# From Eli (08.26.22)
R_2QRPBKfzZwFDAaT W3 601 Mom from email from JCG

# From notes I had before 08.26.22 
R_3LhFdFrq0SFkkKI	601	dad, wave 3	(in wave 3 log)
R_20lyHTLoqvQ4Rpz	601 mom, wave 3 (not sure where this came from)

R_vNctgxUZcfcMwFj	id 601  mom --> mistery what this response id belongs to
R_20lyHTLoqvQ4Rpz	601 mom, wave 3  --> mistery what this response id belongs to
these 2 rows are very similar, but not identical = Eli & I think that it is the same data, but assessor error (08.26.22: not sure where we were looking, these two ids do not appear on logs) 

# Deleting row R_vNctgxUZcfcMwFj	id 601
```{r}

p_w2_clean_3 %>%
  filter(response_id == "R_vNctgxUZcfcMwFj") # will be deleting from wave 2 as it does not appear anywhere in logs

p_w2_clean_3 %>%
  filter(response_id == "R_20lyHTLoqvQ4Rpz") # does not appear in dataset or in logs

p_w2_clean_3 <- p_w2_clean_3 %>%
  filter(response_id != "R_vNctgxUZcfcMwFj")
```

# Number of participants in Wave 2 (AFTER DELETION)
```{r}
# 108 total parent participants in w2, from which...

p_w2_clean_3 %>%
  filter(participant == 1) # 81 were moms

p_w2_clean_3 %>%
  filter(participant == 2) # 25 were dads

p_w2_clean_3 %>%
  filter(participant == 4) # 2 were "other" (1 stepmom, 1 abuela) 

unique(p_w2_clean_3$family_id) # 84 unique id values, thus 84 families in W2. 
```
tO DO: SIDE BY SIDE COMPARE WAVE 1 AND WAVE 2 TO SEE IF ID WITH 2 PARTICIPANTS COINCIDE WITH 2 PARENT-FAMILIES IN WAVE 1 AND WAVE 2 AND THEN EXPORT TO SPSS FILE

```{r}
# all_parent_wave2 %>%
#   haven::write_sav(here("nopublish", "all_parent_wave2.sav")) 
```




HERE


## previous notes:

# wave correct?
R_2qyriEL0QoD8Kln	id 604 in wave 1, says wave = 1

R_6FiHWtagxFiXgbf	id 604 in wave 2, says wave = 1. CHANGE to wave 2, assesor mistakenly put wave 1 and should have put wave 2.  

# new variable
new variable: 217_4_text? The question in this variable was completely changed by w2, levels also changed (no 99 and added a 4 for yearly money). 

----------------------------------------------------------------------------------------------------