---
title: "parent w1"
author: "Alejandra Garcia Isaza"
date: "3/31/2022"
output: html_document
---
Note on March 31, 2022: I copied the code below from "parent_3_waves.Rmd" which I think is the doc I was working on the latest for cleaning. Here I will only work on parents wave 1 (baseline) for dissertation and will update progress made here to the three waves Rmd to continue cleaning.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rio)
library(here)
library(tidyverse)
library(haven)
library(janitor)
library(knitr)
library(surveytoolbox)
library(sjPlot)
library(kableExtra)
library(psych)
library(sjmisc)
```

################## WAVE 1 ##################

# Loading the data & initial cleaning
```{r}
p_w1_raw <- read_sav(here("nopublish", "7b. Juntos Parent W1-SPAN_February 10, 2021_10.24.sav"))

p_w1_raw <- p_w1_raw %>% 
  arrange(ID_1_TEXT) %>% # ordering participants ids
  janitor::clean_names() %>% # cleaning names
  select(-id) %>% 
  rename(c("id" = "id_1_text"))
```

# Checking duplicated ids
```{r}
data.frame(table(p_w1_raw$id)) # 29 ids have 2 or more instances each (ids 415 and 418 have 3 instances each)

dupes <- p_w1_raw %>%
  select(id, response_id) %>%
  arrange(id)
```

# Further cleaning & dealing with duplicates
```{r}
p_w1_clean <- p_w1_raw %>% 
  select(-1:-8, -10:-17, -19, -20) %>% # selecting out columns with metadata
  filter(id != "999") %>% # deleting 2 test rows
  filter(response_id != "R_31L6rXjsdcuBaz0") %>% # deleting 1 duped row "false start" (id 312)
  mutate(id = case_when(response_id == "R_12ch5JnOHKHRgre" ~ "408", # id 418 (top one) recoded as 408
                        response_id == "R_uaijWwHdl8sIOmB" ~ "317", # had a pj before id #
                        TRUE ~ as.character(id))) %>%
  arrange(id)

data.frame(table(p_w1_clean$id)) # duplicated ids match two-parent families, id 415 is a family of 3
```

# Fixing wrong condition assignment and school id
```{r}
# fixing wrong assignment
p_w1_clean_2 <- p_w1_clean %>%
  mutate(school_id = str_sub(id, 1, 1),
         condition = case_when(
           school_id == "1" | school_id == "3" | school_id == "5" ~ "1",
           school_id == "2" | school_id == "4" | school_id == "6" ~ "2"
         )) %>%
  select(school_id, condition, everything()) %>%
  select(- school) # selecting out (i.e. deleting) original faulty school variable
```

# Adding value labels and making vars numeric 
```{r}
p_w1_clean_3 <- p_w1_clean_2 %>%
  mutate(condition = as.numeric(condition),
         condition = set_vall(condition, c("control" = 1, "intervention" = 2)),
         school_id = as.numeric(school_id),
         school_id = set_vall(school_id, c("cascade" = 1, "prairie_mountain" = 2, "kelly" = 3, "ata" = 4, "briggs" = 5, "agnes_stewart" = 6)))
```

# Creating individual id variable
```{r}
p_w1_clean_3 <- p_w1_clean_3 %>% 
  mutate(participant = ifelse(response_id == "R_1ON5POXTzIXrIeb", 4, participant), # changing partic. type to correspond to records
         participant_4_text = ifelse(response_id == "R_1ON5POXTzIXrIeb", "stepmom", participant_4_text), 
         participant = as.numeric(participant),
         participant = set_vall(participant, c("mom" = 1, "dad" = 2, "tutor legal" = 3, "other, specify" = 4)),
         participant_id = paste0(id, "_", participant), .before = 5) %>% # new var as unique identifier
  rename(c("family_id" = "id")) 
```

# renaming vars to correspond to value labels 
```{r}
# item that asks about # of adults in home is spreaded out and had an unusual coding.
p_w1_clean_3 <- p_w1_clean_3 %>%
  rename(c("q211_1" = "q211_2"),
         c("q211_2" = "q211_99"),
         c("q211_3" = "q211_4"),
         c("q211_4" = "q211_5"),
         c("q211_5" = "q211_6"),
         c("q211_99" = "q211_7"),
         c("q211_6_text" = "q211_6_text"))
```

```{r}
view_df(p_w1_clean_3)
```

# Number of participants in Wave 1
```{r}
# 125 total participants in w1, from which...

p_w1_clean_3 %>%
  filter(participant == 1) # 91 were moms

p_w1_clean_3 %>%
  filter(participant == 2) # 32 were dads

p_w1_clean_3 %>%
  filter(participant == 4) # 2 were "other" (1 stepmom, 1 grandparent)

unique(p_w1_clean_3$family_id) # 96 unique id values, thus 96 families in W1. 
```

# ------ Notes on W1 dataset -------
# participants ids 415 and 418 had 3 records each. 
# participant id 415 was a family of three: 2 moms, 1 dad. One of the moms was recoded as "other, specify" in the `participant` variable and "stepmom" in the `participant_4_text` variable. 
# Participant id 418 was "triplicated" and there was no "mom" for id 408 that was only a "dad". This was fixed.
# # Total participants in wave n = 125
# ------------------------------------

# keeping only parent & student linked data
```{r}
p_w1_clean_3 %>%
  filter(family_id == "106") # participant data complete, W1 student data is on W2 dataset

p_w1_clean_3 %>%
  filter(participant_id == "301_2")# participant dropped from study and is -99 throughout. Id not present in w2 and w3.

p_w1_clean_3 %>%
  filter(family_id == "504") # participant data complete, but no student data for wave 1


  
p_w1_clean_4 <- p_w1_clean_3 %>%
  filter(family_id != 301) %>% # deleted bec participant dropped from study and is -99 throughout. Id not present in w2 and w3.
  filter(family_id != 504) %>% # deleted because no corresponding youth data on W1, despite logs indicating youth completed W1 and W2.
  arrange(participant_id) # this is the cleaned complete parent wave 1 dataset

data.frame(table(p_w1_clean_4$family_id)) 

unique(p_w1_clean_4$family_id) # 94 parents

```

# Number of participants in Wave 1 (with matching youth ids)
```{r}
# 123 total participants in w1, from which...

p_w1_clean_4 %>%
  filter(participant == 1) # 90 were moms

p_w1_clean_4 %>%
  filter(participant == 2) # 31 were dads

p_w1_clean_4 %>%
  filter(participant == 4) # 2 were "other" (1 stepmom, 1 grandparent)

unique(p_w1_clean_4$family_id) # 96 unique id values, thus 96 families in W1. 
```




```{r}
y_w1_clean_5_test <- y_w1_clean_5 %>%
  rename(family_id = id)
```


```{r}
parent_youth_w1 <- left_join(p_w1_clean_4, y_w1_clean_5_test, by = "family_id")
```

