---
title: "youth_3_waves"
author: "Alejandra Garcia Isaza"
date: "7/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rio)
library(here)
library(tidyverse)
library(haven)
library(janitor)
library(knitr)
library(surveytoolbox)
library(sjPlot)
library(kableExtra)
library(psych)
library(sjmisc)
```

################## WAVE 1 ##################

# Loading the data & initial cleaning
```{r}
y_w1_raw <- read_sav(here("nopublish", "Youth W1  SPAN_February 10, 2021_10.31.sav"))


y_w1_raw <- y_w1_raw %>% 
  arrange(H1_1_TEXT) %>% # ordering participants ids
  janitor::clean_names() %>% # cleaning names
  select(-h1) %>% 
  rename(c("id" = "h1_1_text"),
         c("wave" = "h5"))

y_w1_raw %>%
  filter(finished == 0) # 508 completely empty (2 records, one complete) 624 99% complete

data.frame(table(y_w1_raw$id)) # 508 duplicated: one false start, one completed.
```

# Further cleaning
```{r}
y_w1_clean <- y_w1_raw %>% 
  select(-1:-8, -10:-17, -20, -21) %>%
  filter(response_id != "R_2wMlBAGtO7lzpRo") %>% # deleting 1 duped row "false start" (id 508)
  arrange(id)

duplicated(y_w1_clean$id)
```

# Fixing wrong condition assignment and school id
```{r}
y_w1_clean_2 <- y_w1_clean %>%
  mutate(school_id = str_sub(id, 1, 1),
         condition = case_when(
           school_id == "1" | school_id == "3" | school_id == "5" ~ "1",
           school_id == "2" | school_id == "4" | school_id == "6" ~ "2"
         )) %>%
  select(school_id, condition, everything()) %>%
  select(- h8) # deleting school var that has wrong assignment
```

# Adding value labels, making vars numeric, and recoding response option in q38
```{r}
y_w1_clean_3 <- y_w1_clean_2 %>%
  mutate(condition = as.numeric(condition),
         condition = set_vall(condition, c("control" = 1, "intervention" = 2)),
         school_id = as.numeric(school_id),
         school_id = set_vall(school_id, c("cascade" = 1, "prairie_mountain" = 2, "kelly" = 3, "ata" = 4, "briggs" = 5, "agnes_stewart" = 6)),
         q38 = ifelse(q38 == 99, 88, q38),
         q38 = set_varl(q38, "In the last 3 months, how frequently has at least one of your parents/guardians… …attended a parent group meeting."),
         q38 = set_vall(q38, c("Never" = 1, "Rarely" = 2, "Sometimes" = 3, "Often" = 4, "I don't know" = 88))) # q38 didn't have a 99 = no response option. 
```


```{r}
# at school (should be 101_1)
q101_1_1 #yes
q101_1_0 #no
q101_1_99 #refuse to answer

# other than school (should be 101_2)
q101_2_1 #yes
q101_2_0 #no
q101_2_99 #refuse to answer

q101b #how stressful 

stress <- y_w1_clean_3 %>% 
  select(id, starts_with("q101_1"), starts_with("q102_2")) %>% 
  pivot_longer(
    cols = starts_with("q10"),
    names_to = c("item", "place", "response"),
    names_sep = "_",
    values_to = "stress_lev",
    values_drop_na = TRUE) %>% 
  filter(stress_lev == 1) %>%
  mutate(stress_lev = as.numeric(stress_lev))

get_dupes(stress, id, place, item)

# id 506 chose yes and no
stress_2 <- stress %>% 
  distinct(id, place, .keep_all = TRUE) # Option 1 = yes, was chosen because the {distinct} function keeps the first row 

# renaming vars
stress_3 <- stress_2 %>% 
  pivot_wider(names_from = place, values_from = response) %>%
  rename(c("101_1" = "1"),
         c("101_2" = "2")) %>%
  select(-item, -stress_lev) %>%
  pivot_longer(
    cols = starts_with("101_"),
    names_to = c("place"),
    values_to = "response",
    values_drop_na = TRUE)

stress_4 <- stress_3 %>%
  pivot_wider(names_from = place, values_from = response)

# test <- left_join(y_w1_clean_3, stress_3)
```

# note: evaluate if it makes sense doing all these because  I have to do all this process with each pair of variables (102_1, 102_2, 103_1, 103_2 ... 107_1, 107_2). 

# Renaming demo vars
```{r}
y_w1_clean_5 <- y_w1_clean_3 %>% # not y_w1_clean_4 for consistency w parent & educator
  rename(c("lang_english" = "q92_1"),
         c("lang_spanish" = "q92_2"),
         c("lang_indigenous" = "q92_3"),
         c("lang_other" = "q92_4"),
         c("lang_no_response" = "q92_99"),
         c("lang_indigenous_text" = "q92_3_text"),
         c("lang_other_text" = "q92_4_text"),
         c("spanish_comfort_home" = "q93"),
         c("spanish_comfort_school" = "q94"),
         c("spanish_comfort_friends" = "q95"),
         c("spanish_comfort_general" = "q96"),
         c("english_comfort_home" = "q97"),
         c("english_comfort_school" = "q98"),
         c("english_comfort_friends" = "q99"),
         c("english_comfort_general" = "q100"),
         c("gender_id" = "q231"),
         c("white" = "q232_1"),
         c("hispanic_latino_spanish" = "q232_2"),
         c("indigenous_americas" = "q232_3"),
         c("black_african_american" = "q232_4"),
         c("asian" = "q232_5"),
         c("american_indian_alaska_native" = "q232_6"),
         c("middle_eastern_north_african" = "q232_7"),
         c("native_hawaiian_pacific_islander" = "q232_8"),
         c("race_ethnicity_other" = "q232_9"),
         c("race_ethnicity_no_response" = "q232_99"),
         c("indigenous_americas_text" = "q232_3_text"),
         c("race_ethnicity_other_text" = "q232_9_text"),
         c("age" = "q233"), 
         c("birth_country" = "q234"),
         c("another_birth_country_text" = "q234_2_text"),
         c("age_first_moved_us" = "q235_1_text")) %>%
  select(-q235, -response_id) # deleted: q235 because they did not have meaningful info & response_id bec. no longer needed

unique(y_w1_clean_5$id) # 94 students
```


```{r}
view_df(y_w1_clean_3)
```

# ------ Notes on W1 dataset -------
# Participant id 508 was the only duplicated id and it was because one record was a "false start". 
# Participant id 612 appeared as if belonged to school 2 = prairie, this was incorrect and was fixed. 
# q38 didn't have a 99 = no response option so it didn't make sense to add it ex post facto. Prob. a survey coding error. 
# Total participants in wave n = 94
# ------------------------------------

################## WAVE 2 ##################

# Loading the data & initial cleaning
```{r}
y_w2_raw <- read_sav(here("nopublish", "Youth W2 SPAN_February 10, 2021_10.32.sav"))


y_w2_raw <- y_w2_raw %>% 
  arrange(H1_1_TEXT) %>% # ordering participants ids
  janitor::clean_names() %>% # cleaning names
  select(-h1) %>% 
  rename(c("id" = "h1_1_text"),
         c("wave" = "h5"))

data.frame(table(y_w2_raw$id)) # 106, 313, 314, 405 (tripli), 408, 412

y_w2_raw %>%
  filter(finished == 0) # 405 duplicated: one with 47% and other with 80% complete; 412 completely empty; remaining with 99% complete 

```

```{r}
y_w2_raw %>%
  filter(id == 106|id == 313|id == 314|id == 405| id == 408| id == 412)

y_w2_clean <- y_w2_raw %>% 
  select(-1:-8, -10:-17, -20, -21) %>%
  filter(response_id != "R_20My7Nw91c2Kh6r") %>% # deleting duped row "false start" (id 405, with 47%)
  filter(response_id != "R_3jebHOrmpEvVusr") %>% # deleting duped row "false start" (id 405, with 80%)
  filter(response_id != "R_2Y5PBgkaoarXoW5") %>% # deleting duped row "false start" (id 412, with 2%)
  filter(response_id != "R_12LVKPebRTdZRF3") %>% # deleting row bec. filled with -99 throughout (id 408)
  arrange(id)

y_w2_clean %>%
  filter(id == 106|id == 313|id == 314)
```

```{r}
# id 106
y_w1_clean_3 %>%
  filter(id == 106)

y_w2_raw %>%
  filter(id == 106)

# id 313
y_w1_clean_3 %>%
  filter(id == 313)

y_w2_raw %>%
  filter(id == 313)

# id 314
y_w1_clean_3 %>%
  filter(id == 314)

y_w2_raw %>%
  filter(id == 314)
```


id 106 duplicated: one is from wave 1 and the other from wave 2.

R_1o4JcWGSoGCYx8w = data from id 106 = from wave 1 --> need to send this data point to wave 1.  
R_1n6evLXEU4dKorq = data from id 106 = from wave 2 --> keep it in this dataset (w2). 

--

id 313 duplicated

R_VU96u3fLkvWZQtP is true 313
R_yBDOsopku8j9smR is 311
--

id 314 duplicated

R_5pckz7xWKcHSapH is true 314
R_3m44tVSOd22OdTN is 312


id 999? (R_1CHCJIPFkuEHBH2) = test run, delete row. 

id JCG? (R_SCw7VH3ueceflUl) = belongs to id 109. Assessor error inputting id. 


################## WAVE 3 ##################

# Loading the data & initial cleaning
```{r}
y_w3_raw <- read_sav(here("nopublish", "Youth W3 SPAN_February 10, 2021_11.05.sav"))


y_w3_raw <- y_w3_raw %>% 
  arrange(H1_1_TEXT) %>% # ordering participants ids
  janitor::clean_names() %>% # cleaning names
  select(-h1) %>% 
  rename(c("id" = "h1_1_text"),
         c("wave" = "h5"))

data.frame(table(y_w3_raw$id)) # 405 (tripli), 613, 620

y_w3_raw %>%
  filter(finished == 0) # 405 2 "false starts" one 99%, other 4%; 613 completely empty, 2%; 620 71% complete; remaining 99% complete

```

```{r}
y_w3_raw %>%
  filter(id == 405|id == 613|id == 620)

y_w3_clean <- y_w3_raw %>% 
  select(-1:-8, -10:-17, -20, -21) %>%
  filter(response_id != "R_10ryStLMAuTyIpH") %>% # deleting duped row "false start" (id 405, with 99%)
  filter(response_id != "R_PzESZy9gZPQEb9D") %>% # deleting duped row "false start" (id 405, with 4%)
  filter(response_id != "R_1rkc9GzcBpTf5Ap") %>% # deleting duped row "false start"  (id 613, with 2%)
  filter(response_id != "R_2RayAqcviYXNSM6") %>% # deleting duped row "false start"  (id 620, with 71%)
  arrange(id)

duplicated(y_w3_clean$id) # no duplicates on wave 3. 
```

id 603 appears as if belongs to school 2 = prairie