---
title: "youth"
author: "Alejandra Garcia Isaza"
date: "2/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rio)
library(here)
library(tidyverse)
library(haven)

library(surveytoolbox)
library(sjPlot)
```


```{r}
y_w1_raw <- haven::read_sav(here("nopublish", "Youth W1  SPAN_February 10, 2021_10.31.sav"))

y_w1_raw <- y_w1_raw %>% 
  arrange(H1_1_TEXT) %>% # ordering participants ids
  janitor::clean_names() %>% # cleaning names
  select(-h1) %>% 
  rename(c("id" = "h1_1_text"))

y_w1_raw
```

```{r}
# checking duplicated Id
sum(duplicated(y_w1_raw$id)) # 1 duplicated ids

data.frame(table(y_w1_raw$id)) # id 508 is duped. One row complete, one row "false start"
```

```{r}
y_w1_clean <- y_w1_raw %>% 
  select(-1:-8, -10:-17, -20, -21) %>%
  filter(response_id != "R_2wMlBAGtO7lzpRo") %>% # deleting 1 duped row "false start" (id 508)
  arrange(id)

duplicated(y_w1_clean$id)

y_w1_clean
```

id 612 appeared as if belonged to school 2 = prairie, this was incorrect. 

```{r}
# fixing wrong assignment
y_w1_clean_2 <- y_w1_clean %>%
  mutate(school_id = str_sub(id, 1, 1),
         condition = case_when(
           school_id == "1" | school_id == "3" | school_id == "5" ~ "1",
           school_id == "2" | school_id == "4" | school_id == "6" ~ "2"
         )) %>%
  select(school_id, condition, everything()) %>%
  select(- h8) # deleting school var that has wrong assignment

y_w1_clean_2
```

```{r}
# making variables numeric and adding labels  
y_w1_clean_3 <- y_w1_clean_2 %>%
  mutate(condition = as.numeric(condition),
         condition = set_vall(condition, c("control" = 1, "intervention" = 2)),
         school_id = as.numeric(school_id),
         school_id = set_vall(school_id, c("cascade" = 1, "prairie_mountain" = 2, "kelly" = 3, "ata" = 4, "briggs" = 5, "agnes_stewart" = 6))) # this worked!!!

y_w1_clean_3

view_df(y_w1_clean_3) # nice plot that serves as codebook
```

```{r}
# exporting to .sav to check proper coding of condition and labels
# y_w1_clean_3 %>%
#   haven::write_sav(here("nopublish", "y_w1_clean_3.sav")) # it worked, the .sav file has all 94 participants with the proper scale and the proper level labels.
```

q38 has option 99 as "I don't know" and there is no "no response" option. Probably an error because other items in scale (q36 - q 40) have 88 = "I don't know" and 99 = "refuse to answer". 

Because code below changes the values of the data, I am leaving this item with the coding as is: 99 as "I don't know". It will anyways be converted as NA for scales. 

```{r}
# collapsing spread variables ---> no relevant variables need to be collapsed for this project. 

# y_w1_clean_4 <- y_w1_clean_3 %>%
#   mutate(q38 = as.numeric(q38),
#          q38 = set_vall(q38, c("Never" = 1, "Rarely" = 2, "Sometimes" = 3, "Often" = 4, "I don’t know" = 88)),
#          q38 = ifelse(q38 == 99, 88, q68))  # for some reason this completely changes the values of the data. 

# y_w1_clean_3$q38
# 
# y_w1_clean_4$q38

```


```{r}
# renaming demo vars

y_w1_clean_5 <- y_w1_clean_3 %>% # not y_w1_clean_4 for consistency w parent & educator
  rename(c("lang_english" = "q92_1"),
         c("lang_spanish" = "q92_2"),
         c("lang_indigenous" = "q92_3"),
         c("lang_other" = "q92_4"),
         c("lang_no_response" = "q92_99"),
         c("lang_indigenous_text" = "q92_3_text"),
         c("lang_other_text" = "q92_4_text"),
         c("spanish_comfort_home" = "q93"),
         c("spanish_comfort_school" = "q94"),
         c("spanish_comfort_friends" = "q95"),
         c("spanish_comfort_general" = "q96"),
         c("english_comfort_home" = "q97"),
         c("english_comfort_school" = "q98"),
         c("english_comfort_friends" = "q99"),
         c("english_comfort_general" = "q100"),
         c("gender_id" = "q231"),
         c("white" = "q232_1"),
         c("hispanic_latino_spanish" = "q232_2"),
         c("indigenous_americas" = "q232_3"),
         c("black_african_american" = "q232_4"),
         c("asian" = "q232_5"),
         c("american_indian_alaska_native" = "q232_6"),
         c("middle_eastern_north_african" = "q232_7"),
         c("native_hawaiian_pacific_islander" = "q232_8"),
         c("race_ethnicity_other" = "q232_9"),
         c("race_ethnicity_no_response" = "q232_99"),
         c("indigenous_americas_text" = "q232_3_text"),
         c("race_ethnicity_other_text" = "q232_9_text"),
         c("age" = "q233"), 
         c("birth_country" = "q234"),
         c("another_birth_country_text" = "q234_2_text"),
         c("age_first_moved_us" = "q235_1_text")) %>%
  select(-q235) # deleted: q235 because they did not have meaningful info

view_df(y_w1_clean_5)

```


```{r}
# exporting to .sav to check proper coding of condition and labels
# y_w1_clean_5 %>%
#   haven::write_sav(here("nopublish", "y_w1_clean_5.sav")) # it worked, variable and value labels were succesfully exported. # export on 02/24/21
```


# Notes on Missing values

77 = "doesn't apply"
88 = "I don't know"
99 = "no response"
-99 = seen but unanswered questions -- (e.g questions that ask participants to respond if they have knowledge about classroom)

```{r}
# recoding missing values as N/A with function

# vector with missing values in dataset
missing_vals <- c(77, 88, 99, -99)

# function that returns true if values in vector are equal to missing_vals. The function takes a vector x, and specified values of missing data
recode_missing <- function(x, missing_vals = c(77, 88, 99, -99)) {
  test <- x %in% missing_vals
  ifelse(test, NA, x)
}

# function that recodes missing values to NA. The function takes a dataframe with variables with missing data, and specified values of missing data
recode_missing_df <- function(df, missing_vals = c(77, 88, 99, -99)) {
  modify(df, ~recode_missing(.x, missing_vals)) # here uses the function created above
}

y_w1_clean_6 <- recode_missing_df(y_w1_clean_5)
```

# Scales

## homework involvement --> NOT using these.
q6-q13 and q15

## other parent-youth education involvement vars --> NOT using these.
q14, q16, q17

## parent support of youth with school & community involvement 
q18-q22

## parent-youth communication about school and ed --> NOT using these.
q23-q35 #####

## youth's perception of parent involvement
q36-q40 (1-4, never - often) higher score, more PI

## youth's involvement in school & community --> NOT using these.
q41-q44

## teacher-student relationship
q45-q55 (1-4, disagree - agree) higher score, better rel.


## youth academic self-efficacy (?) --> NOT using these.
q57-q67 

## school climate
q68-q82 (1-4 disagree - agree, but q81 needs to be reverse coded) higher score, better climate

## school engagement --> NOT using these.
q83-q91 (several need to be reverse coded)

## parent-youth relationship 
q127-q132 (1-4, never - often) higher score, better rel

## parent-youth difficult conversations --> NOT using these.
q136-q143

## monitoring --> NOT using these.
q144-q149 

## discipline --> NOT using these.
q150-q160

## youth behavior --> NOT using these.
q161-q171

## others --> NOT using these.
SDQ, hope, peers, substance use


# Common guidelines for evaluating Cronbach's Alpha
.00 to .69 = Poor
.70 to .79 = Fair 
.80 to .89 = Good 
.90 to .99 = Excellent/Strong


```{r}
y_w1_clean_6_rev_code <- y_w1_clean_6 %>%
  mutate(q81 = likert_reverse(q81, top = 4, bottom = 1))
```


```{r}
####### internal consistency ######

library(psych)

par_supp_scho <- data.frame(y_w1_clean_6_rev_code[,26:30]) #q18-q22
alpha(par_supp_scho) # alpha = .72 --> fair consistency

par_commun_scho <- data.frame(y_w1_clean_6_rev_code[,31:43]) #q23-q35
alpha(par_commun_scho) # alpha = .88 --> good consistency

par_inv <- data.frame(y_w1_clean_6_rev_code[,44:48]) #q36-q40
alpha(par_inv) # alpha = .68 --> poor consistency #### not using this one. 

youth_teach_rel <- data.frame(y_w1_clean_6_rev_code[,53:63]) #q45-q55
alpha(youth_teach_rel) # alpha = .91 --> Excellent consistency

climate <- data.frame(y_w1_clean_6_rev_code[,75:89]) #q68-q82
alpha(climate) # alpha = .92 --> Excellent consistency

youth_par_rel <- data.frame(y_w1_clean_6_rev_code[,179:184]) #q127-q132
alpha(youth_par_rel) # alpha = .76 --> fair consistency
```

```{r}
####### scales ######

y_w1_scales <- y_w1_clean_6_rev_code %>%
  rowwise() %>% 
  mutate(parent_support_school = mean(c(q18, q19, q20, q21, q22), na.rm = TRUE),
         parent_communic_school = mean(c(q23, q24, q25, q26, q27, q28, q29, q30, q31, q32, q33, q34, q35), na.rm = TRUE),
         youth_teacher_rel = mean(c(q45, q46, q47, q48, q49, q50, q51, q52, q53, q54, q55), na.rm = TRUE),
         school_climate = mean(c(q68, q69, q70, q71, q72, q73, q74, q75, q76, q77, q78, q79, q80, q81, q82), na.rm = TRUE),
         youth_parent_rel = mean(c(q127, q128, q129, q130, q131, q132), na.rm = TRUE)) %>%
  select(1:7, 300:304)

final_y_w1 <- left_join(y_w1_clean_5, y_w1_scales)

view_df(final_y_w1)
```

