---
title: "youth_wave_1"
author: "Alejandra Garcia Isaza"
date: "4/31/2022"
output: html_document
---
Note on April 1, 2022: I copied the code below from "youth_3_waves.Rmd" which I think is the doc I was working on the latest for cleaning. Here I will only work on youth wave 1 (baseline) for dissertation and will update progress made here to the three waves Rmd to continue cleaning.

Note on November 3, 2022: I edited this code to complete the cleaning of youth wave 2. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rio)
library(here)
library(tidyverse)
library(haven)
library(janitor)
library(knitr)
library(surveytoolbox)
library(sjPlot)
library(kableExtra)
library(psych)
library(sjmisc)
```

################## WAVE 1 ##################

# Loading the data & initial cleaning
```{r}
y_w1_raw <- read_sav(here("nopublish", "Youth W1  SPAN_February 10, 2021_10.31.sav"))


y_w1_raw <- y_w1_raw %>% 
  arrange(H1_1_TEXT) %>% # ordering participants ids
  janitor::clean_names() %>% # cleaning names
  select(-h1) %>% 
rename(c("id" = "h1_1_text"),
         c("sex" = "h2"),
         c("wave" = "h5"),
         c("int_location" = "h6"),
         c("int_location_4_text" = "h6_4_text"))
```

# Checking incomplete rows & duplicated ids
```{r}
y_w1_raw %>%
  filter(finished == 0) # 508 completely empty (2 records, one false start, one complete) 624 99% complete

y_w1_raw %>% 
  filter(wave == 2) # no rows from w2

y_w1_raw %>% 
  filter(wave == 3) # no rows from w3

data.frame(table(y_w1_raw$id)) # id 508 duped
```

# Further cleaning
```{r}
y_w1_clean <- y_w1_raw %>% 
  select(-1:-8, -10:-17, -20, -21) %>%
  filter(response_id != "R_2wMlBAGtO7lzpRo") %>% # deleting 1 duped row "false start" (id 508)
  arrange(id)

duplicated(y_w1_clean$id)
```

# Fixing wrong condition assignment and school id
```{r}
y_w1_clean_2 <- y_w1_clean %>%
  mutate(school_id = str_sub(id, 1, 1),
         condition = case_when(
           school_id == "1" | school_id == "3" | school_id == "5" ~ "1",
           school_id == "2" | school_id == "4" | school_id == "6" ~ "2"
         )) %>%
  select(school_id, condition, everything()) %>%
  select(- h8) # deleting school var that has wrong assignment
```

# Adding value labels, making vars numeric, and recoding response option in q38
```{r}
y_w1_clean_3 <- y_w1_clean_2 %>%
  mutate(condition = as.numeric(condition),
         condition = set_vall(condition, c("control" = 1, "intervention" = 2)),
         school_id = as.numeric(school_id),
         school_id = set_vall(school_id, c("cascade" = 1, "prairie_mountain" = 2, "kelly" = 3, "ata" = 4, "briggs" = 5, "agnes_stewart" = 6)),
         q38 = ifelse(q38 == 99, 88, q38),
         q38 = set_varl(q38, "In the last 3 months, how frequently has at least one of your parents/guardians… …attended a parent group meeting."),
         q38 = set_vall(q38, c("Never" = 1, "Rarely" = 2, "Sometimes" = 3, "Often" = 4, "I don't know" = 88))) # q38 didn't have a 99 = no response option. 
```


```{r}
unique(y_w1_clean_3$id) # 94 students

data.frame(table(y_w1_clean_3$id)) # no dupes
```

# importing id 106 from youth *wave 2 dataset* to wave 1 dataset
```{r}
just_106 <- read_sav(here("nopublish", "just_106.sav"))
```

# joining id 106 from *wave 2* to wave 1 (where it belongs)
```{r}
y_w1_clean_4 <- full_join(y_w1_clean_3, just_106) %>% 
  arrange(id) # this is the cleaned complete youth wave 1 dataset
```

# Number of participants in Wave 1
```{r}
data.frame(table(y_w1_clean_4$id)) # no duped ids

unique(y_w1_clean_4$id) # 95 youth
```


```{r}
view_df(y_w1_clean_4)
```


```{r}
# y_w1_clean_4 %>%
#   haven::write_sav(here("nopublish", "all_youth_wave1.sav"))
```

# ------ Notes on W1 dataset -------
# Participant id 508 was the only duplicated id and it was because one record was a "false start". 
# Participant id 612 appeared as if belonged to school 2 = prairie, this was incorrect and was fixed. 
# q38 didn't have a 99 = no response option so it didn't make sense to add it ex post facto. Prob. a survey coding error. 
# Total participants in wave n = 95
# ------------------------------------

##############################################
# keeping only parent & student linked data
##############################################

```{r}
y_w1_clean_4 %>%
  filter(id == "301") # no id found in youth dataset, parent (dad) dropped from study

y_w1_clean_4 %>%
  filter(id == "504") # no id found in youth dataset
y_w1_clean_4 %>%
  filter(response_id == "R_2tnelq8IWX0oN9z") # response id tied to id 504 in call logs for W1. Youth completed W1 and W2. Remains a mystery what happened to wave 1 youth data. 

y_w1_clean_4 %>%
  filter(id == "623") # id present in students (complete data), not in parents bec dad dropped

y_w1_clean_5 <- y_w1_clean_4 %>%
  filter(id != 623)# deleted bec dad never completed W1 and was dropped from study (confirmed in logs). Students = 94
```

# cleaning names for joining
```{r}
# adding "y1" to each variable name: "y" stands for youth, "1" for wave 1.
names(y_w1_clean_5) <- paste(names(y_w1_clean_5), "y1", sep = "_")

y_w1_clean_6 <- y_w1_clean_5 %>%
  rename(c("school_id" = "school_id_y1"),
         c("condition" = "condition_y1"),
         c("family_id" = "id_y1")) %>% # renaming these variables to be able to join later
  select(-response_id_y1)

view_df(y_w1_clean_6)

# y_w1_clean_6 %>%
#   haven::write_sav(here("nopublish", "y_w1_clean_6.sav")) 
```
