---
title: "youth_wave_2"
author: "Alejandra Garcia Isaza"
date: "11/03/2022"
output: html_document
---
Note on November 3, 2022: I copied the Rmd called "youth_wave_1.Rmd". This document only contains wave 2 youth data. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rio)
library(here)
library(tidyverse)
library(haven)
library(janitor)
library(knitr)
library(surveytoolbox)
library(sjPlot)
library(kableExtra)
library(psych)
library(sjmisc)
```

################## WAVE 2 ##################

# Loading the data & initial cleaning
```{r}
y_w2_raw <- read_sav(here("nopublish", "Youth W2 SPAN_February 10, 2021_10.32.sav"))


y_w2_raw <- y_w2_raw %>% 
  arrange(H1_1_TEXT) %>% # ordering participants ids
  janitor::clean_names() %>% # cleaning names
  select(-h1) %>% 
  rename(c("id" = "h1_1_text"),
         c("sex" = "h2"),
         c("wave" = "h5"),
         c("int_location" = "h6"),
         c("int_location_4_text" = "h6_4_text"))
```

# Checking incomplete rows & duplicated ids
```{r}
y_w2_raw %>%
  filter(finished == 0) # 8 records incomplete, but five are 99% (202, 306, 404, 610, 613), one 80% 405), and two less than 50% (ids 405 % 412)

y_w2_raw %>% 
  filter(wave == 1) # id 106 "R_1o4JcWGSoGCYx8w" says w1 --> confirmed in logs: this is youth data for w1. Action: need to move to wave 1. 

y_w2_raw %>% 
  filter(wave == 3) # no rows from w3

data.frame(table(y_w2_raw$id)) # 106(one belongs to w1), 313, 314, 405(tripli), 408(one row -99 throughout prb false start, other row is complete), 412, JCG(assesor error)
```


# Further cleaning -- leaving out rows (false starts and coding errors)
```{r}
y_w2_clean <- y_w2_raw %>% 
  select(-1:-8, -10:-17, -20, -21) %>%
  filter(response_id != "R_20My7Nw91c2Kh6r") %>% # deleting duped row "false start" (id 405, with 47%)
  filter(response_id != "R_3jebHOrmpEvVusr") %>% # deleting duped row "false start" (id 405, with 80%)
    filter(response_id != "R_12LVKPebRTdZRF3") %>% # deleting row bec. filled with -99 throughout (id 408)
  filter(response_id != "R_2Y5PBgkaoarXoW5") %>% # deleting duped row "false start" (id 412, with 2%)
  filter(response_id != "R_1CHCJIPFkuEHBH2") %>% # deleting test row (id 999, with 2%)
  arrange(id)

data.frame(table(y_w2_clean$id)) # 106, 313, 314, JCG
```


```{r}
y_w2_clean_1 <- y_w2_clean %>% 
   mutate(id = case_when(response_id == "R_yBDOsopku8j9smR" ~ "311", # id 313 recoded as 311
                        response_id == "R_3m44tVSOd22OdTN" ~ "312", # id 314 recoded as 312
                        response_id == "R_SCw7VH3ueceflUl" ~ "109", # id JCG recoded as 109
                        TRUE ~ as.character(id))) %>%
  arrange(id)

data.frame(table(y_w2_clean_1$id)) # only 106 is dupped bec one belongs to wave 1
```


# Fixing wrong condition assignment and school id
```{r}
y_w2_clean_2 <- y_w2_clean_1 %>%
  mutate(school_id = str_sub(id, 1, 1),
         condition = case_when(
           school_id == "1" | school_id == "3" | school_id == "5" ~ "1",
           school_id == "2" | school_id == "4" | school_id == "6" ~ "2"
         )) %>%
  select(school_id, condition, everything()) %>%
  select(- h8) # deleting school var that has wrong assignment
```

# Adding value labels, making vars numeric, and recoding response option in q38
```{r}
y_w2_clean_3 <- y_w2_clean_2 %>%
  mutate(condition = as.numeric(condition),
         condition = set_vall(condition, c("control" = 1, "intervention" = 2)),
         school_id = as.numeric(school_id),
         school_id = set_vall(school_id, c("cascade" = 1, "prairie_mountain" = 2, "kelly" = 3, "ata" = 4, "briggs" = 5, "agnes_stewart" = 6)),
         q38 = ifelse(q38 == 99, 88, q38),
         q38 = set_varl(q38, "In the last 3 months, how frequently has at least one of your parents/guardians… …attended a parent group meeting."),
         q38 = set_vall(q38, c("Never" = 1, "Rarely" = 2, "Sometimes" = 3, "Often" = 4, "I don't know" = 88))) # q38 didn't have a 99 = no response option.
```

# for joining id 106 that belongs to wave 1 and was mistakenly put in wave 2
```{r}
y_w2_clean_4 <- y_w2_clean_3 %>%
  filter(response_id != "R_1o4JcWGSoGCYx8w") # this is the youth wave 2 cleaned dataset (without row of w1 data from id 106)
  
just_106 <- y_w2_clean_3 %>%
  filter(response_id == "R_1o4JcWGSoGCYx8w") # this is just id 106

# just_106 %>%
#   haven::write_sav(here("nopublish", "just_106.sav"))
```

# Number of participants in Wave 2
```{r}
unique(y_w2_clean_4$id) # 81 students

data.frame(table(y_w2_clean_4$id)) # no dupes
```

```{r}
view_df(y_w2_clean_4)
```


```{r}
# y_w2_clean_4 %>%
#   haven::write_sav(here("nopublish", "all_youth_wave2.sav"))
```

